---
title: "Informational Power Through Pivotality: How Consensus Can Benefit a Hegemon"
author: "Manoel Galdino (University of São Paulo; mgaldino@usp.br)"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    number_sections: TRUE
    toc: FALSE
    latex_engine: xelatex
header-includes:
  - \usepackage{setspace}
  - \setstretch{1.5}
  - \usepackage{parskip}
  - \setlength{\parskip}{6pt}
  - \setlength{\parindent}{0pt}
  - \usepackage{float}
  - \usepackage{amsmath,amssymb,amsthm}
  - \newtheorem{definition}{Definition}
  - \newtheorem{lemma}{Lemma}
  - \newtheorem{proposition}{Proposition}
  - \newtheorem{theorem}{Theorem}
  - \newtheorem{corollary}{Corollary}
  - \newtheorem{remark}{Remark}
  - \newtheorem{example}{Example}
  - \usepackage{tikz}
  - \usetikzlibrary{arrows.meta,positioning,calc}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{pdflscape}
  - \usepackage{array}
fontsize: 12pt
geometry: margin=2.5cm
mainfont: Times New Roman
bibliography: references.bib
link-citations: true
keywords: "International Organizations, Bargaining, Hegemon, Informational Power, Screening, Consensus"
abstract: |
  When can consensus benefit a powerful state even without granting it proposal power? I study a two-round bargaining game in which only weak states propose, the hegemon privately knows its outside payoff, and a no vote by the hegemon is an immediate and irreversible opt-out. Under unanimity, weak proposers must satisfy the hegemon's private participation constraint. In the regular interior domain, a unanimity PBE exists only under derived incentive and belief conditions; whenever it exists, every regular on-path equilibrium pools at the high outside payoff. Under majority rule, equilibrium is generally set-valued: weak coalitions can exclude the hegemon, but larger groups can also sustain separating current passage. Institutional comparisons are therefore conditional on the common PBE-existence domain. On that domain, weak-state formation under unanimity is nested in formation under majority, while unanimity weakly raises the hegemon's payoff whenever both institutions form, with equality under majority pooling. Consensus can thus create informational power through pivotality even when agenda power is deliberately assigned to weaker states.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
history_path <- "tables/clean_optout_gate0_histories_piH0.tsv"
gate0_histories <- utils::read.delim(
  history_path,
  check.names = FALSE,
  quote = "",
  comment.char = "",
  stringsAsFactors = FALSE
)
```

\newpage

# Introduction

When can consensus benefit a powerful state? A control-based intuition says
that majority rule should be more attractive to powerful actors because a
winning coalition can act without unanimous consent. Consensus seems to move
power in the opposite direction by giving every participant a veto.

This paper identifies a different source of power. Consensus can favor a
privately informed hegemon because it makes that actor's approval an
informational constraint on weaker states. Weak states know the institutional
package they can offer but not the hegemon's privately known outside payoff.
When the hegemon is pivotal, the weak proposer must decide whether and how to
insure agreement against that hidden participation constraint.

The model separates outside-option power, pivotality, and agenda power. The
hegemon \(H\) is never recognized as proposer: \(\pi_H=0\) in both bargaining
rounds. Only weak states propose. A proposal is a relative institutional
package \(y\), such as a quota share, production flexibility, an exception, or
an enforcement advantage. It is feasible in every state and reduces the unit
surplus left to the weak coalition one-for-one. If \(H\) votes no in Round 1,
it opts out immediately and irreversibly, receives \(o_\theta\) without
discounting, and cannot re-enter. If \(H\) votes yes but simultaneous weak
votes make the ballot fail, \(H\) remains active and the game can continue.

The clean timing changes the equilibrium architecture. First, terminal
unanimity retains the familiar screening choice between a low-only package and
pooling. Second, in the regular interior domain, Round-1 unanimity does not
admit a global pooling/low-only/delay reduction. A global PBE exists only if
\(\beta o_1\geq o_0\) and a strict security-value condition holds; whenever
it exists, every regular on-path equilibrium pools at \(y=o_1\). Third,
majority is not a unique no-screening benchmark. Its exact security value is
the maximum of exclusion, low-only, and pooling guarantees, and its equilibrium
correspondence depends on group size. Finally, institutional comparisons are
made only on the common PBE-existence domain. There, weak-state entry under
unanimity is nested in entry under majority, and unanimity weakly benefits
\(H\) whenever both institutions form.

These are conditional results, not claims that unanimity always dominates
majority or that the model characterizes all PBEs under unrestricted beliefs.
The analysis uses perfect Bayesian equilibrium together with a
weak-vote-passive assessment: an uninformed weak voter's unexpected ballot is
not treated as direct evidence about \(H\)'s type. The assessment does not
impose sequential-equilibrium consistency across globally off-path
information sets.

# Consensus, Information, and Institutional Design {#literature}

The argument contributes first to rational-design accounts of international
institutions. Existing work emphasizes how institutional rules help states
manage uncertainty, commitment, and distributional conflict
[@koremenos2001rational]. Consensus is often interpreted as a device for
protecting participants from unfavorable outcomes or encouraging broad
participation [@gould2016consensus]. The mechanism here reverses the usual
direction of the informational benefit: uncertainty can make consensus
valuable to the privately informed pivotal actor.

Second, the paper qualifies self-binding and informal-power accounts.
Self-binding theories explain why powerful states may accept constraints to
make cooperation credible [@keohane1984after; @ikenberry2001after].
Informal-power accounts show how powerful states can dominate outcomes even
under formally egalitarian rules [@steinberg2002shadow; @stone2011controlling;
@gruber2000ruling]. Here the hegemon neither proposes nor needs to control
drafting. Its leverage comes from a private outside payoff that becomes
politically productive when its approval is necessary.

Third, the model builds on bargaining theories in which recognition and voting
rules determine distribution [@baron1989bargaining; @kalandrakis2006proposal;
@eraslan2019legislative]. Setting \(\pi_H=0\) removes formal proposal power
from the hegemon and isolates pivotality. The informational problem is related
to veto bargaining [@bardhi2018modes; @kim2025persuasion], but there is no
information designer or commitment to an experiment. Information matters
because an institutional quota determines whether a privately informed vote
is necessary for implementation.

Two recent papers provide close but distinct comparisons. @piazolo2025legislative
study a setting in which privately informed responders can signal through
rejection and bargaining may break down. @glynia2026unanimity analyze a
one-shot ultimatum environment with a fixed proposer and a trade-off between
insuring unanimity and gambling on a majority coalition. The present paper
asks whether consensus can benefit a powerful non-proposer. Its distinctive
features are weak-state agenda control, an immediate and irreversible opt-out
by the hegemon, and an explicitly conditional comparison over equilibrium
correspondences.

# A Motivating Terminal Example {#example}

Consider a terminal bargaining round with one hegemon and two weak states.
The weak coalition has a unit surplus. A higher package \(y\) benefits \(H\)
and reduces the weak residual one-for-one. The low type has outside payoff
\(o_0\), the high type has outside payoff \(o_1\), and
\(0<o_0<o_1<1\). Because implementation is certain if \(H\) votes yes, type
\(\theta\) accepts exactly when \(y\geq o_\theta\).

Under unanimity, a weak proposer compares a low-only package \(y=o_0\), worth
\((1-\mu)(1-o_0)\), with a pooling package \(y=o_1\), worth \(1-o_1\). It
pools when
\[
\mu>\frac{o_1-o_0}{1-o_0}.
\]
Pooling gives the low type an informational rent \(o_1-o_0\).

This terminal example is deliberately limited. In Round 1, ballots are
simultaneous, a no vote by \(H\) is an immediate opt-out, and a yes vote
followed by weak-caused failure leads to a distinct continuation. The direct
cutoff \(y=o_\theta\) therefore applies only when current implementation after
\(H\)-yes is certain. The full model derives the expected voting incentive
constraint rather than treating the terminal cutoff as a global rule.

# The Model {#model}

\begin{definition}[Clean immediate-opt-out game]\label{def:game}
There are \(N\geq3\) states: one hegemon \(H\) and \(m=N-1\) weak states.
Nature draws a persistent type \(\theta\in\{0,1\}\). Only \(H\) observes its
type; weak states share prior \(\mu=\Pr(\theta=1)\). The weak coalition has a
fixed institutional surplus normalized to one, and weak outside payoffs are
zero. The primitive domain is
\[
0\leq o_0<o_1\leq\bar y\leq1,
\qquad \beta\in(0,1].
\tag{1}\label{eq:primitive-domain}
\]
If a current agreement includes \(H\), its payoff is \(y\); hence
\(b_0=b_1=0\). The outside payoff \(o_\theta\) is external to the
institutional pie.

The game has two rounds. In every reached round, a new weak proposer is drawn
uniformly and independently; \(\pi_H=0\) in both rounds. If weak state \(i\)
proposes while \(H\) is active, its contract is
\[
s_i=(y,(x_j)_{j\ne i}),\qquad
y\in[0,\bar y],\quad x_j\geq0.
\]
If \(H\) participates, the proposer receives
\[
x_i^H=1-y-\sum_{j\ne i}x_j\geq0.
\tag{2}\label{eq:h-budget}
\]
If majority passage excludes \(H\), the conditional component \(y\) is
reabsorbed and the proposer receives
\[
x_i^{WO}=1-\sum_{j\ne i}x_j=x_i^H+y.
\tag{3}\label{eq:weak-only-budget}
\]
Thus \(H\)'s outside payoff is never charged to weak surplus, and \(H\) never
receives both \(y\) and \(o_\theta\) in the same history.

The proposer is counted as voting yes. Every other active player observes the
proposal and votes simultaneously; individual votes become public only after
the ballot closes. Unanimity requires \(N\) yes votes. Majority requires
\[
q=\left\lfloor\frac N2\right\rfloor+1.
\tag{4}\label{eq:quota}
\]
The protocol is identical across rules except for the quota and is not a
sequential roll call.

A no vote by type \(\theta\) of \(H\) is an immediate and irreversible
opt-out. In Round 1 it pays \(o_\theta\) immediately and without discounting.
No current or future agreement can then include \(H\). Under unanimity the
branch ends because the original quota \(N\) is impossible. Under majority, a
current weak-only agreement passes if the weak yes votes already reach \(q\);
otherwise the game proceeds to a weak-only Round 2 with the original quota
\(q\). If \(H\) votes yes and the Round-1 ballot fails only because weak votes
do not reach the quota, \(H\) remains active and the game proceeds to Round 2.
Round 2 is terminal, and all Round-2 payoffs are multiplied by \(\beta\) in
Round-1 units.

Before Round-1 recognition, weak states make one collective, all-or-nothing
formation decision and pay an external sunk cost \(\chi\) each if the
institution forms. For rule \(R\), define
\[
V_W^R(\mu)=\frac1m
\mathbb E\!\left[\sum_{j=1}^m u_j^R\mid\text{formation}\right].
\tag{5}\label{eq:entry-value}
\]
Formation occurs iff \(V_W^R(\mu)\geq\chi\), with formation at equality.
\end{definition}

The immediate-opt-out timing is summarized in Figure
\ref{fig:timing}. The distinction between an \(H\)-no and an \(H\)-yes
followed by weak rejection is the core protocol distinction.

\begin{figure}[H]
\centering
\resizebox{\textwidth}{!}{%
\begin{tikzpicture}[
  node distance=1.15cm and 1.05cm,
  box/.style={draw, rounded corners=1pt, align=center, inner sep=5pt, minimum width=2.45cm},
  branch/.style={draw, rounded corners=1pt, align=left, inner sep=5pt, minimum width=4.0cm},
  arrow/.style={-{Latex[length=2.1mm]}, thick}
]
\node[box] (nature) {Nature draws \(\theta\)\\\(H\) observes};
\node[box, right=of nature] (entry) {Collective weak\\formation};
\node[box, right=of entry] (proposal) {R1 weak proposer\\chooses \(s_i\)};
\node[box, right=of proposal] (ballot) {Simultaneous ballot\\votes later revealed};

\node[branch, right=1.4cm of ballot, yshift=1.45cm] (pass) {Quota reached with \(H\)-yes:\\current agreement includes \(H\)};
\node[branch, right=1.4cm of ballot] (optout) {\(H\)-no: immediate \(o_\theta\), no re-entry;\\\(U\) ends, \(M\) may pass or continue weak-only};
\node[branch, right=1.4cm of ballot, yshift=-1.45cm] (continue) {\(H\)-yes but weak-caused failure:\\\(H\) remains active and R2 follows};

\draw[arrow] (nature) -- (entry);
\draw[arrow] (entry) -- (proposal);
\draw[arrow] (proposal) -- (ballot);
\draw[arrow] (ballot) -- (pass);
\draw[arrow] (ballot) -- (optout);
\draw[arrow] (ballot) -- (continue);
\end{tikzpicture}
}
\caption{Timing of the clean immediate-opt-out baseline. Ballots are simultaneous. A no vote by \(H\) is an immediate and irreversible opt-out; a yes vote followed by weak-caused failure leaves \(H\) active for Round 2.}
\label{fig:timing}
\end{figure}

\begin{definition}[Weak-vote-passive assessment]\label{def:passive}
The solution concept is PBE. On-path beliefs obey Bayes' rule. A unilateral
vote deviation by an uninformed weak state is not treated as direct evidence
about \(\theta\); beliefs instead follow the public proposal and \(H\)'s
simultaneous type-contingent strategy and revealed vote. A separating vote by
\(H\) can update beliefs when \(H\) remains active.

Every zero-probability continuation in which \(H\) remains active must state
its off-path posterior. Globally off-path ballot beliefs and later
zero-probability continuation beliefs are components of weak PBE and need not
be linked by Bayes' rule. The assessment is not a refinement, D1, the
Intuitive Criterion, a sequential-equilibrium restriction, or a
characterization of all PBEs.

Players vote yes when indifferent. If a weak proposer is indifferent among
payoff-maximizing proposals, the selected proposal minimizes \(H\)'s expected
payoff. No third tie-break is imposed.
\end{definition}

For an R1 ballot information set \(I=(h,s_i)\), let \(v_W\) denote the
simultaneous weak-vote vector and \(A_R^Y(v_W)\) indicate implementation when
\(H\) votes yes. The correct incentive constraint is
\[
\sum_{v_W}\sigma_W(v_W\mid I)
\left[
A_R^Y(v_W)y+
\{1-A_R^Y(v_W)\}\beta C_{H,2}^R(\theta,h_2)
\right]\geq o_\theta.
\tag{6}\label{eq:h-ic-r1}
\]
The right-hand side is always the immediate opt-out payoff. Only when
\(A_R^Y(v_W)=1\) with probability one does this reduce to
\[
y\geq o_\theta.
\tag{7}\label{eq:direct-cutoff}
\]
Appendix A states the corresponding weak-voter IC and records every voting
history.

The main theorems use the regular interior domain
\[
\mu\in(0,1),\qquad \beta\in(0,1),\qquad 0<o_0<o_1<1.
\tag{8}\label{eq:regular-domain}
\]
The boundaries \(o_0=0\), \(o_1=1\), \(\beta=1\), and degenerate priors are
separated in Appendix C rather than imported by continuity.

\begin{table}[H]
\centering
\small
\begin{tabular}{p{0.28\textwidth}p{0.62\textwidth}}
\toprule
Object & Scope \\
\midrule
Terminal unanimity payoff & All-PBE payoff result in the terminal subgame. \\
Regular R1 unanimity & Global PBE existence and on-path payoff characterization under weak PBE and the declared completion beliefs. \\
Regular R1 majority & Pure on-path proposal characterization; tied mixtures inherit the stated payoff bounds. \\
Institutional comparison & Selection-free only on the derived common PBE-existence domain. \\
Boundaries and endpoints & Separate propositions or one-sided equilibrium-outcome limits; not continuations of the regular theorem by assumption. \\
\bottomrule
\end{tabular}
\caption{Scope of the clean baseline results.}
\label{tab:result-scope}
\end{table}

# Backward Induction {#results}

## Terminal Round

\begin{lemma}[Terminal unanimity]\label{lem:r2-unanimity}
At posterior \(\nu\), define
\[
P=1-o_1,\qquad
L(\nu)=(1-\nu)(1-o_0),\qquad
S(\nu)=\max\{P,L(\nu)\}.
\tag{9}\label{eq:r2-u-value}
\]
Every terminal unanimity PBE gives the weak coalition total payoff \(S(\nu)\).
The proposer offers \(y=o_0\) when
\[
\nu\leq\nu_2^*=\frac{o_1-o_0}{1-o_0}
\tag{10}\label{eq:r2-cutoff}
\]
and \(y=o_1\) above the cutoff. At equality, the proposal tie-break selects
low-only. Each weak state's ex ante Round-2 value is \(S(\nu)/m\).
\end{lemma}

\begin{proposition}[Terminal majority]\label{prop:r2-majority}
In terminal majority bargaining, all weak states approve. A weak proposer
sets named weak offers to zero and retains the unit pie whether \(H\) is active
or has already opted out. Therefore
\[
C_{W,2}^M=\frac1m,\qquad
c=\frac{\beta}{m},\qquad
C_{H,2}^M(\theta)=o_\theta.
\tag{11}\label{eq:r2-m-values}
\]
This result is selection-free, and \(H\)'s outside payoff remains external to
the weak pie.
\end{proposition}

The terminal screening result is the clean one-shot bridge. It does not imply
that Round-1 acceptance follows a global cutoff: equation
\eqref{eq:h-ic-r1} still integrates simultaneous weak votes and the
continuation after \(H\)-yes.

## Round 1 Under Unanimity

For the regular domain, write
\[
D_0=1-o_0,\quad L=(1-\mu)D_0,\quad
\delta=\frac{\beta(m-1)}m,\quad a=1-\delta,\quad
d=\frac{\beta P}{m}.
\tag{12}\label{eq:u-notation}
\]
If pooling is on path, Bayes' rule prices a weak-voter deviation at
\(\beta S(\mu)/m\), and the maximal proposer payoff is
\[
\pi_P=P-\delta S(\mu).
\tag{13}\label{eq:onpath-pooling}
\]
If low-only is on path, \(H\)-yes reveals the low type, so the maximal
proposer payoff is
\[
\pi_L=aL.
\tag{14}\label{eq:onpath-low}
\]

Globally off path, the proposer can guarantee two different security values:
\[
G_L=(1-\mu)(D_0-\delta P),
\qquad
G_P=aP.
\tag{15}\label{eq:u-security}
\]
The gap between the off-path low-only guarantee and its on-path payoff is
strict:
\[
G_L-\pi_L=(1-\mu)\delta(D_0-P)>0.
\tag{16}\label{eq:shaving-gap}
\]
This discontinuity is why the old global pooling/low-only/delay reduction does
not survive the clean game.

\begin{lemma}[Off-path completion and guarantees]\label{lem:completion}
Suppose the regular domain holds and \(\beta o_1\geq o_0\). There is a weak-PBE
completion that holds every globally off-path proposal to at most
\(\max\{G_L,G_P\}\). Conversely, the proposer can guarantee \(G_L\) with
\(y=o_0\) and \(x_j=d\) for every non-proposer, and can guarantee \(G_P\)
with \(y=o_1\) and the same named offers. If \(\beta o_1<o_0\), the feasible
proposal \(y=o_1,x_j=0\) has no sequentially rational ballot completion, so
the full game has no PBE.
\end{lemma}

Define
\[
D_U=D_0-\delta P>0,
\qquad
\mu_E=1-\frac{aP}{D_U}.
\tag{17}\label{eq:u-existence-cutoff}
\]

\begin{theorem}[Regular unanimity existence and payoff]\label{thm:u-regular}
On the regular interior domain, a global unanimity PBE exists if and only if
\[
\boxed{\beta o_1\geq o_0\quad\text{and}\quad G_P>G_L,}
\tag{18}\label{eq:u-existence}
\]
equivalently, \(\beta o_1\geq o_0\) and \(\mu>\mu_E\). At equality
\(G_P=G_L\), the proposal tie-break selects a lower-\(H\)-payoff deviation
whose on-path approval price jumps, so no fixed point exists.

Every regular unanimity PBE in the existence domain has the same on-path
outcome:
\[
y=o_1,\qquad
T_W^U=P,\qquad
V_W^U=\frac{P}{m},\qquad
E[u_H^U]=o_1.
\tag{19}\label{eq:u-regular-payoffs}
\]
\end{theorem}

Theorem \ref{thm:u-regular} is a pooling result under the clean immediate
opt-out contract. Low-only and rejection are not regular on-path equilibrium
classes. Rejection can reappear when \(\beta=1\), and low-only can reappear
when \(o_0=0\); Appendix C gives the exact boundary statements. Nothing in the
theorem asserts uniqueness of strategies or characterizes all PBEs under a
stronger consistency concept.

## Round 1 Under Majority

Let
\[
k=q-1,\qquad c=\frac\beta m,
\qquad
E=1-kc,
\qquad
B_M=(1-\mu)(1-o_0)+\mu c,
\qquad
F_M=\max\{E,B_M,P\}.
\tag{20}\label{eq:m-security}
\]
The three guarantees correspond to exclusion, a low-only offer, and pooling.
For a non-proposing weak voter,
\[
EU_j(Y)-EU_j(N)
=(x_j-c)\Pr(j\text{ is pivotal}).
\tag{21}\label{eq:m-weak-ic}
\]
The exact security value is \(F_M\), but implementability at that value
depends on group size.

\begin{proposition}[Regular majority correspondence]\label{prop:majority}
On the regular interior domain:

1. If \(N=3\), the proposer payoff is exactly \(F_M\). Exclusion is selected
   when \(E\) maximizes; low-only is selected when \(B_M\) maximizes; pooling
   requires \(P>\max\{E,B_M\}\). A tie \(E=B_M\) remains set-valued.
2. If \(N=4\), a majority PBE exists if and only if
   \[
   E\geq B_M\quad\text{or}\quad P>B_M.
   \tag{22}\label{eq:m-n4-existence}
   \]
   Exclusion obtains when \(E\geq\max\{B_M,P\}\), and pooling obtains when
   \(P>\max\{B_M,E\}\). The boundary \(B_M=P>E\) has no PBE.
3. If \(N\geq5\), a majority PBE always exists and the exact proposer-payoff
   set is
   \[
   V_{\mathrm{prop}}^M\in[F_M,1].
   \tag{23}\label{eq:m-large-value}
   \]
   Every regular pure on-path proposal is exclusion, separating current
   passage, or pooling. A separating proposal
   \(o_0\leq y<o_1\) has
   \[
   T_W^M=1-(1-\mu)y,\qquad
   E[u_H^M]=(1-\mu)y+\mu o_1.
   \tag{24}\label{eq:m-separating}
   \]

Across all existing regular majority PBEs,
\[
F_M\leq T_W^M\leq1,
\qquad
\frac{F_M}{m}\leq V_W^M\leq\frac1m.
\tag{25}\label{eq:m-weak-bounds}
\]
\end{proposition}

The historical No-Cheap-H condition also changes. Since
\[
B_M-E=kc-o_0+\mu(o_0+c-1),
\tag{26}\label{eq:no-cheap-gap}
\]
the uniform sufficient condition is
\[
o_0\geq kc=(q-1)\frac{\beta}{m}.
\tag{27}\label{eq:no-cheap}
\]
It selects exclusion for \(N=3\) and guarantees an exclusion equilibrium for
\(N=4\). For \(N\geq5\), it does not eliminate separating current passage.
Majority no-screening is therefore not an all-PBE result.

# Entry and Conditional Institutional Comparison {#entry}

The comparison is restricted to the common regular PBE domain. Conditions
\eqref{eq:u-existence} are required for unanimity. For \(N=4\), condition
\eqref{eq:m-n4-existence} is also required; for \(N=3\) and \(N\geq5\),
majority existence is automatic once the regular domain holds. These are
derived existence conditions rather than additional primitive assumptions.

\begin{proposition}[Entry nesting]\label{prop:nesting}
On the common PBE domain, every majority equilibrium satisfies
\[
V_W^M\geq\frac{F_M}{m}
\geq\frac{P}{m}
=V_W^U.
\tag{28}\label{eq:entry-nesting}
\]
Consequently, collective formation sets satisfy
\[
\mathcal F_U\subseteq\mathcal F_M
\tag{29}\label{eq:formation-nesting}
\]
selection-free. Given a selected majority payoff \(v_M=V_W^M\):

1. both institutions form if \(\chi\leq P/m\);
2. only majority forms if \(P/m<\chi\leq v_M\);
3. neither forms if \(\chi>v_M\).

There is no region in which only unanimity forms.
\end{proposition}

\begin{proposition}[Hegemon payoff bounds]\label{prop:h-comparison}
Let
\[
\bar o=(1-\mu)o_0+\mu o_1.
\]
On the common PBE domain, every regular majority equilibrium satisfies
\[
\bar o\leq E[u_H^M]\leq o_1=E[u_H^U].
\tag{30}\label{eq:h-payoff-bounds}
\]
Thus, conditional on both institutions forming, unanimity weakly benefits
\(H\), with equality under majority pooling. If only majority forms,
majority weakly benefits \(H\) relative to nonformation under unanimity. If
neither forms, \(H\) receives \(\bar o\) under both rules.
\end{proposition}

Propositions \ref{prop:nesting} and \ref{prop:h-comparison} provide an
institutional classification, not global dominance. They do not collapse the
majority correspondence to a scalar payoff and do not compare rules at
parameter vectors where one of the required PBEs does not exist.

# Discussion {#discussion}

## OPEC and pivotal participation {#opec}

The OPEC interpretation is institutional rather than monetary. The package
\(y\) can represent a quota share, production flexibility, an exception, or
an enforcement arrangement that benefits a pivotal producer and reduces the
residual institutional surplus available to other members. The private object
\(o_\theta\) is the producer's outside payoff from opting out: it can summarize
opportunity costs, alternative production strategies, spare capacity, and
other privately known advantages of nonparticipation.

The model does not claim that a literal OPEC meeting implements the exact
two-round protocol. It isolates a mechanism: when every agreement needs the
pivotal producer's approval, weaker members must insure against a private
participation constraint. In the regular existence region, this pressure
produces pooling at the high outside payoff even though the hegemon never
proposes. Under majority, weak coalitions can exclude \(H\), but the
equilibrium correspondence can also include separating current passage when
the group is large enough. The empirical implication is therefore not that
majority mechanically removes information rents; it is that institutional
rules change whether and how a private participation constraint enters the
coalition problem.

## Observable implications and neighboring mechanisms {-}

Four implications follow directly from the clean results. First, terminal
unanimity switches from low-only to pooling as the posterior probability of
the high outside-payoff type rises. Second, regular Round-1 unanimity is more
demanding: equilibrium may fail to exist, but every existing on-path outcome
pools at \(o_1\). Third, majority outcomes are group-size dependent and
set-valued; for \(N\geq5\), exclusion and separating current passage can both
survive. Fourth, entry nesting is selection-free only on the common
PBE-existence domain.

These implications differ from agenda power and endogenous institutional
choice. A positive recognition probability for \(H\) would give the informed
actor proposal power and introduce a signaling subgame. Allowing \(H\) to
choose the voting rule after observing its type could make the rule itself a
signal. Neither channel is part of the baseline.

## Scope {#scope}

The baseline fixes \(\pi_H=0\), \(b_\theta=0\), a unit weak pie, binary
private information, simultaneous ballots, and exogenous voting rules. A no
vote by \(H\) is an immediate opt-out; it is not a choice of
\(\beta C_\theta\), a hybrid \(\max\{o_\theta,\beta C_\theta\}\), or a
delayed-continuation threshold. The package space is always feasible and does
not revive the archived feasibility/C-B-R architecture.

The weak-vote-passive assessment is also a substantive scope condition. It is
informationally natural because weak states do not observe \(\theta\), but weak
PBE leaves globally off-path beliefs freer than sequential equilibrium.
Theorem \ref{thm:u-regular} states exactly where this freedom is used, and no
claim of uniqueness over all PBEs is made.

Boundary parameters matter. Low-only and rejection can reappear at
\(o_0=0\) or \(\beta=1\), and majority needs separate boundary bounds. These
cases are recorded in Appendix C rather than treated as if they belonged to
the regular interior theorem. Continuous types, heterogeneous weak states,
\(\pi_H>0\), endogenous rule choice, and sequential roll-call voting require
new derivations.

# Conclusion

Consensus is often described only as a constraint on power. The clean
immediate-opt-out model shows another possibility. When a powerful actor is
privately informed and pivotal, its approval becomes an informational
constraint on weaker proposers. This creates informational power even when
the hegemon has no agenda power.

The result is conditional. Regular unanimity equilibrium exists only on a
derived domain and pools whenever it exists. Majority equilibrium is generally
a correspondence rather than a unique no-screening benchmark. On their common
existence domain, weak-state formation under unanimity is nested in formation
under majority, while unanimity weakly raises the hegemon's payoff whenever
both institutions form. The institutional ranking therefore follows from
existence, formation, and equilibrium selection—not from a blanket claim that
one voting rule dominates.

The broader implication is that formal equality and asymmetric power can
coexist through information. Consensus can make a privately known outside
payoff politically valuable by turning a veto into a participation constraint.
That mechanism is distinct from both proposal power and constitutional choice,
and it survives precisely because the baseline keeps those channels separate.

# (APPENDIX) Appendix {-}

# Appendix A: Protocol and Histories {-}

## A.1 Public histories, beliefs, and voting incentives {-}

Let \(h\) denote the public history before a proposal and let
\(I=(h,s_i)\) denote a ballot information set. Once all simultaneous votes
are revealed, the public history is
\[
h_2=(h,s_i,a_H,v_W,\text{quota outcome},\text{opt-out status}).
\tag{A1}\label{eq:public-history}
\]
Continuation strategies can therefore condition on the proposal and complete
vote vector, not only on a posterior. Whenever Bayes' denominator is positive,
\[
\nu(h_2)=
\frac{\mu(h)\Pr(h_2\mid\theta=1,h)}
{\mu(h)\Pr(h_2\mid\theta=1,h)
+\{1-\mu(h)\}\Pr(h_2\mid\theta=0,h)}.
\tag{A2}\label{eq:bayes}
\]
The posterior \(\nu(h_2)\) and the continuation payoff
\(C_{H,2}^R(\theta,h_2)\) are distinct objects.

At an R2 information set \(I_2\), let \(p_R(I_2)\) be the probability that
\(H\)-yes implements the current proposal, given simultaneous weak strategies.
Then
\[
EU_H^R(Y\mid\theta,I_2)-EU_H^R(N\mid\theta,I_2)
=\beta p_R(I_2)(y-o_\theta).
\tag{A3}\label{eq:h-r2-difference}
\]
If \(p_R(I_2)>0\), the sign is that of \(y-o_\theta\). If
\(p_R(I_2)=0\), \(H\) is indifferent for every \(y\) and votes yes by the
tie rule. This terminal identity does not replace the Round-1 expected IC in
\eqref{eq:h-ic-r1}.

For a non-proposing weak voter \(j\), let
\(u_j(a_j,v_{-j};I)\) denote the payoff determined by the history table.
Approval requires
\[
\sum_{v_{-j}}\Pr(v_{-j}\mid I)u_j(Y,v_{-j};I)
\geq
\sum_{v_{-j}}\Pr(v_{-j}\mid I)u_j(N,v_{-j};I).
\tag{A4}\label{eq:weak-ic}
\]
The expectation includes \(H\)'s simultaneous type-contingent vote. Named
offers are paid whenever an agreement passes, even if the named recipient
voted no.

## A.2 Exhaustive history contract {-}

Table panels \ref{tab:gate0-timing}--\ref{tab:gate0-continuations} are
generated from the versioned source
\texttt{tables/clean\_optout\_gate0\_histories\_piH0.tsv}. Together they cover the
twenty-one history classes used by the proofs.

```{r gate0-history-timing, results='asis'}
history_timing <- dplyr::select(
  gate0_histories,
  history_id,
  round,
  rule,
  `active players`,
  quota,
  `quota reached?`,
  `H vote`
)
kableExtra::kbl(
  history_timing,
  format = "latex",
  booktabs = TRUE,
  longtable = TRUE,
  escape = TRUE,
  caption = "Gate 0 histories, panel A: timing, electorate, quota, and votes.",
  label = "gate0-timing"
) |>
  kableExtra::kable_styling(
    latex_options = "repeat_header",
    font_size = 7
  ) |>
  kableExtra::column_spec(1, width = "0.55in") |>
  kableExtra::column_spec(2, width = "0.45in") |>
  kableExtra::column_spec(3, width = "0.35in") |>
  kableExtra::column_spec(4, width = "1.55in") |>
  kableExtra::column_spec(5, width = "0.75in") |>
  kableExtra::column_spec(6, width = "1.55in") |>
  kableExtra::column_spec(7, width = "0.85in") |>
  kableExtra::landscape()
```

```{r gate0-history-implementation, results='asis'}
history_implementation <- dplyr::select(
  gate0_histories,
  history_id,
  `participation offer to H?`,
  `failure cause`,
  `H included?`,
  `H remains?`,
  `y implemented?`,
  `destination of y`
)
kableExtra::kbl(
  history_implementation,
  format = "latex",
  booktabs = TRUE,
  longtable = TRUE,
  escape = TRUE,
  caption = "Gate 0 histories, panel B: implementation and the disposition of the participation package.",
  label = "gate0-implementation"
) |>
  kableExtra::kable_styling(
    latex_options = "repeat_header",
    font_size = 7
  ) |>
  kableExtra::column_spec(1, width = "0.55in") |>
  kableExtra::column_spec(2, width = "1.25in") |>
  kableExtra::column_spec(3, width = "1.8in") |>
  kableExtra::column_spec(4, width = "0.8in") |>
  kableExtra::column_spec(5, width = "0.95in") |>
  kableExtra::column_spec(6, width = "1.15in") |>
  kableExtra::column_spec(7, width = "1.65in") |>
  kableExtra::landscape()
```

```{r gate0-history-payoffs, results='asis'}
history_payoffs <- dplyr::select(
  gate0_histories,
  history_id,
  `H payoff`,
  `weak payoffs`,
  `discount factor`
)
kableExtra::kbl(
  history_payoffs,
  format = "latex",
  booktabs = TRUE,
  longtable = TRUE,
  escape = TRUE,
  caption = "Gate 0 histories, panel C: payoffs and time units.",
  label = "gate0-payoffs"
) |>
  kableExtra::kable_styling(
    latex_options = "repeat_header",
    font_size = 8
  ) |>
  kableExtra::column_spec(1, width = "0.65in") |>
  kableExtra::column_spec(2, width = "1.75in") |>
  kableExtra::column_spec(3, width = "3.7in") |>
  kableExtra::column_spec(4, width = "2.1in") |>
  kableExtra::landscape()
```

```{r gate0-history-continuations, results='asis'}
history_continuations <- dplyr::select(
  gate0_histories,
  history_id,
  `relevant posterior`,
  `terminal or continuation`,
  `next subgame`
)
kableExtra::kbl(
  history_continuations,
  format = "latex",
  booktabs = TRUE,
  longtable = TRUE,
  escape = TRUE,
  caption = "Gate 0 histories, panel D: beliefs and continuation subgames.",
  label = "gate0-continuations"
) |>
  kableExtra::kable_styling(
    latex_options = "repeat_header",
    font_size = 8
  ) |>
  kableExtra::column_spec(1, width = "0.65in") |>
  kableExtra::column_spec(2, width = "4.1in") |>
  kableExtra::column_spec(3, width = "2.05in") |>
  kableExtra::column_spec(4, width = "1.8in") |>
  kableExtra::landscape()
```

The R2-unanimity row with \(H\) already absent documents a structurally
unreachable class. Once \(H\) opts out, the original unanimity quota cannot be
met and that branch has already ended.

# Appendix B: Proofs of Regular Results {-}

## B.1 Terminal rounds {-}

\textit{Proof of Lemma \ref{lem:r2-unanimity}.}
In terminal unanimity bargaining, every non-proposing weak voter approves: a
no vote either leaves failure unchanged or changes failure into an agreement
paying a nonnegative named offer, and ties go to yes. Hence \(H\) is pivotal,
and equation \eqref{eq:h-r2-difference} gives the cutoff
\(y\geq o_\theta\).

A package below \(o_0\) fails for both types. Within
\([o_0,o_1)\), \(y=o_0\) maximizes weak surplus and is accepted only by the
low type, giving \(L(\nu)\). At or above \(o_1\), \(y=o_1\) maximizes weak
surplus and pools, giving \(P\). Comparing these values yields
\eqref{eq:r2-cutoff}. At equality the weak proposer is indifferent and the
proposal tie-break selects the smaller transfer to \(H\). Uniform recognition
divides the total weak value by \(m\). \(\square\)

\textit{Proof of Proposition \ref{prop:r2-majority}.}
Terminal failure gives weak states zero, so weak approval is weakly dominant.
All weak states approve. Because \(q\leq m\), a weak proposer can set every
named offer to zero, pass with weak votes alone, and retain the unit pie. This
construction is valid whether \(H\) is active or has opted out. If \(H\) is
active and votes no, it receives \(o_\theta\) externally; if it votes yes,
the proposal can set \(y=0\). Uniform recognition gives the representative
weak payoff \(1/m\), and discounting gives \(c=\beta/m\) in Round-1 units.
\(\square\)

## B.2 Unanimity completion and existence {-}

\textit{Proof of Lemma \ref{lem:completion}.}
First suppose \(\beta o_1\geq o_0\). For each globally off-path proposal in
the upper-bound construction, retain a ballot belief assigning probability
\(1-\mu\in(0,1)\) to the low type. After a history with low-type
\(H\)-yes and exactly one weak no, declare a continuation posterior
\(\nu=1>\nu_2^*\), so terminal R2 selects pooling. Because the proposal itself
has probability zero, weak PBE does not link these two beliefs by Bayes' rule.
The low type then obtains \(\beta o_1\) after weak-caused failure and is
willing to vote yes; the high type opts out.

Complete globally off-path proposals as follows.

1. If \(y<o_0\), prescribe all weak voters yes and both types of \(H\) no.
   The proposer obtains zero.
2. If \(o_0\leq y<o_1\) and every named weak offer is at least
   \(d=\beta P/m\), prescribe all weak voters yes, low-type \(H\) yes, and
   high-type \(H\) no. The best such proposal gives at most \(G_L\).
3. If \(y\geq o_1\) and every named weak offer is at least \(d\), prescribe
   all weak voters and both types of \(H\) yes. The best such proposal gives
   at most \(G_P\).
4. If some \(x_j<d\), designate one underpaid weak rejector, prescribe
   low-type \(H\) yes and high-type \(H\) no, and use the pooling
   continuation. The rejector strictly prefers continuation, and the
   proposer's payoff is at most
   \((1-\mu)\beta P/m<G_L\).

Two or more weak no votes cannot occur because each rejector would be
nonpivotal and would have to vote yes by the tie rule. A nondegenerate ballot
mixture cannot enlarge the set because indifference also selects yes.

For the reverse direction, consider \(y=o_0\) and \(x_j=d\) for every
non-proposer. In any sequentially rational completion, all weak voters vote
yes. Two no votes violate the tie rule. With exactly one weak no, that voter
can strictly prefer rejection only if terminal continuation exceeds \(d\),
which requires R2 low-only. But low-only pays low-type \(H\)
\(\beta o_0<o_0\) after \(H\)-yes, so the low type opts out and makes the
weak rejector nonpivotal. With all weak voters yes, low-type \(H\) votes yes,
high-type \(H\) votes no, and the proposer receives exactly \(G_L\).

The same argument applies to \(y=o_1\) with \(x_j=d\). A lone weak rejector
again needs low-only continuation. Both types then prefer immediate opt-out
to the discounted terminal outside payoff after \(H\)-yes. The rejector
becomes nonpivotal, all weak voters vote yes, both types of \(H\) approve, and
the proposer obtains \(G_P\). Thus \(G_L\) and \(G_P\) are genuine
guarantees and the construction gives the matching upper bound.

If \(\beta o_1<o_0\), the feasible proposal \(y=o_1,x_j=0\) has no
sequentially rational ballot completion. All weak yes makes each weak voter
prefer a positive R2 continuation; two weak no votes violate the tie rule; and
with exactly one weak no both types of \(H\) opt out. Hence no PBE exists on
that open subdomain. \(\square\)

\textit{Proof of Theorem \ref{thm:u-regular}.}
Lemma \ref{lem:completion} shows that the proposer can force both \(G_L\) and
\(G_P\), and that every proposal can be completed at no more than their
maximum. Yet an on-path low-only proposal must pay each non-proposer
\(\beta D_0/m>d\), because \(H\)-yes reveals the low type. Therefore
\[
G_L-\pi_L=(1-\mu)\delta(D_0-P)>0.
\]
Low-only cannot be a regular PBE: off path the proposer shaves named weak
offers to \(d\), but once that proposal becomes on path Bayes' rule restores
the higher approval price.

Pooling attains its security value only when \(P>L\), in which case
\(\pi_P=aP=G_P\). Direct algebra gives
\[
\mu_E-\nu_2^*
=\frac{\delta P(D_0-P)}{D_0D_U}>0,
\tag{B1}\label{eq:cutoff-order}
\]
so \(G_P>G_L\) implies \(P>L\). At \(G_P=G_L\), the intermediate deviation
ties the proposer payoff and pays less to \(H\), so the proposal tie-break
selects it; once on path, its approval price jumps and the proposed fixed point
fails. Under the strict inequality, pooling strictly dominates every
completed deviation and Lemma \ref{lem:completion} constructs a global PBE.
The pooling package gives the payoff vector in
\eqref{eq:u-regular-payoffs}. \(\square\)

## B.3 Majority security and group-size characterization {-}

\textit{Proof of Proposition \ref{prop:majority}.}
Let \(X=\sum_{j\ne i}x_j\), and let \(r\) count non-proposers offered at least
\(c\). If at least \(q-1\) weak states vote yes, \(H\)-yes implements and type
\(\theta\) approves iff \(y\geq o_\theta\). If fewer than \(q-1\) weak
states vote yes, \(H\)-yes still fails and pays only
\(\beta o_\theta<o_\theta\), so both types opt out. Outside the all-weak-yes
profile, the only regular pure ballot size is therefore \(q-1\).

The upper-bound completion has three cases. If \(y<o_0\) and \(r\leq k-1\),
a \(q-1\)-weak-yes rejection completion pays the proposer \(c\); if
\(r\geq k\), weak-only passage gives at most \(1-X\leq E\). If
\(o_0\leq y<o_1\), declare the globally off-path ballot belief
\(\nu=1\). With \(r\leq k-1\), choose \(q-1\) weak yes votes; the true low
state passes, the true high state continues, and the proposer obtains at most
\(B_M\). With \(r\geq k\), weak-only passage is forced and yields less than
\(E\). If \(y\geq o_1\), all weak voters are nonpivotal when they all vote
yes, and the proposer receives at most \(P\).

Conversely, paying \(c\) to \(k\) weak partners guarantees \(E\);
\(y=o_0,x=0\) guarantees at least \(B_M\); and \(y=o_1,x=0\) guarantees
\(P\). Hence \(F_M\) is the exact security value.

For \(N=3\), all-weak-yes is a minimum winning coalition. Each of the three
security constructions is implementable when it maximizes, so proposer payoff
is \(F_M\). Exclusion and low-only remain set-valued when \(E=B_M\) because
both give \(H\) its outside payoff; pooling survives only under the stated
strict inequality. Their respective weak totals are
\[
T_W^E=1,\qquad
T_W^B=(1-\mu)(1-o_0)+\mu\beta,\qquad
T_W^P=P.
\tag{B2}\label{eq:m-n3-totals}
\]

For \(N=4\), the low-only proposal guarantees \(B_M\) off path but has no
Bayes-consistent pure on-path ballot. Only exclusion and pooling can attain the
security value. This gives condition \eqref{eq:m-n4-existence} and the two
selection regions. When \(B_M=P>E\), the unattainable deviation ties pooling
in proposer payoff and gives \(H\) less, so the proposal tie-break destroys
the pooling fixed point.

For \(N\geq5\), all-weak-yes is oversized because \(q\leq m-1\). For any
\(v\in[F_M,1]\), select an exclusion proposal with \(y=0\) and total named
weak offers \(X=1-v\). Every non-proposer is nonpivotal at all-weak-yes and
votes yes by the tie rule; the weak-only proposal passes and gives the proposer
\(v\). Complete every deviation with the security construction at no more
than \(F_M\). This proves sufficiency throughout the interval, including its
floor.

Any regular pure on-path proposal is then one of three classes. Exclusion has
\(y<o_0\), weak total one, and pays \(H\) its outside payoff. Separating
current passage has \(o_0\leq y<o_1\): low-type \(H\) votes yes, high-type
\(H\) votes no, and the oversized weak coalition passes in both states. With
named offers \(X\), proposer payoff and total weak payoff are
\[
v=1-X-(1-\mu)y,\qquad
T_W^M=1-(1-\mu)y.
\]
The canonical \(y=o_0,X=0\) class exists iff
\(1-(1-\mu)o_0\geq E\). Pooling uses \(y=o_1,X=0\) and exists iff
\(P>\max\{E,B_M\}\). These classes and their mixtures imply
\eqref{eq:m-weak-bounds}. \(\square\)

Equation \eqref{eq:no-cheap-gap} is affine in \(\mu\). Its endpoint gaps are
\(kc-o_0\) and \(qc-1\leq0\), giving condition
\eqref{eq:no-cheap}. The group-size implications follow directly from
Proposition \ref{prop:majority}.

## B.4 Institutional comparison {-}

\textit{Proof of Proposition \ref{prop:nesting}.}
On the common PBE domain, Theorem \ref{thm:u-regular} gives
\(V_W^U=P/m\). Proposition \ref{prop:majority} gives
\(V_W^M\geq F_M/m\), and \(F_M\geq P\) by definition. This proves
\eqref{eq:entry-nesting}, formation-set inclusion, and the three cost regions.
\(\square\)

\textit{Proof of Proposition \ref{prop:h-comparison}.}
Exclusion pays \(H\) exactly \(\bar o\). Separating current passage pays
\[
(1-\mu)y+\mu o_1\in[\bar o,o_1)
\]
because \(y\in[o_0,o_1)\). Pooling pays \(o_1\). Convex combinations of
payoff-tied majority classes preserve these bounds. Unanimity pays \(o_1\) on
the common domain by Theorem \ref{thm:u-regular}. This proves
\eqref{eq:h-payoff-bounds} and the conditional classification. \(\square\)

# Appendix C: Boundaries and Endpoint Limits {-}

## C.1 Zero low outside payoff {-}

Let \(\beta<1\), \(o_0=0\), \(0<o_1\leq1\), and
\(\mu\in(0,1)\). Retain
\[
P=1-o_1,\qquad L=1-\mu,\qquad
\delta=\frac{\beta(m-1)}m,\qquad a=1-\delta,
\]
and define
\[
K_0=\max\{aL,P-\delta\}.
\tag{C1}\label{eq:k0}
\]
The second term is inactive when negative.

\begin{proposition}[Unanimity when \(o_0=0\)]\label{prop:u-o0-boundary}
A global PBE always exists. If \(L\geq P\), low-only is the selected accepted
class and gives \(T_W^U=L\). If \(P>L\), canonical pooling exists and gives
\(T_W^U=P\); low-only also exists if and only if
\[
P-\delta\leq aL.
\tag{C2}\label{eq:o0-multiplicity}
\]
At equality the proposal tie-break favors low-only. If \(P>L\), an overpaid
pooling class with \(o_1<y\leq\bar y\) and \(T=1-y\) exists exactly when
\[
T\in[1-\bar y,P),
\qquad
T-\delta P>K_0.
\tag{C3}\label{eq:o0-overpay}
\]
Rejection is never on path on this boundary.
\end{proposition}

\textit{Proof.}
Terminal low-only gives each weak state \(1/m\), so after a globally off-path
proposal it can make the approval price \(\beta/m\). On-path low-only gives
the proposer \(aL\). A pooling proposal that pays \(\beta/m\) to every
non-proposer forces approval and gives \(P-\delta\). These two constructions
give the exact security value \(K_0\). Deliberate R1 rejection gives its
original proposer only \((1-\mu)\beta/m<aL\).

If \(L\geq P\), low-only attains the security value. If \(P>L\), canonical
pooling gives \(aP>K_0\), while low-only remains optimal under another
off-path completion exactly under \eqref{eq:o0-multiplicity}. At \(L=P\),
the tie-break selects low-only because it pays less to \(H\). An overpaid
pooling proposal gives proposer payoff \(T-\delta P\) and survives precisely
when it strictly exceeds both security components. \(\square\)

Low-only gives \(H\) expected payoff \(\mu o_1\), canonical pooling gives
\(o_1\), and overpaid pooling gives \(1-T\).

## C.2 No terminal weak surplus {-}

\begin{proposition}[Unanimity when \(o_1=1\)]\label{prop:u-o1-boundary}
If \(\beta<1\), \(o_0>0\), \(o_1=1\), and \(\mu\in(0,1)\), no unanimity
PBE exists.
\end{proposition}

\textit{Proof.}
Here \(P=0\). The proposal \(y=o_0,x_j=0\) forces weak approval. Two weak no
votes make each rejector nonpivotal. With exactly one weak no, a positive
continuation requires terminal low-only, but low-type \(H\) then receives only
\(\beta o_0<o_0\) after voting yes and opts out, again making the rejector
nonpivotal. With zero continuation, the weak yes tie rule applies. Low-type
\(H\) therefore approves, high-type \(H\) opts out, and the deviation
guarantees the proposer \(L=(1-\mu)(1-o_0)\). On-path low-only yields only
\(aL<L\), pooling leaves zero weak surplus, and rejection gives the original
proposer at most \(\beta L/m<L\). No on-path class attains the guarantee.
\(\square\)

## C.3 No discounting {-}

Let \(\beta=1\), \(0\leq o_0<o_1\leq1\), and
\(\mu\in(0,1)\). Define
\[
K_1=\max\left\{\frac Lm,
P-\frac{(m-1)(1-o_0)}m\right\}.
\tag{C4}\label{eq:k1}
\]

\begin{proposition}[Unanimity when \(\beta=1\)]\label{prop:u-beta1-boundary}
Rejection or continuation is admissible and gives the R1 proposer
\(S(\mu)/m\). If \(L\geq P\), low-only and rejection survive and pooling is
selected against. If \(P>L\), pooling and rejection survive; low-only also
survives if and only if
\[
P-\frac{(m-1)(1-o_0)}m\leq\frac Lm.
\tag{C5}\label{eq:beta1-low}
\]
For \(P>L\), an overpaid pooling class with \(o_1<y\leq\bar y\) and
\(T=1-y\) exists exactly when
\[
T-\frac{m-1}{m}P>K_1.
\tag{C6}\label{eq:beta1-overpay}
\]
\end{proposition}

\textit{Proof.}
With no discounting, after weak-caused failure low-type \(H\) is indifferent
between terminal low-only and immediate opt-out, and high-type \(H\) is
indifferent between terminal and immediate opt-out. The yes tie rule sustains
the punitive continuation. Low-only therefore gives and can guarantee
\(L/m\), while a pooling proposal that pays
\((1-o_0)/m\) to every non-proposer guarantees the second component of
\(K_1\) when feasible. A deliberately rejected proposal gives
\(S(\mu)/m\). The comparisons in \eqref{eq:beta1-low} and
\eqref{eq:beta1-overpay} then give the stated classes; equality is selected
against the proposal that pays \(H\) more. Mixtures over payoff- and
\(H\)-payoff-equivalent classes can survive. \(\square\)

This is the only baseline locus on which the historical rejection or delay
label survives on path. The intersections \(o_0=0\) and \(o_1=1\) are
included in Proposition \ref{prop:u-beta1-boundary}.

## C.4 Majority boundary bounds {-}

At \(\beta=1\), or at \(o_0=0\) with \(\beta<1\), \(H\)-yes can be optimal
even when only \(q-2\) weak states vote yes and the current ballot still fails.
Rejection classes can therefore reappear. For \(N\geq4\), the corresponding
algebraic floors are
\[
F_{M,1}=\max\left\{
1-kc,\,
(1-\mu)[1-o_0-(k-1)c]+\mu c,\,
1-o_1-(k-1)c
\right\},
\tag{C7}\label{eq:m-beta1-floor}
\]
and
\[
F_{M,0}=\max\left\{
1-kc,\,
(1-\mu)[1-(k-1)c]+\mu c,\,
1-o_1-(k-1)c
\right\}.
\tag{C8}\label{eq:m-o0-floor}
\]
These are boundary bounds, not complete equilibrium characterizations and not
substitutes for Proposition \ref{prop:majority}. No institutional comparison
in the paper extrapolates the regular theorem to these loci.

## C.5 Degenerate priors as one-sided limits {-}

Literal games with \(\mu=0\) or \(\mu=1\) require an extra convention about
zero-probability types. The paper instead reports one-sided limits of regular
equilibrium outcomes. Let
\[
z=(v_{\mathrm{prop}},T_W,V_W,E[u_H]),\qquad V_W=T_W/m,
\]
and let \(\mathcal E_R(\mu)\) be the regular on-path PBE-outcome
correspondence. For \(e\in\{0,1\}\), define
\[
\mathcal L_e^R=
\left\{
\lim_{n\to\infty}z_n:
\mu_n\in(0,1),\ \mu_n\to e,\
z_n\in\mathcal E_R(\mu_n)
\right\}.
\tag{C9}\label{eq:endpoint-limit}
\]
An empty correspondence means that no sequence of regular PBEs approaches the
endpoint, not that a separately specified degenerate game lacks a PBE.

For unanimity,
\[
\mathcal L_0^U=\varnothing,\qquad
\mathcal L_1^U=
\begin{cases}
\{(aP,P,P/m,o_1)\},&\beta o_1\geq o_0,\\
\varnothing,&\beta o_1<o_0.
\end{cases}
\tag{C10}\label{eq:u-endpoint-limits}
\]

For majority, write
\[
B_0^+=D_0,\qquad B_1^-=c,\qquad
F_0^+=\max\{E,D_0\},\qquad F_1^-=\max\{E,P\}.
\tag{C11}\label{eq:m-endpoint-floors}
\]

\begin{table}[H]
\centering
\small
\begin{tabular}{p{0.12\textwidth}p{0.39\textwidth}p{0.39\textwidth}}
\toprule
Group & \(\mu\to0^+\) & \(\mu\to1^-\) \\
\midrule
\(N=3\) &
Low-only \((D_0,D_0,D_0/m,o_0)\) if \(D_0>E\); otherwise exclusion
\((E,1,1/m,o_0)\). &
Pooling \((P,P,P/m,o_1)\) if \(P>E\); otherwise exclusion
\((E,1,1/m,o_1)\). \\
\(N=4\) &
Empty if \(D_0>E\); otherwise exclusion \((E,1,1/m,o_0)\). &
Pooling \((P,P,P/m,o_1)\) if \(P>E\); otherwise exclusion
\((E,1,1/m,o_1)\). \\
\bottomrule
\end{tabular}
\caption{One-sided majority equilibrium-outcome limits for \(N=3\) and \(N=4\).}
\label{tab:majority-endpoint-limits}
\end{table}

At \(D_0=E\), the lower lateral tie belongs to exclusion; at \(P=E\), the
upper lateral tie also belongs to exclusion. For \(N\geq5\),
\[
\mathcal L_0^M=
\begin{cases}
\{(v,1,1/m,o_0):v\in[E,1]\},&D_0<E,\\
\{(v,1,1/m,o_0):v\in[D_0,1]\}
\cup
\{(D_0,t,t/m,o_0):t\in[D_0,1]\},&D_0\geq E,
\end{cases}
\tag{C12}\label{eq:m-lower-limit}
\]
and
\[
\mathcal L_1^M=
\{(v,1,1/m,o_1):v\in[F_1^-,1]\}
\cup
\begin{cases}
\{(P,P,P/m,o_1)\},&P>E,\\
\varnothing,&P\leq E.
\end{cases}
\tag{C13}\label{eq:m-upper-limit}
\]

\textit{Derivation.}
The affine guarantee satisfies \(B_M(\mu)\to D_0\) at \(0^+\) and
\(B_M(\mu)\to c\) at \(1^-\). Substitution into the exact small-group
characterizations gives Table \ref{tab:majority-endpoint-limits}. For
\(N\geq5\), exclusion supplies the proposer intervals. Near zero, a
separating class has weak total \(1-(1-\mu)y\); attaining the floor forces
\(y\to o_0\) when \(D_0\geq E\). Near one, exclusion and separation converge
to weak total one and \(H\) payoff \(o_1\), while canonical pooling adds a
distinct limit exactly when \(P>E\). These exhaust the regular classes.
\(\square\)

There is no common lower-endpoint comparison because
\(\mathcal L_0^U\) is empty. At the upper endpoint, a common limit exists iff
\(\beta o_1\geq o_0\); entry nesting carries to every common selection, and
all corresponding \(H\)-payoff limits equal \(o_1\).

## C.6 One-shot bridge {-}

Suppressing Round 2 leaves exactly the terminal subgame in Lemma
\ref{lem:r2-unanimity}. The weak proposer compares
\((1-\mu)(1-o_0)\) with \(1-o_1\), so the low-only and pooling regions are
separated by \(\nu_2^*\). This is an interpretation of the proved terminal
subgame, not a new complete-information benchmark and not a second baseline
game.

# Appendix D: Notation {-}

\scriptsize
\begin{longtable}{p{0.14\textwidth}p{0.16\textwidth}p{0.47\textwidth}p{0.16\textwidth}}
\caption{Notation for the clean immediate-opt-out \(\pi_H=0\) baseline.}
\label{tab:notation}\\
\toprule
Symbol & Type & Meaning & First use \\
\midrule
\endfirsthead
\toprule
Symbol & Type & Meaning & First use \\
\midrule
\endhead
\(N\) & Primitive & Total number of states & Definition \ref{def:game} \\
\(m=N-1\) & Derived & Number of weak states & Definition \ref{def:game} \\
\(q\) & Rule parameter & Original strict-majority quota & Equation \eqref{eq:quota} \\
\(k=q-1\) & Derived & Additional yes votes required by a weak proposer under majority & Equation \eqref{eq:m-security} \\
\(\theta\) & Type & Persistent hegemon type, low \(0\) or high \(1\) & Definition \ref{def:game} \\
\(\mu\) & Belief & Prior probability of the high type & Definition \ref{def:game} \\
\(o_\theta\) & Primitive payoff & Immediate outside payoff after \(H\)-no; direct cutoff under certain implementation & Definition \ref{def:game} \\
\(y\) & Contract component & Participation package paid to \(H\) only when included & Definition \ref{def:game} \\
\(\bar y\) & Primitive bound & Maximum participation package & Equation \eqref{eq:primitive-domain} \\
\(x_j\) & Contract component & Named weak-state offer & Definition \ref{def:game} \\
\(\beta\) & Primitive & Discount factor for Round-2 payoffs in Round-1 units & Definition \ref{def:game} \\
\(\pi_H\) & Recognition parameter & Hegemon recognition probability, fixed at zero in the baseline & Introduction \\
\(\chi\) & Entry cost & External sunk cost per weak state & Equation \eqref{eq:entry-value} \\
\(P\) & Derived value & Terminal pooling weak total, \(1-o_1\) & Equation \eqref{eq:r2-u-value} \\
\(L(\nu)\) & Derived value & Terminal low-only weak total & Equation \eqref{eq:r2-u-value} \\
\(S(\nu)\) & Derived value & Terminal unanimity weak security value & Equation \eqref{eq:r2-u-value} \\
\(D_0\) & Derived value & Low-type weak residual, \(1-o_0\) & Equation \eqref{eq:u-notation} \\
\(\delta\) & Derived value & Aggregate discounted approval weight, \(\beta(m-1)/m\) & Equation \eqref{eq:u-notation} \\
\(a\) & Derived value & \(1-\delta\) & Equation \eqref{eq:u-notation} \\
\(G_L,G_P\) & Security values & Exact unanimity low-only and pooling guarantees & Equation \eqref{eq:u-security} \\
\(\mu_E\) & Existence cutoff & Regular unanimity existence threshold & Equation \eqref{eq:u-existence-cutoff} \\
\(c\) & Continuation value & Weak voter's discounted majority continuation, \(\beta/m\) & Equation \eqref{eq:r2-m-values} \\
\(E\) & Security value & Majority exclusion guarantee, \(1-kc\) & Equation \eqref{eq:m-security} \\
\(B_M\) & Security value & Majority low-only guarantee & Equation \eqref{eq:m-security} \\
\(F_M\) & Security value & Exact majority proposer security value & Equation \eqref{eq:m-security} \\
\(\bar o\) & Derived payoff & Prior expected outside payoff of \(H\) & Proposition \ref{prop:h-comparison} \\
\(\mathcal F_R\) & Formation set & Beliefs or selections for which rule \(R\) forms at cost \(\chi\) & Proposition \ref{prop:nesting} \\
\(K_0,K_1\) & Boundary values & Unanimity security values at \(o_0=0\) and \(\beta=1\) & Appendix C \\
\(\mathcal L_e^R\) & Limit correspondence & One-sided cluster set of regular equilibrium outcomes & Equation \eqref{eq:endpoint-limit} \\
\bottomrule
\end{longtable}
\normalsize

# References {-}
