---
title: "Informational Power and Institutional Design: When a Hegemon May Choose Consensus"
author: "Manoel Galdino (University of São Paulo; mgaldino@usp.br)"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    number_sections: TRUE
    toc: FALSE
    latex_engine: xelatex
header-includes:
  - \usepackage{setspace}
  - \setstretch{1.5}
  - \usepackage{parskip}
  - \setlength{\parskip}{6pt}
  - \setlength{\parindent}{0pt}
  - \usepackage{float}
  - \usepackage{amsmath,amssymb,amsthm}
  - \newtheorem{definition}{Definition}
  - \newtheorem{lemma}{Lemma}
  - \newtheorem{proposition}{Proposition}
  - \newtheorem{theorem}{Theorem}
  - \newtheorem{corollary}{Corollary}
  - \newtheorem{remark}{Remark}
  - \newtheorem{example}{Example}
  - \usepackage{tikz}
  - \usetikzlibrary{trees,arrows.meta,positioning,calc,shapes.geometric}
  - \usepackage{pdflscape}
  - \usepackage{booktabs}
fontsize: 12pt
geometry: margin=2.5cm
mainfont: Times New Roman
bibliography: references.bib
link-citations: true
abstract: |
  When and why would a hegemon prefer consensus to majority rule in an international organization? I argue that consensus can be a technology of hegemonic power rather than a constraint upon it. In a formal model, a hegemon privately observes the value of cooperation and uses Bayesian persuasion to influence weaker states' entry decisions. Under majority rule, weak states can bypass the hegemon, eliminating any screening problem. Under unanimity, weak states must include the hegemon in every bargaining coalition without knowing its type, creating a screening cutoff at which their behavior changes discretely. This generates an upward jump in the hegemon's value function---a non-convexity that Bayesian persuasion exploits. The institutional comparison has a single-crossing property: there is a unique prior threshold, in closed form, above which the hegemon prefers unanimity and below which majority can dominate by making entry easier. The mechanism reframes institutional design as a trade-off between informational leverage under unanimity and easier participation under majority, providing a distributive explanation for consensus rules in organizations like the GATT/WTO.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(knitr)
```

# Introduction

Most international organizations operate by consensus, even when powerful states drive their creation [@gould2022consensus]. Why would a powerful state ever prefer consensus over majority rule, where decisions can pass with only a simple majority? This is a puzzle, as the international arena is where material asymmetries—that is, differences in resources and influence among states—should matter most. If institutional rules reflect power, powerful states might be expected to favor arrangements that let them turn that power into a distributive advantage, meaning an ability to secure a greater share of benefits.

In bargaining, majority rule allows a powerful actor to form coalitions, exclude rivals, and keep the surplus [@baron1989bargaining; @kalandrakis2006proposal; @eraslan2019legislative]. Consensus does the opposite. By giving every participant a veto, it forces powerful states to surrender the very tools they might value most [@keohane1984after]. Yet, some major international organizations, such as the GATT/WTO, are built around consensus [@steinberg2002shadow; @koremenos2001rational].

Existing answers do not fully resolve the puzzle. A common response claims that powerful states still influence outcomes through informal agenda power, even under consensus. Yet this answer is conceptually unsatisfying: under unanimity, all participants are veto players, so informal agenda influence cannot enable exclusion and surplus capture as majority rule does. More broadly, existing accounts either see consensus as constraining powerful states or as a concealment for conventional power, but they do not explain when and why a hegemon would prefer unanimity to a rule that allows exclusion.

This paper theorizes when and why a hegemon may prefer unanimity to majority rule in international organizations. The central claim is that consensus is not merely a constraint or procedural facade; rather, it is an institutional technology of power. Using Bayesian persuasion [@kamenica2019bayesian], I show that consensus can make informational power more politically productive than the coalition arithmetic of majority rule. While the multi-receiver persuasion literature has examined information design under unanimity [@bardhiguo2018], and @kim2025persuasion analyzes Bayesian persuasion in bilateral veto bargaining, neither offers the institutional comparison between voting rules that is central here. Thus, the key institutional choice is not between power and constraint, but between two distinct modes of exercising power.

Two forms of power are central to the analysis. *Agenda power* is the ability to control proposals, shape coalitions, and determine which alternatives reach the floor---the familiar Baron-Ferejohn logic in which the capacity to exclude is the source of surplus capture. *Informational power* is the ability to shape beliefs about the value of cooperation; in international politics, large powers typically possess deeper technical expertise and greater ability to generate authoritative interpretations of proposed rules, whereas weaker states face limited analytical capacity. This paper's focus on information *design* under different voting rules complements the literature on information *aggregation*, where @feddersen1998convicting show that unanimity is the worst rule for aggregating dispersed private information. The present model reverses the direction: information is concentrated in a single actor (the hegemon), and unanimity amplifies rather than impedes its strategic value [on veto structures and bargaining power more generally, see @diermeier1999bicameralism]. Under majority rule, agenda power dominates: weak states can exclude the hegemon, so beliefs affect expected values but create no strategic discontinuity. Under unanimity, informational power dominates: every actor becomes pivotal, and the hegemon's private information creates a screening problem that generates a non-convexity in its value function.

To formalize the argument, I develop a model in which a hegemon chooses between two institutional packages that differ only in the voting rule: majority and unanimity. Proposal rights are symmetric under both rules, isolating the effect of the voting rule itself. Before weaker states decide whether to enter, the hegemon privately observes the value of cooperation and can shape their beliefs about it. The central difference is that the two rules reward this informational advantage differently. Under majority rule, weak states can form coalitions without making the hegemon pivotal, eliminating the screening problem. Under unanimity, weak states must include the hegemon in every proposal and decide how much to offer without knowing the probability of acceptance. This screening generates a cutoff at which weak proposers switch from aggressive to conservative treatment of the hegemon, creating a discontinuity in the hegemon’s value function that Bayesian persuasion exploits. This creates a persuasion opportunity absent under majority, although majority may still be preferred when it facilitates participation under less optimistic priors.

The analysis yields three main results. First, I characterize the screening cutoff under unanimity and show it has no analog under majority rule, where the hegemon’s value function is linear in beliefs (Propositions 1--3). Second, I show that this screening creates an additional non-convexity in the hegemon’s value function, providing a persuasion opportunity absent under majority (Proposition 4). Third, I establish that the institutional comparison has a single-crossing property: there is at most one prior at which the ranking between unanimity and majority changes, and I provide a closed-form expression for the threshold (Lemma 1 and Theorem 1).

The substantive implication is that consensus is most valuable to a hegemon when the prospects of multilateral cooperation are promising enough that weaker states are willing to come to the table. Under those conditions, unanimity amplifies the hegemon’s informational advantage: weaker states must include it in every deal without knowing what cooperation is truly worth, and the hegemon can shape their beliefs to exploit that uncertainty. When cooperation looks too uncertain for weaker states to participate at all, majority rule can be preferred—not because it gives the hegemon better bargaining terms, but because it lowers the bar for participation. These two forces produce a clean trade-off with a unique tipping point: above it, the hegemon prefers consensus; below it, it may prefer majority. The ranking never oscillates—the institutional comparison is not incomplete but conditional in a sharp and tractable way. Moreover, the boundary is robust: when the hegemon's bilateral alternatives exceed the threshold, unanimity loses its conditional advantage only near certainty about the magnitude of the gains from cooperation—precisely where private information is close to irrelevant. Under identifiable informational conditions, consensus becomes an institutional form through which hegemonic power is exercised, not a constraint upon it.

In sum, this paper contributes to the study of international organizations in three ways. First, it offers a new mechanism explaining when and why a powerful state may rationally choose unanimity rather than just accept it. Second, it reframes institutional design as a trade-off between informational leverage under unanimity and easier participation under majority, rather than solely between power and constraint. Third, it provides a formal link between voting rules and the strategic value of private information.

The rest of the paper proceeds as follows. Section 2 presents a motivating example that previews the mechanism in a simplified three-player environment. Section 3 presents the general model. Sections 4 and 5 derive the bargaining logic under majority and unanimity, showing that unanimity creates a screening cutoff and a discrete jump in the hegemon's payoff that majority eliminates. Section 6 connects bargaining to entry and persuasion. Section 7 establishes the main result. Section 8 discusses the GATT/WTO application and scope conditions. Section 9 concludes.


# A Motivating Example {#example}

Before turning to the general model, consider a stripped-down environment with three players to build intuition for the mechanism. There is one hegemon $H$ and two weak states $W_1, W_2$. The value of cooperation is either $V(0) = 1$ (low) or $V(1) = 2$ (high); $H$ observes the state, the weak states do not. Each weak state’s outside option is $0$; $H$’s outside option is $\alpha V(\theta)$ with $\alpha = 0.2$. A weak state is selected to make a take-it-or-leave-it offer specifying $H$’s share $y_H$; if $H$ rejects, each player receives their disagreement payoff.

**Majority: exclusion eliminates screening.** Under majority rule, $W_1$ and $W_2$ can pass a proposal without $H$’s vote. The proposing weak state excludes $H$ from the coalition; $H$ captures $0.2 V(\theta)$ from its bilateral alternative regardless of what $W$ believes. The hegemon’s expected payoff is $0.2 + 0.2\mu$, where $\mu = \Pr(\theta = 1)$---linear in beliefs, with no strategic discontinuity.

**Unanimity: inclusion creates screening.** Under unanimity, $H$’s vote is required. The proposing weak state must decide how much to offer $H$ without knowing $\theta$:

- *Aggressive offer* ($y_H = 0.2$): type $\theta = 0$ is indifferent and accepts; type $\theta = 1$ rejects (wants $0.4$). $W$ gets $0.8$ if $\theta = 0$, gets $0$ if $\theta = 1$.
- *Conservative offer* ($y_H = 0.4$): both types accept. $W$ gets $0.6$ if $\theta = 0$, gets $1.6$ if $\theta = 1$.

The weak state prefers the aggressive offer when $0.8(1-\mu) > 0.6(1-\mu) + 1.6\mu$, which simplifies to $\mu < 1/9$. The screening cutoff $\mu^* = 1/9$ is the belief at which $W$ switches from aggressive to conservative treatment of $H$.

**The jump and overpayment.** Below the cutoff, $H$’s expected payoff is the same as under majority: $0.2 + 0.2\mu$. Above the cutoff, $H$’s expected payoff jumps to $0.4$---constant, independent of $\mu$. The source is overpayment: on the conservative branch, type $\theta = 0$ receives $0.4$ but would accept $0.2$. Weak states cannot avoid this cost because they must buy $H$’s vote without knowing the type. The jump at $\mu^* = 1/9$ is $0.18$, or about 16\% of expected surplus at the cutoff.

**Bayesian persuasion exploits the jump.** The jump creates a non-convexity in $H$’s value function that Bayesian persuasion can exploit. Suppose the prior is $p < 1/9$. Without persuasion, $H$’s payoff under both rules is $0.2 + 0.2p$. But under unanimity, $H$ can design a public signal that pushes $W$’s posterior to $1/9$ with probability $9p$ (inducing conservative behavior and payoff $0.4$) and to $0$ with probability $1 - 9p$ (payoff $0.2$). The resulting expected payoff is $0.2 + 1.8p$, strictly exceeding the majority payoff of $0.2 + 0.2p$ for every $p > 0$. Under majority, the payoff is already linear, so concavification adds nothing. This is the central asymmetry: unanimity creates a non-convexity that makes information strategically productive; majority does not.

**Preview.** The general model (Section \@ref(model) onwards) enriches this logic with $N$ players, random proposer, two-round Baron-Ferejohn bargaining with discounting ($\beta < 1$), and an explicit entry stage with cost $c$. The richer environment produces two screening cutoffs (one per round), a richer non-convexity structure, and an entry threshold that creates a second channel through which the voting rule matters. Theorem \ref{thm:main} shows that the pattern here---unanimity dominates at sufficiently optimistic priors, majority can dominate only through easier entry at pessimistic priors---generalizes with a single-crossing property.



# The Model {#model}

## Environment

\begin{definition}[Institutional Design Game]\label{def:game}
The game $\Gamma(N, p, r, \alpha, \beta, c)$ has the following primitives.
\begin{itemize}
\item There are $N \geq 3$ players: one hegemon $H$ and $N-1$ weak states $W_1, \dots, W_{N-1}$.
\item Nature draws $\theta \in \{0, 1\}$ with prior $\Pr(\theta = 1) = p \in (0,1)$.
\item The state determines the value of multilateral cooperation: $V(0) = 1$ and $V(1) = r > 1$.
\item The hegemon observes $\theta$; the weak states do not.
\item Disagreement payoffs: each $W$ receives $0$;\footnote{The normalization $d_W = 0$ is without loss of generality: it requires only that weak states' outside options are symmetric and strictly lower than the hegemon's. The pie $V(\theta)$ is then interpreted as the surplus above weak states' common outside option.} $H$ receives $\alpha V(\theta)$, with $\alpha \in (0, 1/r)$.\footnote{The proportional form $d_H = \alpha V(\theta)$ captures bilateral alternatives that scale with the value of multilateral cooperation---the same factors that make multilateral cooperation valuable also improve the hegemon's outside option. The restriction $\alpha < 1/r$ ensures bilateral alternatives are strictly dominated by multilateral cooperation for all states of the world.}
\item The common discount factor is $\beta \in (0,1)$.
\item Each weak state that joins the institution pays an entry cost $c > 0$.
\item The voting rule is $R \in \{M, U\}$: under majority ($M$), decisions pass with $q = \lfloor N/2 \rfloor + 1$ votes; under unanimity ($U$), decisions require all $N$ votes.\footnote{The model analyzes formal unanimity (every member has a veto). In practice, ``consensus'' at the WTO refers to the norm of not calling a formal vote, which is distinct from a formal unanimity requirement. The results apply most directly to settings where each member has an effective veto, whether de jure or de facto.} Both rules have symmetric proposal rights (each player recognized with probability $1/N$).
\end{itemize}
\end{definition}

Let $V_e(\mu) \equiv 1 + \mu(r-1)$ denote the expected value of cooperation at posterior belief $\mu$, and define $x \equiv (N-1)\alpha r$ as a notational shorthand.

## Timing

The game has three stages.

1. **Stage 0 (Institutional choice).** The hegemon chooses the voting rule $R \in \{M, U\}$.

2. **Stage 1 (Information and entry).** The hegemon commits to a public signal structure $\pi: \{0,1\} \to \Delta(S)$. Nature draws $\theta$; the hegemon observes $\theta$ and the signal realization $s \in S$ is generated according to $\pi$. Weak states observe $s$, update to posterior $\mu = \Pr(\theta = 1 \mid s)$, and simultaneously decide whether to enter (paying cost $c$).

3. **Stage 2 (Bargaining).** If the institution forms, members play a two-round Baron-Ferejohn bargaining game over a pie of size $V(\theta)$:
    - **Round 1**: a proposer is selected uniformly at random (probability $1/N$). The proposer offers a division. All members vote according to rule $R$. If accepted, payoffs are realized. If rejected, the game proceeds to Round 2 with payoffs discounted by $\beta$.
    - **Round 2 (terminal)**: a new proposer is selected uniformly at random. The proposer offers a division. All members vote. If accepted, payoffs are realized. If rejected, each player receives their disagreement payoff.

## Solution concept

The solution concept is Perfect Bayesian Equilibrium (PBE). Strategies are sequentially rational given beliefs, and beliefs are updated by Bayes' rule on the equilibrium path. Off-path beliefs satisfy sequential rationality.\footnote{Tie-breaking convention: when $H$ is indifferent between accepting and rejecting an offer (offer equals reservation value), $H$ accepts.}

## Information structure

Throughout the game, the hegemon's information sets are singletons: $H$ knows $\theta$ at every decision node. Each weak state's information sets pool the two branches of $\theta$: at any decision node, $W_i$ knows the public signal realization and the history of proposals and responses, but not $\theta$. After the public signal, weak states hold posterior $\mu$ over $\theta$. This asymmetry is what generates the screening problem when a weak state proposes under unanimity.

Figures \ref{fig:gametree-a} and \ref{fig:gametree-b} present the extensive form of the game: Figure~\ref{fig:gametree-a} covers institutional choice, Bayesian persuasion, and entry (Stages 0--1); Figure~\ref{fig:gametree-b} details the bargaining subgame under unanimity (Stage 2).

\begin{landscape}

\begin{figure}[H]
\centering

\begin{tikzpicture}[
  grow=right,
  level distance=4cm,
  edge from parent/.style={draw,-{Latex[length=2.5mm]}},
  every node/.style={font=\small},
  player/.style={draw, circle, fill=white, inner sep=2pt, minimum size=7mm},
  nature/.style={draw, circle, fill=yellow!15, inner sep=2pt, minimum size=7mm},
  terminal/.style={draw, rectangle, fill=green!8, inner sep=3pt},
  subgame/.style={draw, rectangle, fill=orange!12, inner sep=4pt, rounded corners=2pt},
  infoset/.style={dashed, thick, gray},
  level 1/.style={sibling distance=5.5cm},
  level 2/.style={sibling distance=2.8cm},
  level 3/.style={sibling distance=2.2cm}
]

\node [player, label={[font=\large]left:{Stage 0}}] (root) {$H$}
%% ---- Majority branch ----
child {
  node [nature] (natM) {$N$}
  child {
    node [player] (wM0) {$W$}
    child { node [subgame] {Bargaining $(M)$} edge from parent node[above] {enter} }
    child { node [terminal] {$\alpha,\; 0$} edge from parent node[below] {not} }
    edge from parent node[above] {$\theta{=}0$}
  }
  child {
    node [player] (wM1) {$W$}
    child { node [subgame] {Bargaining $(M)$} edge from parent node[above] {enter} }
    child { node [terminal] {$\alpha r,\; 0$} edge from parent node[below] {not} }
    edge from parent node[below] {$\theta{=}1$}
  }
  edge from parent node[below] {\large $R = M$}
}
%% ---- Unanimity branch ----
child {
  node [nature] (natU) {$N$}
  child {
    node [player] (wU0) {$W$}
    child { node [subgame] {Bargaining $(U)$} edge from parent node[above] {enter} }
    child { node [terminal] {$\alpha,\; 0$} edge from parent node[below] {not} }
    edge from parent node[above] {$\theta{=}0$}
  }
  child {
    node [player] (wU1) {$W$}
    child { node [subgame] {Bargaining $(U)$} edge from parent node[above] {enter} }
    child { node [terminal] {$\alpha r,\; 0$} edge from parent node[below] {not} }
    edge from parent node[below] {$\theta{=}1$}
  }
  edge from parent node[above] {\large $R = U$}
};

% Information sets
\draw [infoset] (wM0) -- node[midway, left, font=\small, text=gray] {$W$} (wM1);
\draw [infoset] (wU0) -- node[midway, left, font=\small, text=gray] {$W$} (wU1);

% BP annotation between branches
\node[font=\normalsize\itshape, text=black, fill=white, inner sep=2pt]
  at ($(root) + (2cm, 0cm)$) {$H$ designs $\pi$};

\end{tikzpicture}

\caption{\footnotesize Extensive form: Stages 0--1 (institutional choice, information design, and entry). At Stage~0, $H$ selects voting rule $R \in \{M,U\}$ and commits to a Bayesian persuasion signal $\pi$. Nature draws $\theta \in \{0,1\}$; weak states observe signal realization $s$, form posterior $\mu = \Pr(\theta{=}1 \mid s)$, and decide entry (cost $c$). Dashed lines: $W$'s information sets ($W$ does not observe $\theta$). Non-entry payoffs: $(u_H, u_W) = (\alpha V(\theta),\, 0)$.}\label{fig:gametree-a}
\end{figure}

\end{landscape}

\begin{landscape}

\begin{figure}[H]
\centering

\begin{tikzpicture}[
  grow=right,
  level distance=3.5cm,
  edge from parent/.style={draw,-{Latex[length=2.5mm]}},
  every node/.style={font=\small},
  player/.style={draw, circle, fill=white, inner sep=2pt, minimum size=7mm},
  nature/.style={draw, circle, fill=yellow!15, inner sep=2pt, minimum size=7mm},
  terminal/.style={draw, rectangle, fill=green!8, inner sep=3pt},
  subgame/.style={draw, rectangle, fill=orange!12, inner sep=4pt, rounded corners=2pt},
  infoset/.style={dashed, thick, gray},
  level 1/.style={sibling distance=6cm},
  level 2/.style={sibling distance=3.2cm},
  level 3/.style={sibling distance=3.2cm},
  level 4/.style={sibling distance=2cm}
]

\node [nature, label={[font=\large]left:{R1}}] (r1) {$N$}
%% ---- W proposes ----
child {
  node [player] (wp) {$W$}
  %% Aggressive offer
  child {
    node [player] (hagg) {$H$}
    child {
      %% theta=1 rejects → R2 subgame (inline)
      node [nature, label={[font=\small]above:{R2}}] (r2) {$N$}
      child {
        node [terminal] {$\alpha r,\;\; r{-}\alpha r{-}c$}
        edge from parent node[above] {$W$ pr.}
      }
      child {
        node [terminal] {$r,\;\; {-}c$}
        edge from parent node[below] {$H$ pr.}
      }
      edge from parent node[above=5pt, pos=0.6] {rej ($\theta{=}1$)}
    }
    child {
      %% theta=0 accepts
      node [terminal] {$\alpha,\;\; 1{-}\alpha{-}c$}
      edge from parent node[below=5pt, pos=0.6] {acc ($\theta{=}0$)}
    }
    edge from parent node[above] {agg.\ $(y_H{=}\alpha)$}
  }
  %% Conservative offer
  child {
    node [terminal] (cons) {$\alpha r,\;\; V(\theta){-}\alpha r{-}c$}
    edge from parent node[below=2pt] {cons.\ $(y_H{=}\alpha r)$}
  }
  edge from parent node[above=3pt] {$W$ proposes $(\tfrac{N{-}1}{N})$}
}
%% ---- H proposes ----
child {
  node [player] (hp) {$H$}
  child {
    node [player] (wv) {$W$}
    child {
      node [subgame] {R2 subgame}
      edge from parent node[above] {rej}
    }
    child {
      node [terminal] {$V(\theta){-}(N{-}1)x_W,\; x_W{-}c$}
      edge from parent node[below] {acc}
    }
    edge from parent node[below] {$x_W{=}\beta V_W^{R2}$}
  }
  edge from parent node[below=3pt] {$H$ proposes $(\tfrac{1}{N})$}
};

% Screening annotation
\node[font=\normalsize\itshape, text=black, fill=white, inner sep=2pt]
  at ($(wp) + (0cm, -2.8cm)$)
  {$W$ screens: agg.\ iff $\mu < \mu_s$};

\end{tikzpicture}

\caption{\footnotesize Extensive form: Stage 2 (bargaining under unanimity, Rounds 1--2). Payoffs are $(u_H, u_W)$. A proposer is selected uniformly ($1/N$). When $W$ proposes, $W$ chooses between aggressive ($y_H = \alpha$) and conservative ($y_H = \alpha r$). Under the aggressive offer, type $\theta{=}0$ accepts (tie-breaking) and $\theta{=}1$ rejects, revealing the state and leading to R2 (payoffs discounted by $\beta$). In R2, $W$ knows $\theta{=}1$ (complete information); $W$ offers $\alpha r$, $H$ offers $0$; disagreement yields $(\alpha r, -c)$. When $H$ proposes in R1, $H$ offers $\beta V_W^{R2}$ to each $W$; $W$ accepts in equilibrium. Under majority rule, $H$'s vote is not pivotal, eliminating the screening mechanism (Proposition~\ref{prop:majority}).}\label{fig:gametree-b}
\end{figure}

\end{landscape}

## Preview of the main result

The paper's central result (Theorem \ref{thm:main}) establishes that the institutional comparison has the \emph{single-crossing property}: there exists at most one prior $p^*$ such that the hegemon prefers unanimity for $p > p^*$ and majority for $p < p^*$. The mechanism rests on three building blocks developed in Sections \@ref(majority)--\@ref(entry_bp): (1) majority produces no screening, so the hegemon's payoff is linear in beliefs (Section \@ref(majority)); (2) unanimity forces weak states to screen the hegemon's type, creating a discrete upward jump in the hegemon's expected payoff (Section \@ref(consensus)); and (3) this jump generates non-convexity that Bayesian persuasion exploits, more than compensating for the harder entry conditions under unanimity (Section \@ref(entry_bp)). The following sections develop each building block before assembling them into the institutional comparison in Section \@ref(comparison).

# Majority Rule: Linear Bargaining Without Screening {#majority}

Under majority rule, a proposal passes with $q = \lfloor N/2 \rfloor + 1$ votes. The key structural feature is that weak states can form a winning coalition without including the hegemon. When a weak state proposes, it assembles a majority from the other weak states alone; the hegemon, excluded from the coalition, captures its bilateral alternative $\alpha V(\theta)$ regardless of what weak states believe about the state of the world.

\begin{proposition}[Majority rule produces no screening]\label{prop:majority}
Under majority rule with symmetric proposal rights, the hegemon's expected continuation value $E_\theta[V_H^{R1}(\theta, \mu, M)]$ is affine in posterior beliefs $\mu$. There is no screening cutoff.
\end{proposition}

\begin{proof}
See Appendix B.1.
\end{proof}

Because the hegemon's vote is never needed, weak states never face a choice between offers that depend on inferring the hegemon's type. The hegemon's private information affects the value of the agreement but creates no strategic discontinuity. Majority transforms bargaining into coalition arithmetic: the proposer buys the cheapest votes, and the hegemon is simply paid its outside option. Bayesian persuasion under majority therefore operates only through the entry channel---it can induce weak states to participate, but it cannot extract screening rents from the bargaining process itself.


# Consensus: Screening Through Pivotal Inclusion {#consensus}

Under unanimity, every proposal requires the approval of all members, including the hegemon. When a weak state proposes, it must secure $H$'s vote without knowing whether the pie is high or low. As in the motivating example (Section \@ref(example)), this creates a screening problem: the proposer must choose between an aggressive offer that only the low type accepts and a conservative offer that both types accept. The conservative offer overpays the low type---the price of uncertainty.

This screening logic operates in both bargaining rounds. In each round, the weak proposer's indifference between aggressive and conservative offers defines a screening cutoff: a belief threshold above which the proposer switches from aggressive to conservative treatment. The cutoffs have closed-form expressions (Appendix A).

\begin{proposition}[R1 screening cutoff under consensus]\label{prop:cutoff_R1}
Under unanimity, there exists a unique screening cutoff $\mu_s^{R1} \in (0,1)$ such that the weak proposer prefers the aggressive offer for $\mu < \mu_s^{R1}$ and the conservative offer for $\mu > \mu_s^{R1}$. In the principal regime, the cutoff is:
\begin{equation}\label{eq:cutoff_R1}
\mu_s^{R1} = \frac{\phi - \sqrt{\phi^2 - 4(N-2)}}{2(N-2)}, \qquad \phi \equiv \frac{rN - \beta(N-1+r)}{\beta(r-1)}.
\end{equation}
In this regime, the cutoff does not depend on $\alpha$. For $\beta = 1$, it simplifies to $\mu_s^{R1} = 1/(N-2)$.
\end{proposition}

\begin{proof}
See Appendix B.2.
\end{proof}

The independence of the screening cutoff from the hegemon's outside option strength is substantively important. What determines how aggressively weak states treat the hegemon is not how valuable bilateral alternatives are, but how consequential the hegemon's private information is for the value of cooperation, how patient the bargainers are, and how many states must agree.

\begin{proposition}[Screening creates an informational rent]\label{prop:jump}
Under unanimity, the hegemon's expected continuation payoff in R1 has an upward jump at $\mu_s^{R1}$:
\begin{equation}\label{eq:jump_R1}
\text{Jump} = (1 - \mu_s^{R1}) \cdot \frac{(N-1)\beta(r-1)}{N^2} > 0.
\end{equation}
\end{proposition}

\begin{proof}
See Appendix B.3.
\end{proof}

At the screening cutoff, weak proposers switch from treating the hegemon as the low type to hedging against the possibility of the high type. The conservative offer pays the low type as if it could be the high type, creating a discrete upward shift in the hegemon's expected payoff. The screening rent is largest when the hegemon's private information is most consequential for the value of cooperation (large $r$), when bargaining is patient enough to make the continuation threat credible (large $\beta$), and in moderately-sized organizations---large enough that the hegemon's vote matters, but not so large that the cost of unanimous agreement dilutes the per-member advantage ($(N-1)/N^2$ peaks at intermediate $N$).

\begin{figure}[H]
\centering
\begin{tikzpicture}[x=8cm, y=10cm,
  >=Stealth,
  every node/.style={font=\small}]

% Axes
\draw[->, thick] (-0.05,0) -- (1.08,0) node[right] {$\mu$};
\draw[->, thick] (0,-0.02) -- (0,0.78) node[above] {$E[V_H^{R1}(\mu)]$};

% Tick marks on x-axis
\draw (0,-0.01) -- (0,0.01) node[below=3pt] {$0$};
\draw (1,-0.01) -- (1,0.01) node[below=3pt] {$1$};

% Screening cutoff vertical dashed line
\draw[dashed, gray] (0.197,0) -- (0.197,0.68);
\node[below=3pt] at (0.197,0) {$\mu_s^{R1}$};

% --- Majority (red, linear) ---
% From (0, 0.38) to (1, 0.52): affine, no jump
\draw[red, thick] (0,0.38) -- (1,0.52);
\node[red, right] at (1.01,0.52) {\footnotesize Majority (linear)};

% --- Unanimity (blue, piecewise with jump) ---
% Aggressive branch: from (0, 0.44) to (mu_s, 0.544)
% Conservative branch: from (mu_s, 0.602) to (1, 0.72)

% Aggressive branch
\draw[blue, thick] (0,0.44) -- (0.197,0.544);
% Open circle at right endpoint (left limit)
\draw[blue, thick, fill=white] (0.197,0.544) circle (1.5pt);

% Conservative branch
\draw[blue, thick] (0.197,0.602) -- (1,0.72);
% Closed circle at left endpoint (right limit)
\draw[blue, thick, fill=blue] (0.197,0.602) circle (1.5pt);

% --- Jump annotation ---
\draw[<->, thick, blue!70!black] (0.22,0.544) -- (0.22,0.602);
\node[right, blue!70!black, font=\footnotesize] at (0.23,0.573) {Screening rent};

% --- Branch labels ---
\node[blue, above, rotate=28, font=\footnotesize] at (0.08,0.48) {Aggressive};
\node[blue, above, rotate=8, font=\footnotesize] at (0.62,0.67) {Conservative};

% --- Concavification shading ---
% The concave envelope connects aggressive start to conservative end,
% passing above the jump. Show the region between envelope and value function.
% Envelope: line from (0, 0.44) to the conservative branch tangent point.
% For schematic clarity, draw envelope as dashed line from (0,0.44) to ~(0.42, 0.635)
% which is tangent to conservative branch, then follows conservative branch.
\draw[dashed, blue!50!black, thin] (0,0.44) -- (0.42,0.635);

% Shade the region between the dashed envelope and the actual value function
\fill[blue!10, opacity=0.5]
  (0,0.44) -- (0.197,0.544) -- (0.197,0.602) -- (0.42,0.635) -- cycle;
% Redraw the dashed line on top
\draw[dashed, blue!50!black, thin] (0,0.44) -- (0.42,0.635);

% BP gains annotation
\node[blue!50!black, font=\scriptsize, align=center] at (0.12,0.56) {BP\\gains};

\end{tikzpicture}
\caption{Schematic of the hegemon's expected R1 payoff as a function of posterior belief $\mu$, based on Example~\ref{ex:magnitudes} parameters ($N=5$, $r=1.5$, $\alpha=0.3$, $\beta=0.9$). Under majority (red), the payoff is linear in $\mu$ with no discontinuity. Under unanimity (blue), the payoff has two branches separated by an upward jump at the screening cutoff $\mu_s^{R1} \approx 0.197$: to the left, weak proposers play aggressively; to the right, they switch to conservative offers that overpay the low type. The shaded region shows where the concave envelope (dashed) lies above the value function, creating gains from Bayesian persuasion.}
\label{fig:screening-schematic}
\end{figure}

\begin{example}[Screening jump: concrete magnitudes]\label{ex:magnitudes}
Let $N=5$, $r=1.5$, $\alpha=0.3$, $\beta=0.9$. The R1 screening cutoff is $\mu_s^{R1} \approx 0.197$. At beliefs just below the cutoff, weak proposers play aggressively and the hegemon's expected payoff is $E[V_H^{R1}] \approx 0.544$. Just above the cutoff, the payoff jumps to $\approx 0.602$---an increase of about $5.3\%$ of expected surplus. Under majority at the same belief, $E[V_H^{R1}] \approx 0.428$, with no jump anywhere. The screening mechanism gives the hegemon $27\%$ more than majority on the aggressive branch and $41\%$ more on the conservative branch. The discrete gap between branches is the non-convexity that Bayesian persuasion exploits.
\end{example}


# Entry and Bayesian Persuasion {#entry_bp}

A weak state enters the institution if its expected bargaining payoff exceeds the entry cost $c > 0$.\footnote{The entry cost $c$ may represent fixed costs of negotiating accession, per-round costs of preparing delegations, or political costs of committing to multilateral disciplines. In established organizations, $c$ is best interpreted as the cost of \emph{substantive engagement}: allocating scarce diplomatic and analytical resources to a particular negotiating round. Since $V(\theta)$ is normalized independently of $N$, per-capita payoffs scale as $O(1/N)$. For large-$N$ applications, $c$ should be read as a per-capita cost: writing $c = \tilde{c}/N$, the entry condition becomes $N \cdot V_W \geq \tilde{c}$, where $N \cdot V_W$ is $O(1)$. All equilibrium cutoffs ($\mu_s^{R1}$, $\mu_s^{R2}$, $\alpha^*$, $p^*$) are homogeneous of degree zero in the surplus and thus independent of this normalization.} Because unanimity gives the hegemon a higher payoff conditional on participation (as established in Section \@ref(comparison)), it leaves less surplus for weak states, making entry harder: $\tau(U) \geq \tau(M)$.

The hegemon's reduced-form value function is $v(\mu, R) = E_\theta[V_H^{R1}(\theta, \mu, R)]$ when weak states enter and zero otherwise. Under majority, this function is affine above the entry threshold---a single non-convexity. Under unanimity, the screening jump at $\mu_s^{R1}$ creates an additional non-convexity.

\begin{proposition}[Unanimity creates an additional persuasion opportunity]\label{prop:nonconvexity}
The value function $v(\mu, U)$ has a screening non-convexity at $\mu_s^{R1}$ that is absent in $v(\mu, M)$.
\end{proposition}

\begin{proof}
See Appendix B.4.
\end{proof}

By the standard Bayesian persuasion result [@kamenica2011bayesian], the hegemon's optimal payoff under rule $R$ is $\Pi_H^*(R) = \operatorname{cav} v(p, R)$---the concavification of the value function evaluated at the prior. Under majority, concavification exploits only the entry threshold: the hegemon can induce participation but cannot extract screening rents. Under unanimity, concavification can exploit both entry and screening, inducing posteriors that straddle $\mu_s^{R1}$ and place mass on both sides of the jump. This dual exploitation is the paper's distinctive contribution. Figure \@ref(fig:bp-illustration) illustrates both value functions and their concavifications.

```{r bp_illustration, echo=FALSE, fig.width=10, fig.height=4.5, fig.cap="Value functions and concavification under majority (red) and unanimity (blue). Dashed lines show concave envelopes. Left panel (a): baseline parameters ($N=5$, $r=1.5$, $\\alpha=0.3$, $\\beta=0.9$, $c=0.1$) where the screening jump under unanimity creates a large concavification gain. Right panel (b): parameters near the boundary of the unanimity-preferred region ($N=5$, $r=1.2$, $\\alpha=0.3$, $\\beta=0.7$, $c=0.1$) where the mechanism is weaker and the concavification gains are smaller."}

source("scripts/model_functions.R", local = TRUE)

# Helper to plot one panel
plot_panel <- function(r_val, alpha_val, N_val, beta_val, c_val, title) {
  mus_fine <- seq(0.001, 0.999, by = 0.001)

  v_U <- sapply(mus_fine, function(m) {
    vw <- VW_R1_unanimity(r_val, alpha_val, m, N_val, beta_val)
    if (vw >= c_val) VH_R1_unanimity(r_val, alpha_val, m, N_val, beta_val) else 0
  })

  v_M <- sapply(mus_fine, function(m) {
    vw <- VW_R1_majority(r_val, alpha_val, m, N_val, beta_val)
    if (vw >= c_val) VH_R1_majority(r_val, alpha_val, m, N_val, beta_val) else 0
  })

  cav_U <- concavify(mus_fine, v_U)
  cav_M <- concavify(mus_fine, v_M)

  ylim <- range(c(v_U, v_M, cav_U, cav_M))

  plot(mus_fine, v_U, type = "l", col = "blue", lwd = 2,
       xlab = expression(paste("Posterior ", mu)),
       ylab = expression(paste("v(", mu, ", R)")),
       ylim = ylim, las = 1, main = title)
  lines(mus_fine, v_M, col = "red", lwd = 2)
  lines(mus_fine, cav_U, col = "blue", lwd = 1.5, lty = 2)
  lines(mus_fine, cav_M, col = "red", lwd = 1.5, lty = 2)

  # Mark screening cutoff
  phi_val <- (r_val * N_val - beta_val * (N_val - 1 + r_val)) / (beta_val * (r_val - 1))
  disc <- phi_val^2 - 4*(N_val-2)
  if (disc > 0) {
    mu_s_R1 <- (phi_val - sqrt(disc)) / (2*(N_val-2))
    abline(v = mu_s_R1, col = "gray50", lty = 3)
    text(mu_s_R1 + 0.02, ylim[2]*0.9, expression(mu[s]^{R1}), col = "gray40", cex = 0.8)
  }

  legend("topleft", c(expression(v(mu, U)), expression(v(mu, M)),
                       expression(cav~v(mu, U)), expression(cav~v(mu, M))),
         col = c("blue", "red", "blue", "red"),
         lwd = c(2, 2, 1.5, 1.5), lty = c(1, 1, 2, 2), cex = 0.8, bg = "white")
}

par(mfrow = c(1, 2), mar = c(4, 4, 2.5, 1))

# Panel (a): baseline — strong mechanism
plot_panel(r_val = 1.5, alpha_val = 0.3, N_val = 5, beta_val = 0.9, c_val = 0.1,
           title = "(a) Strong mechanism")

# Panel (b): near boundary — weak mechanism
plot_panel(r_val = 1.2, alpha_val = 0.3, N_val = 5, beta_val = 0.7, c_val = 0.1,
           title = "(b) Weak mechanism")
```


# Institutional Choice {#comparison}

This section establishes the paper's main result: conditions under which the hegemon prefers unanimity.

## Conditional payoff comparison

Before turning to the full institutional comparison with persuasion, I establish a result that is central to the analysis: conditional on the institution forming, unanimity gives the hegemon a strictly higher expected continuation payoff than majority for all interior beliefs.

\begin{lemma}[Conditional payoff dominance of unanimity]\label{lem:conditional}
Define the threshold
\[
\alpha^*(N, \beta) \equiv \frac{\beta(q-1)}{N(N-1) - \beta(N^2 - N - q + 1)}, \qquad q = \lfloor N/2 \rfloor + 1.
\]
Then $\alpha < \alpha^*(N, \beta)$ if and only if, for every $\mu \in (0,1]$,
\[
E_\theta[V_H^{R1}(\theta, \mu, U)] > E_\theta[V_H^{R1}(\theta, \mu, M)].
\]
\end{lemma}

\begin{proof}
See Appendix B.5.
\end{proof}

The boundary $\alpha^*$ is tight: at $\alpha = \alpha^*$, the advantage vanishes exactly at $\mu = 1$, where the hegemon's private information is irrelevant. The intuition has two parts. When the hegemon proposes, unanimity depresses weak states' continuation values (because rejection leads to a game where the hegemon is pivotal), allowing the hegemon to buy agreement more cheaply. When a weak state proposes, unanimity forces the proposer to secure the hegemon's vote, generating the screening rent identified in Proposition \ref{prop:jump}. Both forces favor the hegemon.

\begin{remark}[Robustness beyond $\alpha^*$]\label{rem:mu_bar}
When $\alpha \geq \alpha^*$, the conditional advantage of unanimity fails only near $\mu = 1$. In Region III, the payoff difference decomposes as
\[
D(\mu) = \underbrace{D(1)}_{\text{without info}} + (1 - \mu) \cdot \underbrace{\Gamma}_{\text{value of uncertainty}},
\]
where $D(1) = r[C_{\text{buy}} - C_{\text{out}}(1-\beta)]/N^2 \leq 0$ is the payoff comparison when the hegemon's type is known, and $\Gamma = (r-1)[C_{\text{out}} - C_{\text{buy}} + (N-1)\beta]/N^2 > 0$ is the marginal value of uncertainty for unanimity. The belief threshold above which majority dominates is $\bar\mu = 1 - |D(1)|/\Gamma$. Numerically, $\bar\mu$ is remarkably stable: for $N = 164$ and $\beta = 0.9$, increasing $\alpha$ from $0.05$ to $0.49$ (an 18-fold increase beyond $\alpha^*$) only lowers $\bar\mu$ from $0.87$ to $0.71$ when $r = 1.5$. The mechanism fails only near certainty about the magnitude of the gains from cooperation---precisely where private information has least value (Figure \@ref(fig:heatmap-alpha-mu) in Section \@ref(scope) maps the full $(\alpha, \mu)$ frontier).
\end{remark}

An immediate corollary is that entry is always harder under unanimity ($\tau(U) \geq \tau(M)$): because the hegemon captures more surplus, less is available for weak states. Without persuasion, the institutional comparison is therefore ambiguous: unanimity dominates when the institution forms under both rules, but majority dominates when only it can induce entry.

## The main result

With Bayesian persuasion, the hegemon's payoff under each rule is $\Pi_H^*(R) = \operatorname{cav} v(p, R)$. The institutional choice reduces to comparing these concavified values.

\begin{theorem}[Threshold prior for institutional choice]\label{thm:main}
Suppose $\alpha < \alpha^*(N, \beta)$. The institutional comparison has the \emph{single-crossing property}: there exists at most one prior $p^*$ at which the ranking between unanimity and majority changes. Specifically:
\begin{itemize}
\item[\textbf{(a)}] When entry costs are low enough that the institution forms under unanimity at all priors, unanimity strictly dominates for every $p \in (0,1)$.
\item[\textbf{(b)}] When entry occurs under majority but not unanimity at low priors, there exists a unique threshold $p^*$ such that majority dominates for $p < p^*$ and unanimity dominates for $p > p^*$.
\end{itemize}
The threshold $p^*$ has a closed-form expression (Appendix B.6). In all cases, unanimity dominates at sufficiently optimistic priors, and any advantage of majority arises exclusively through the entry margin.
\end{theorem}

\begin{proof}
See Appendix B.6.
\end{proof}

The theorem resolves the institutional choice puzzle. Majority's only advantage is that it makes participation cheaper for weak states, because the hegemon captures less surplus. But this advantage is relevant only when the prior is pessimistic enough that entry is the binding constraint. At optimistic priors---where the value of cooperation is likely high and the institution would form under either rule---unanimity strictly dominates because the screening mechanism gives the hegemon an informational rent that majority cannot replicate.

The single-crossing property means there is no oscillation: the ranking switches at most once. Consensus is not sometimes better and sometimes worse depending on delicate parameter configurations; it is better whenever the institution is viable.

## Numerical characterization

```{r parameter_regions, echo=FALSE, fig.width=8, fig.height=4, fig.cap="Parameter regions where the hegemon prefers unanimity (blue) vs majority (red) at prior $p = 0.05$, restricted to $\\alpha < \\alpha^*(N, \\beta)$ (the domain where Lemma 1 applies). Left: $r$ vs $\\alpha$ plane with $\\beta=0.9$, $N=5$, $c=0.1$. Right: $\\beta$ vs $r$ plane with $\\alpha=0.3$, $N=5$, $c=0.1$."}

compute_preference <- function(r, alpha, N, beta, c, p) {
  # Value under unanimity
  vw_u <- tryCatch(VW_R1_unanimity(r, alpha, p, N, beta), error = function(e) NA)
  vh_u <- tryCatch(VH_R1_unanimity(r, alpha, p, N, beta), error = function(e) NA)

  # Value under majority
  vw_m <- tryCatch(VW_R1_majority(r, alpha, p, N, beta), error = function(e) NA)
  vh_m <- tryCatch(VH_R1_majority(r, alpha, p, N, beta), error = function(e) NA)

  if (is.na(vw_u) | is.na(vh_u) | is.na(vw_m) | is.na(vh_m)) return(NA)

  # Compute slopes (concavification at low p)
  # Find S_U: max v(mu, U)/mu for mu where entry occurs
  mus_grid <- seq(0.01, 0.99, by = 0.005)

  slopes_U <- sapply(mus_grid, function(m) {
    vw <- VW_R1_unanimity(r, alpha, m, N, beta)
    if (vw >= c) {
      vh <- VH_R1_unanimity(r, alpha, m, N, beta)
      return(vh / m)
    } else {
      return(0)
    }
  })

  slopes_M <- sapply(mus_grid, function(m) {
    vw <- VW_R1_majority(r, alpha, m, N, beta)
    if (vw >= c) {
      vh <- VH_R1_majority(r, alpha, m, N, beta)
      return(vh / m)
    } else {
      return(0)
    }
  })

  S_U <- max(slopes_U, na.rm = TRUE)
  S_M <- max(slopes_M, na.rm = TRUE)

  if (S_U > S_M) return(1) else return(0)
}

# Panel 1: r vs alpha
# alpha* threshold for Lemma 1: alpha* = beta*(q-1) / [N*(N-1) - beta*(N^2-N-q+1)]
alpha_star <- function(N, beta) {
  q <- floor(N/2) + 1
  beta * (q - 1) / (N * (N - 1) - beta * (N^2 - N - q + 1))
}
a_star_val <- alpha_star(5, 0.9)

rs <- seq(1.1, 3, by = 0.1)
alphas <- seq(0.05, min(0.45, a_star_val - 0.025), by = 0.025)

par(mfrow = c(1, 2), mar = c(4, 4, 2, 1))

mat1 <- matrix(NA, length(rs), length(alphas))
for (i in seq_along(rs)) {
  for (j in seq_along(alphas)) {
    if (alphas[j] < 1/rs[i] & alphas[j] < a_star_val) {
      mat1[i, j] <- tryCatch(
        compute_preference(rs[i], alphas[j], 5, 0.9, 0.1, 0.05),
        error = function(e) NA
      )
    }
  }
}

image(rs, alphas, mat1, col = c("firebrick3", "steelblue3"),
      xlab = "r (type gap)", ylab = expression(alpha), main = expression(paste(beta, "=0.9, N=5")),
      las = 1)
# Add alpha < 1/r boundary
curve(1/x, from = 1.1, to = 3, add = TRUE, lty = 2, lwd = 1.5)
# Add alpha* boundary
abline(h = a_star_val, lty = 3, lwd = 1.5, col = "gray40")
legend("topright", c("H prefers U", "H prefers M", expression(alpha == 1/r), expression(alpha^"*")),
       fill = c("steelblue3", "firebrick3", NA, NA), border = c("black", "black", NA, NA),
       lty = c(NA, NA, 2, 3), lwd = c(NA, NA, 1.5, 1.5), col = c(NA, NA, "black", "gray40"),
       cex = 0.7, bg = "white")

# Panel 2: beta vs r
betas <- seq(0.5, 0.99, by = 0.025)

mat2 <- matrix(NA, length(betas), length(rs))
for (i in seq_along(betas)) {
  for (j in seq_along(rs)) {
    mat2[i, j] <- tryCatch(
      compute_preference(rs[j], 0.3, 5, betas[i], 0.1, 0.05),
      error = function(e) NA
    )
  }
}

image(betas, rs, mat2, col = c("firebrick3", "steelblue3"),
      xlab = expression(beta), ylab = "r", main = expression(paste(alpha, "=0.3, N=5")),
      las = 1)
legend("topleft", c("H prefers U", "H prefers M"),
       fill = c("steelblue3", "firebrick3"), cex = 0.7, bg = "white")
```

Figure \@ref(fig:parameter-regions) shows that the hegemon prefers unanimity across a broad range of parameters when the prior is low ($p = 0.05$). In the blue regions, unanimity dominates globally---case (a) of Theorem \ref{thm:main}. The unanimity-dominance region is larger when the hegemon's private information is more consequential (higher $r$) and when bargaining is more patient (higher $\beta$). In the red regions, case (b) applies: the threshold $p^*$ lies above $0.05$, so majority dominates at this pessimistic prior, but unanimity would dominate at higher priors.


# Discussion {#discussion}

## Illustration: The GATT/WTO {#gatt_wto}

The GATT/WTO illustrates the interaction between consensus and informal agenda influence. Since 1947, the multilateral trading system has operated predominantly by consensus, even though formal voting provisions exist [@steinberg2002shadow]. The United States and other major trading powers were central to the system's creation and could plausibly have imposed majority rule.

The model's informational asymmetry assumption finds support in the trade context. Evaluating the effects of tariff schedules, anticipating distributive consequences, and assessing long-term implications of rule changes require sustained investment in analytical infrastructure. Major trading powers maintain large permanent delegations with deep sectoral expertise; many developing country delegations operate with far smaller teams, particularly during complex technical phases of negotiations. This capacity asymmetry means that weaker states often face genuine uncertainty about the value of proposed agreements.

In practice, major trading powers exercise informal agenda influence through the Green Room process, control over drafting, and technical expertise. The model's baseline uses symmetric proposal rights to isolate the voting-rule comparison, but the qualitative mechanism---screening under unanimity, no screening under majority---does not depend on equal recognition probabilities. Monopoly agenda control would eliminate participation (as argued in Section \@ref(scope)), so the relevant question is whether moderate informal influence coexists with the screening channel. The persistence of consensus is consistent with the coexistence of informal influence and formal equality.

The model's mechanism explains a pattern that existing accounts describe but do not fully explain: powerful states in the WTO appear to benefit from consensus despite lacking formal agenda control. The mechanism is not merely that informal practices survive consensus, but that consensus creates the strategic environment---pivotal inclusion under uncertainty---in which informational power becomes most valuable. The model does not claim that the GATT/WTO was designed for this reason, or that all aspects of WTO politics reduce to this mechanism. It shows that consensus and informational asymmetry interact in a way that can benefit powerful states, offering one explanation for why hegemonic preference for consensus is not paradoxical.

The informational mechanism generates a distinctive observable implication: the hegemon's advantage from consensus should be largest in negotiations where informational asymmetry is substantial and where weak states lack independent analytical capacity. In the WTO context, this suggests that consensus should matter most in complex regulatory areas---services, intellectual property, investment---where evaluating proposals requires deep technical expertise, and least in areas where distributive consequences are transparent, such as linear tariff cuts. This prediction distinguishes the informational mechanism from legitimacy accounts (which predict consensus matters uniformly) and from self-enforcement accounts (which predict consensus matters most when enforcement is weak, regardless of information).^[Appendix C shows that with $K > 2$ types, unanimity generates $K-1$ screening boundaries while majority remains linear. More issue dimensions increase the number of screening regions but also make the screening problem harder for the hegemon, suggesting that the mechanism is strongest in focused negotiations.]

**The GATT-to-WTO transition.** The model's institutional comparison offers a lens for understanding the evolution of decision-making rules in the multilateral trading system. The GATT 1947 provided for formal majority voting---simple majority for joint action, two-thirds for amendments---and voting was used regularly in the organization's early years. Beginning in the late 1950s, however, consensus gradually displaced voting as the operative decision-making norm; since 1959, virtually all GATT/WTO legislative decisions have been taken by consensus [@steinberg2002shadow, 344]. The transition coincided with the *en masse* accession of developing countries, which fundamentally changed both the size and the informational composition of the membership.

The model suggests a structural interpretation. In the early GATT, membership was small and relatively homogeneous---predominantly industrialized economies with similar analytical capacities. In such an environment, informational asymmetry between the hegemon and other members was modest, and the primary challenge was inducing participation in a new and untested institution. These are precisely the conditions under which Theorem \ref{thm:main} predicts majority dominance: when forming the institution is the binding constraint (low prior, high effective entry cost), the easier participation conditions under majority outweigh the screening advantages of unanimity. As the membership expanded and diversified, two shifts occurred simultaneously: the organization became more viable (lowering the effective entry barrier), and the informational gap between the hegemon and newer members widened (increasing $r$). Both shifts move the institutional comparison toward unanimity. The consensus norm that emerged was not merely a concession to the principle of sovereign equality---it was a rational adaptation to an environment in which informational power had become the hegemon's most valuable asset.

More broadly, the model predicts that the transition from majority to consensus should be most pronounced in *new regimes*---institutional settings where the benefits of cooperation are genuinely uncertain and where participants lack prior experience to assess the value of proposed agreements. When a new international organization operates within an already-established regime (with known templates, precedents, and track records), the effective uncertainty about the value of cooperation is low regardless of the formal novelty of the institution. It is the novelty of the *regime*, not the organization, that generates the high entry costs and informational asymmetries under which majority initially dominates. The early GATT exemplifies this: multilateral trade liberalization was an untested project with no historical precedent, and the informational demands of evaluating tariff schedules, anticipating distributive consequences, and assessing long-term welfare effects were substantial. These conditions made majority rule the natural starting point and consensus the natural destination.

## Scope {#scope}

**Why symmetric proposal rights?** Under either voting rule, if the hegemon were the exclusive proposer and weak states had no outside option from bargaining ($d_W = 0$), weak states' expected bargaining payoff would be zero. No weak state would pay an entry cost $c > 0$, and the institution would never form. Monopoly proposal power is self-defeating when participation is voluntary. The relevant institutional choice is therefore between voting rules, holding proposal rights symmetric.

**Why not compare consensus with hegemonic agenda control?** Because agenda control combined with any participation constraint eliminates the institution itself. The puzzle is not why a hegemon does not impose dictatorship, but why it would choose a rule that makes its own approval necessary when majority would allow others to bypass it.

**When consensus favors the hegemon.** The mechanism requires that: (i) informational asymmetry is relevant ($H$ knows $\theta$, $W$ does not); (ii) the hegemon has a valuable outside option ($\alpha > 0$); (iii) entry is costly and belief-dependent; (iv) the voting rule makes $H$ pivotal.

The model assumes the hegemon can commit to an information structure before observing $\theta$. This is a strong assumption that warrants discussion. Three considerations support its plausibility in the IO context. First, \emph{reputation}: in repeated interactions, a hegemon that deviates from an announced disclosure policy suffers credibility costs in future negotiations, which can sustain commitment in equilibrium. Second, \emph{institutional transparency rules}: organizations like the WTO have formal provisions for notifications, trade policy reviews, and dispute settlement that constrain and channel information disclosure, serving as partial commitment devices. Third, \emph{upper bound interpretation}: the Bayesian persuasion payoff represents the maximum value of the hegemon's informational advantage. Even if full commitment is unavailable, the qualitative institutional comparison---unanimity creates screening opportunities that majority does not---may survive under weaker information structures, such as evidence games [@glazer2004study] or verifiable disclosure [@milgrom1981good].

**When majority dominates.** The biconditional in Lemma \ref{lem:conditional} implies that $\alpha < \alpha^*$ is the largest parametric region where majority can never dominate through a higher conditional-on-entry payoff. Any advantage of majority must come through the entry margin: $\tau(U) > \tau(M)$. This is more likely when entry costs are high (creating a large gap where the institution forms only under majority), when $\alpha$ approaches $\alpha^*$ (weakening the conditional dominance), or when information is approximately public. For $\alpha \geq \alpha^*$, majority can also dominate through conditional bargaining payoffs near $\mu = 1$, because unanimity's cost of buying $N-1$ votes outweighs the screening advantage.

Figure \@ref(fig:alpha-star) plots $\alpha^*(N, \beta)$ as a function of organization size for three levels of patience. The threshold decreases with $N$ because unanimity requires buying $N-1$ votes, while majority requires only $q-1$. For large organizations ($N > 50$), $\alpha^*$ becomes small: the conditional dominance result applies only when the hegemon's bilateral alternative is modest relative to multilateral cooperation. This does not eliminate the mechanism: even when Lemma \ref{lem:conditional} fails globally, the screening jump at $\mu_s^{R1}$ persists, and Bayesian persuasion can still exploit it at low priors. The condition restricts when unanimity dominates \emph{everywhere}, not when it dominates \emph{somewhere}.

The decreasing relationship between $\alpha^*$ and $N$ has a substantive implication. A hegemon whose bilateral alternatives are moderate relative to multilateral cooperation---$\alpha$ well below $\alpha^*$---benefits from consensus. But as bilateral alternatives strengthen, the screening mechanism no longer compensates for the cost of buying unanimous agreement. The proliferation of preferential trade agreements can be understood, in the model's terms, as an increase in $\alpha$: bilateral and regional deals strengthen the hegemon's outside option relative to multilateral cooperation [@bhagwati2008termites]. As $\alpha$ rises past $\alpha^*$, the hegemon's institutional preference shifts away from consensus---not because multilateral cooperation is less valuable, but because the hegemon can capture a larger share of that value bilaterally. This is consistent with the observation that hegemonic support for consensus-based institutions tends to weaken precisely when bilateral alternatives become more attractive.

```{r alpha-star-plot, echo=FALSE, fig.width=7, fig.height=4.5, fig.cap="Threshold $\\alpha^*(N, \\beta)$ for conditional payoff dominance (Lemma 1) as a function of organization size $N$, for three discount factors. Below each curve, unanimity gives the hegemon a strictly higher continuation payoff than majority for all beliefs; above it, the dominance fails at $\\mu = 1$ (Lemma 1, necessity). The threshold decreases with $N$: in larger organizations, unanimity's vote-buying cost grows relative to the screening advantage, so the mechanism requires a weaker bilateral alternative."}

alpha_star <- function(N, beta) {
  q <- floor(N / 2) + 1
  num <- beta * (q - 1)
  den <- N * (N - 1) - beta * (N^2 - N - q + 1)
  if (den <= 0) return(NA)
  return(num / den)
}

Ns <- seq(10, 150, by = 10)
betas <- c(0.7, 0.8, 0.9)
cols <- c("gray50", "steelblue3", "firebrick3")

par(mar = c(4, 4.5, 1, 1))
plot(NULL, xlim = c(10, 150), ylim = c(0, 0.35),
     xlab = "Organization size (N)", ylab = expression(alpha^"*"),
     las = 1, xaxt = "n")
axis(1, at = seq(10, 150, by = 20))
abline(h = seq(0, 0.35, by = 0.05), col = "gray90", lty = 1)

for (i in seq_along(betas)) {
  vals <- sapply(Ns, function(n) alpha_star(n, betas[i]))
  lines(Ns, vals, col = cols[i], lwd = 2.5, type = "b", pch = 16, cex = 0.6)
}

legend("topright",
       legend = as.expression(lapply(betas, function(b) bquote(beta == .(b)))),
       col = cols, lwd = 2.5, pch = 16, pt.cex = 0.6, cex = 0.85, bg = "white")
```

The biconditional characterization in Lemma \ref{lem:conditional} implies that this boundary is sharp. At $\alpha = \alpha^*$, the dominance difference $D(\mu)$ equals zero exactly at $\mu = 1$; for $\alpha > \alpha^*$, majority strictly dominates near $\mu = 1$ (where informational asymmetry is irrelevant and the vote-buying cost is decisive), while unanimity can still dominate at lower beliefs where the screening advantage is strong. In this regime, the institutional comparison depends on the prior $p$ and cannot be resolved by a single lemma. The paper's main results are stated for $\alpha < \alpha^*$, which ensures a clean separation between the conditional and entry channels.

Figure \@ref(fig:heatmap-alpha-mu) maps the conditional dominance region in the $(\alpha, \mu)$ plane. For each parameter combination, the blue region indicates beliefs at which unanimity gives the hegemon a higher continuation payoff; the red region indicates beliefs at which majority dominates. The solid curve traces the frontier $\bar\mu(\alpha)$ from Remark \ref{rem:mu_bar}. Two patterns stand out. First, the frontier is remarkably flat: increasing $\alpha$ well beyond $\alpha^*$ shifts $\bar\mu$ only modestly downward, confirming that the screening advantage is robust across most of the belief space. Second, the red region is always concentrated near $\mu = 1$---near certainty about the magnitude of the gains from cooperation---where informational asymmetry is irrelevant and the cost of buying unanimous agreement dominates.

```{r heatmap-alpha-mu, echo=FALSE, fig.width=10, fig.height=8, fig.cap="Conditional payoff dominance in the $(\\alpha, \\mu)$ plane. Blue: unanimity dominates ($D(\\mu) > 0$). Red: majority dominates ($D(\\mu) < 0$). Solid curve: belief threshold $\\bar\\mu(\\alpha)$ above which majority dominates (Remark 1). Dashed line: $\\alpha^*$. Left column: $r = 1.5$. Right column: $r = 2$. Top row: $N = 30$. Bottom row: $N = 164$. All panels: $\\beta = 0.9$."}

source("scripts/model_functions.R")

alpha_star_fn <- function(N, beta) {
  q <- floor(N / 2) + 1
  beta * (q - 1) / (N * (N - 1) - beta * (N^2 - N - q + 1))
}

D_payoff <- function(alpha, mu, r, N, beta) {
  vh_u <- tryCatch(VH_R1_unanimity(r, alpha, mu, N, beta), error = function(e) NA)
  vh_m <- tryCatch(VH_R1_majority(r, alpha, mu, N, beta), error = function(e) NA)
  if (is.na(vh_u) || is.na(vh_m)) return(NA)
  return(vh_u - vh_m)
}

find_mu_bar <- function(alpha, r, N, beta, tol = 1e-6) {
  a_star <- alpha_star_fn(N, beta)
  if (alpha < a_star) return(NA)
  d1 <- D_payoff(alpha, 1, r, N, beta)
  if (is.na(d1) || d1 > 0) return(NA)
  d_low <- D_payoff(alpha, 0.01, r, N, beta)
  if (is.na(d_low) || d_low <= 0) return(0)
  lo <- 0.01; hi <- 1
  while (hi - lo > tol) {
    mid <- (lo + hi) / 2
    d_mid <- D_payoff(alpha, mid, r, N, beta)
    if (is.na(d_mid)) { hi <- mid; next }
    if (d_mid > 0) lo <- mid else hi <- mid
  }
  return((lo + hi) / 2)
}

plot_heatmap <- function(r, N, beta, n_grid = 150, main_title = NULL) {
  alpha_max <- min(1/r - 0.01, 0.5)
  alphas <- seq(0.005, alpha_max, length.out = n_grid)
  mus <- seq(0.01, 1, length.out = n_grid)
  D_matrix <- matrix(NA, nrow = n_grid, ncol = n_grid)
  for (i in seq_along(alphas))
    for (j in seq_along(mus))
      D_matrix[i, j] <- D_payoff(alphas[i], mus[j], r, N, beta)

  a_star <- alpha_star_fn(N, beta)
  if (is.null(main_title))
    main_title <- bquote(r == .(r) ~ ", " ~ beta == .(beta) ~ ", N=" ~ .(N))

  cols <- ifelse(D_matrix > 0, "#3182BD80", "#E6550D80")
  cols[is.na(D_matrix)] <- "grey90"
  image(alphas, mus, D_matrix, col = NA, xlab = expression(alpha),
        ylab = expression(mu), main = main_title, cex.main = 1)
  da <- diff(alphas[1:2]); dm <- diff(mus[1:2])
  for (i in seq_along(alphas))
    for (j in seq_along(mus))
      rect(alphas[i] - da/2, mus[j] - dm/2,
           alphas[i] + da/2, mus[j] + dm/2,
           col = cols[i, j], border = NA)

  alpha_curve <- seq(a_star, alpha_max, length.out = 100)
  mu_bar_curve <- sapply(alpha_curve, function(a) find_mu_bar(a, r, N, beta))
  valid <- !is.na(mu_bar_curve) & mu_bar_curve > 0
  if (any(valid))
    lines(alpha_curve[valid], mu_bar_curve[valid], lwd = 2.5, col = "black")

  abline(v = a_star, lty = 2, col = "black", lwd = 1.5)
  text(a_star, 0.05, expression(alpha * "*"), pos = 4, cex = 0.9)

  legend("topleft",
         legend = c("Unanimity dom.", "Majority dom.",
                    expression(bar(mu)(alpha))),
         fill = c("#3182BD80", "#E6550D80", NA),
         border = c("black", "black", NA),
         lty = c(NA, NA, 1), lwd = c(NA, NA, 2.5), col = c(NA, NA, "black"),
         merge = TRUE, cex = 0.65, bg = "white")
}

par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))
plot_heatmap(r = 1.5, N = 30,  beta = 0.9)
plot_heatmap(r = 2.0, N = 30,  beta = 0.9)
plot_heatmap(r = 1.5, N = 164, beta = 0.9)
plot_heatmap(r = 2.0, N = 164, beta = 0.9)
```

**Alternative explanations.** The informational mechanism is complementary to, not a substitute for, other rationalist explanations of consensus. Legitimacy and compliance accounts emphasize that consent enhances durability of agreements, but address why agreements are honored rather than why a hegemon benefits from the bargaining structure. Self-enforcing agreement models [@maggi2006selfenforcing] show that unanimity can be optimal when enforcement is weak, but apply primarily to symmetric states and collective action problems rather than to asymmetric players with private information. Issue linkage accounts explain why broad agendas favor consensus but not why a hegemon with superior information would prefer unanimity in a single-issue environment. This paper isolates a distributive-informational mechanism within the rationalist framework common to these accounts, in the spirit of @fearon1995rationalist.

**Broader implications.** Institutions do not merely reflect power; they select which form of power becomes politically productive. Majority rule rewards coalition arithmetic; unanimity rewards informational influence. This reframing implies a dynamic: as weaker states invest in independent analytical capacity, the informational asymmetry narrows, the returns to persuasion decline, and the hegemon's institutional incentive to prefer consensus should weaken. These dynamic implications are suggestive extrapolations from the static mechanism, not formal predictions of the model. A proper treatment would require a repeated game with endogenous information acquisition, which is beyond the scope of this paper. The value of consensus to a hegemon should covary with the distribution of technical expertise across the membership.


# Conclusion

This paper has shown when and why consensus can be a rational choice for a hegemon. The key is to distinguish between two modes of exercising power under different voting rules. Majority rule with symmetric proposals lets weak states bypass the hegemon, eliminating the screening problem. Unanimity forces weak states to include the hegemon under uncertainty, creating a screening cutoff that generates a non-convexity in the hegemon's value function. Bayesian persuasion exploits that non-convexity. The institutional comparison reduces to a single threshold: above it, unanimity dominates; below it, majority can dominate through easier entry. The ranking never oscillates.

The argument does not imply that all consensus institutions are designed for this reason, or that informal power disappears under consensus. It shows something more specific: under identifiable conditions on informational asymmetry and the prospects of cooperation, unanimity becomes the preferred institutional arrangement for a hegemon. Consensus may be less a concession by power than a way of exercising power through a different channel---the channel of information.

The mechanism depends on a discrete type space. With $K > 2$ types, Appendix C shows that unanimity generates $K-1$ screening boundaries while majority remains linear---the non-convexity structure becomes richer, not weaker. With a continuous type space, however, the offer schedule would generically be smooth, eliminating the discrete jumps that Bayesian persuasion exploits. Discrete types are substantively natural in this setting---states evaluate cooperation in terms of qualitatively distinct regimes rather than marginal variations---but whether informational power retains strategic value under continuous types is an open question.

Several extensions remain. First, if weaker states can invest in independent analytical capacity, informational power erodes endogenously over time. Second, heterogeneous outside options across weak states (middle powers) would enrich the screening dynamics. Third, repeated play within an established consensus institution could generate learning that narrows the informational gap. These directions point toward a richer theory of how informational power rises and declines within institutional frameworks.


# (APPENDIX) Appendix {-}

# Appendix A: Bargaining Derivations {-}

## Notation summary {-}

\begin{table}[H]
\centering\small
\begin{tabular}{lll}
\toprule
Symbol & Meaning & Defined in \\
\midrule
$N$ & Number of players ($1$ hegemon $+ N-1$ weak states) & Def.\ 1 \\
$\theta \in \{0,1\}$ & State of the world & Def.\ 1 \\
$p$ & Prior $\Pr(\theta=1)$ & Def.\ 1 \\
$\mu$ & Posterior belief $\Pr(\theta=1 \mid s)$ & Sec.\ 4 \\
$V(\theta)$ & Value of cooperation: $V(0)=1$, $V(1)=r$ & Def.\ 1 \\
$r > 1$ & High-type value & Def.\ 1 \\
$\alpha \in (0, 1/r)$ & Outside-option share: $d_H = \alpha V(\theta)$ & Def.\ 1 \\
$\beta \in (0,1)$ & Common discount factor & Def.\ 1 \\
$c > 0$ & Entry cost per weak state & Def.\ 1 \\
$V_e(\mu)$ & Expected cooperation value $1 + \mu(r-1)$ & Sec.\ 4 \\
$x$ & Shorthand $(N-1)\alpha r$ & Sec.\ 4 \\
$q$ & Majority quota $\lfloor N/2 \rfloor + 1$ & Sec.\ 5 \\
$R \in \{M, U\}$ & Voting rule (majority / unanimity) & Def.\ 1 \\
$\mu_s^{R2}$ & R2 screening cutoff & App.\ A.2 \\
$\mu_s^{R1}$ & R1 screening cutoff & Prop.\ 2 \\
$\phi$ & Auxiliary for R1 cutoff: $[rN - \beta(N\!-\!1\!+\!r)]/[\beta(r\!-\!1)]$ & Prop.\ 2 \\
$\tau(R)$ & Entry threshold under rule $R$ & Sec.\ 7 \\
$v(\mu, R)$ & Hegemon's reduced-form value function & Sec.\ 7 \\
$\alpha^*$ & Iff threshold for conditional dominance (Lemma 1) & Lemma 1 \\
$C_{\text{buy}}, C_{\text{out}}$ & Vote-buying cost, outside-option cost (proof) & App.\ B.5 \\
$\lambda_M$ & Majority payoff coefficient & Thm.\ 1 \\
$S_U, S_M$ & Max-slope of value function under $U$, $M$ & Thm.\ 1 \\
$p^*$ & Threshold prior for institutional comparison & Thm.\ 1 \\
\bottomrule
\end{tabular}
\end{table}

## A.1 Majority: R2 and R1 {-}

Under majority, weak proposers form a winning coalition from other weak states, excluding $H$. The hegemon captures $\alpha V(\theta)$ from its bilateral alternative. $H$'s payoff is type-specific and does not depend on $W$'s beliefs.

R2 continuation values:
\begin{align}
V_H^{R2}(\theta, M) &= \frac{V(\theta)[1+(N-1)\alpha]}{N} \\
V_W^{R2}(\mu, M) &= \frac{(1-\alpha)V_e(\mu)}{N}
\end{align}

R1 values follow by backward induction with the same structure (no screening, linear).

## A.2 Unanimity: R2 {-}

When $H$ proposes: offers $0$ to all $W$'s, keeps $V(\theta)$.

When $W$ proposes: the screening subgame has cutoff $\mu_s^{R2} = \alpha(r-1)/(r-\alpha)$. Below it, $W$ plays aggressive ($y_H = \alpha$); above, conservative ($y_H = \alpha r$).

The hegemon's R2 continuation values under unanimity are:
\begin{align}
V_H^{R2}(\theta=1, \mu) &= \frac{r[1+(N-1)\alpha]}{N} \quad \text{(constant in $\mu$)}, \label{eq:VH1_R2_U} \\
V_H^{R2}(\theta=0, \mu < \mu_s^{R2}) &= \frac{1 + (N-1)\alpha}{N}, \label{eq:VH0_R2_agg} \\
V_H^{R2}(\theta=0, \mu > \mu_s^{R2}) &= \frac{1 + (N-1)\alpha r}{N} \quad \text{(overpaid)}. \label{eq:VH0_R2_con}
\end{align}

Weak states' R2 continuation values under unanimity:
\begin{align}
V_W^{R2}(\mu < \mu_s^{R2}) &= \frac{(1-\mu)(1-\alpha)}{N} \\
V_W^{R2}(\mu > \mu_s^{R2}) &= \frac{V_e(\mu) - \alpha r}{N}
\end{align}

Continuity at $\mu_s^{R2}$ is verified by direct substitution.

## A.3 Unanimity: R1 screening derivation {-}

The proposing $W$ compares:
\begin{align}
F_1^{con}(\mu) &= V_e(\mu) - \frac{\beta(r+x)}{N} - \omega(\mu) \\
F_1^{agg}(\mu) &= (1-\mu)\left[1 - \frac{\beta(1+x)}{N} - \omega(\mu)\right] + \frac{\mu\beta r(1-\alpha)}{N}
\end{align}
where $\omega(\mu) = (N-2)\beta V_W^{R2}(\mu)$.

The difference $\Delta_1(\mu) = F_1^{agg} - F_1^{con}$ is piecewise quadratic. On the high branch ($\mu > \mu_s^{R2}$), the $\alpha$-dependent terms cancel, yielding the quadratic:
\[
(N-2)\beta(r-1)\mu^2 + [\beta(N-1+r) - rN]\mu + \beta(r-1) = 0.
\]

Dividing by $\beta(r-1)$:
\[
(N-2)\mu^2 - \phi\mu + 1 = 0, \qquad \phi = \frac{rN - \beta(N-1+r)}{\beta(r-1)}.
\]

The smaller root is $\mu_s^{R1}$ (equation \eqref{eq:cutoff_R1}).

## A.4 Jump magnitude {-}

At the R1 cutoff, type $\theta=0$ receives $\beta(1+x)/N$ (aggressive) vs $\beta(r+x)/N$ (conservative) when $W$ proposes. The difference $\beta(r-1)/N$, weighted by $(N-1)/N$ (prob $W$ proposes) and $(1-\mu_s^{R1})$ (prob $\theta=0$), gives the jump in equation \eqref{eq:jump_R1}.

## A.5 Case selection {-}

The R1 cutoff is on the high branch (Case 2, $\alpha$-independent) when $\Delta_1(\mu_s^{R2}) > 0$. This is equivalent to $\alpha < \bar\alpha(r, \beta, N)$, which holds throughout the empirically relevant parameter range.

## A.6 Budget checks {-}

Under majority, the budget identity $E[V_H] + (N-1)V_W = V_e(\mu)$ holds exactly: $V_H[1+(N-1)\alpha]/N + (N-1)(1-\alpha)V_e(\mu)/N = V_e(\mu) \cdot [1+(N-1)\alpha + (N-1)(1-\alpha)]/N = V_e(\mu)$. Under unanimity, the budget identity $E[V_H] + (N-1)V_W = V_e(\mu)$ holds on the conservative R1 branch, where all deals pass in R1. On the aggressive R1 branch, surplus destruction from discounting when $\theta=1$ rejects in R1 and the game proceeds to R2 yields $E[V_H] + (N-1)V_W = V_e(\mu) - \frac{(N-1)\mu r(1-\beta)}{N}$, so the identity holds with inequality: $E[V_H] + (N-1)V_W \leq V_e(\mu)$.

## A.7 Entry thresholds {-}

Under majority, $V_W^{R1}(\mu, M) = \kappa_M V_e(\mu)$ where $\kappa_M = (1-\alpha)[N(N-1)+\beta(q-1)]/(N^2(N-1))$. The entry threshold is $\tau(M) = \max\{0,\; (c/\kappa_M - 1)/(r-1)\}$.

Under unanimity, $V_W^{R1}$ is piecewise. On the conservative R1 branch ($\mu > \mu_s^{R1}$, which also implies $\mu > \mu_s^{R2}$ in the principal case):
$$V_W^{R1}(\mu, U) = \frac{(N+\beta)\,V_e(\mu) - \beta r(1+N\alpha)}{N^2}.$$
This is affine in $\mu$. Setting $V_W^{R1} = c$ and solving:
\begin{equation}\label{eq:tau_U_con}
\tau_U^{con} = \frac{cN^2 - \bigl[N + \beta - \beta r(1+N\alpha)\bigr]}{(N+\beta)(r-1)}.
\end{equation}

On the aggressive R1 branch ($\mu < \mu_s^{R1}$), $V_W^{R1}$ is quadratic in $\mu$ (the product of $(1-\mu)$ with the affine terms in $\omega(\mu)$ generates a second-order term), and the entry threshold solves a quadratic equation with explicit roots.^[Because $V_W^{R1}$ has a downward jump at $\mu_s^{R1}$ (opposite to $V_H$'s upward jump), the entry set $\{\mu : V_W^{R1} \geq c\}$ can be disconnected for intermediate values of $c$: entry may occur both below $\mu_s^{R1}$ (aggressive regime) and above some $\mu_2 > \mu_s^{R1}$ (conservative regime), with a gap around the cutoff. The affine formula \eqref{eq:tau_U_con} gives the entry threshold in the conservative regime and is the relevant quantity when $c > \max_{\mu \leq \mu_s^{R1}} V_W^{R1}(\mu, \text{agg})$.]

The numerical results in the paper compute $\tau(U)$ by pointwise evaluation of $V_W^{R1}$.


# Appendix B: Proofs {-}

## B.1 Proof of Proposition 1 (Majority produces no screening) {-}

Under majority rule, $W$'s proposal passes with $q-1$ votes from other weak states. $H$'s vote is not required. $W$ excludes $H$ from the winning coalition; $H$ captures $\alpha V(\theta)$ from its bilateral alternative regardless of $\mu$. There is no screening and no jump. All continuation values are affine in $V_e(\mu)$, hence affine in $\mu$. $\square$

## B.2 Proof of Proposition 2 (R1 cutoff existence and uniqueness) {-}

See Appendix A.3 for the derivation. Existence: $\Delta_1(0) = \beta(r-1)/N > 0$ and $\Delta_1(1) = r(\beta-1) < 0$; by continuity, a root exists. Uniqueness: on the relevant branch, $\Delta_1$ is a convex quadratic with exactly one root in $(0,1)$ (the set where $\Delta_1 \leq 0$ is an interval by convexity; since $\Delta_1(\mu_s^{R2}) > 0$ and $\Delta_1(1) < 0$, the first root is unique). $\square$

## B.3 Proof of Proposition 3 (Jump) {-}

Direct computation from the difference in H's R1 payoff between the conservative and aggressive branches. See Appendix A.4. $\square$

## B.4 Proof of Proposition 4 (Additional non-convexity) {-}

Under majority, $v(\mu, M)$ is affine for $\mu \geq \tau(M)$, hence concave---the concave envelope coincides with $v$ above the threshold. Under unanimity, the jump at $\mu_s^{R1}$ means $v(\mu, U)$ is not concave on $[\tau(U), 1]$: $\lim_{\mu \uparrow \mu_s^{R1}} v(\mu, U) < \lim_{\mu \downarrow \mu_s^{R1}} v(\mu, U)$. The concave envelope must therefore lie strictly above $v(\mu, U)$ in a neighborhood of $\mu_s^{R1}$. Since $v(\mu, U)$ is affine on each branch (the expected payoff is linear in $\mu$ conditional on the weak proposer's strategy), the upward jump is the sole source of non-concavity. $\square$

## B.5 Proof of Lemma 1 (Conditional payoff dominance) {-}

Let $C_{\text{buy}} \equiv \beta(q-1)(1-\alpha)$ (the vote-buying cost under majority) and $C_{\text{out}} \equiv N(N-1)\alpha$ (the aggregate outside-option cost). Under majority, $E[V_H^{R1}(\mu, M)] = \lambda_M V_e(\mu)$ where $\lambda_M = [N(1+(N-1)\alpha) - C_{\text{buy}}]/N^2$.

**Step 1: Decomposition.** Define:
\begin{align}
D_{\text{base}}(\mu) &\equiv \frac{(C_{\text{buy}}-C_{\text{out}})V_e(\mu) + C_{\text{out}}\beta r}{N^2}, \\
\delta_{R2}(\mu) &\equiv \frac{(N-1)\beta[\mu(r-\alpha)-\alpha(r-1)]}{N^2}, \\
\delta_{R1}(\mu) &\equiv \frac{(N-1)\beta(r-1)(1-\mu)}{N^2}.
\end{align}

The difference $D(\mu) \equiv E[V_H^{R1}(\mu, U)] - \lambda_M V_e(\mu)$ decomposes as:
$$D(\mu) = D_{\text{base}}(\mu) + \mathbf{1}\{\mu < \mu_s^{R2}\}\cdot\delta_{R2}(\mu) + \mathbf{1}\{\text{R1 conservative}\}\cdot\delta_{R1}(\mu).$$

\textit{Verification of the decomposition.} On Region II (conservative R2, aggressive R1), $D = D_{\text{base}}$ by construction. The R2 correction $\delta_{R2}$ accounts for the difference between aggressive and conservative $V_W^{R2}$ under unanimity; it vanishes at $\mu_s^{R2}$ because the two R2 branches are continuous there: $\delta_{R2}(\mu_s^{R2}) = 0$. The R1 correction $\delta_{R1}$ accounts for the overpayment to type $\theta=0$ when $W$ plays conservative in R1; it satisfies $\delta_{R1}(\mu) \geq 0$ for all $\mu \leq 1$, with equality only at $\mu = 1$.

**Step 2: $D_{\text{base}} > 0$ on $[0,1]$.** $D_{\text{base}}$ is affine, so it suffices to check two endpoints.

*At $\mu = 1$:*
$$D_{\text{base}}(1) = \frac{r[C_{\text{buy}} - C_{\text{out}}(1-\beta)]}{N^2} = \frac{r[\beta(q-1)(1-\alpha) - N(N-1)\alpha(1-\beta)]}{N^2}.$$
Setting $C_{\text{buy}} - C_{\text{out}}(1-\beta) = 0$ and solving for $\alpha$ yields $\alpha = \alpha^*(N, \beta)$. Since $\alpha < \alpha^*$, $D_{\text{base}}(1) > 0$.

*At $\mu = 0$:*
$$D_{\text{base}}(0) = \frac{C_{\text{buy}} - C_{\text{out}} + C_{\text{out}}\beta r}{N^2} = \frac{C_{\text{buy}} + C_{\text{out}}(\beta r - 1)}{N^2}.$$
Since $D_I(0) = D_{\text{base}}(0) + \delta_{R2}(0) = D_{\text{base}}(0) - (N-1)\beta\alpha(r-1)/N^2$, we have $D_{\text{base}}(0) = D_I(0) + (N-1)\beta\alpha(r-1)/N^2 > D_I(0)$. To show $D_I(0) > 0$: write $D_I(0) = [\beta(q-1) - \alpha d_0]/N^2$ where $d_0 = N(N-1) - \beta((N-1)^2 r + N - q)$. If $d_0 \leq 0$, then $D_I(0) \geq \beta(q-1)/N^2 > 0$. If $d_0 > 0$, the threshold $\bar\alpha_0 = \beta(q-1)/d_0$ satisfies $\bar\alpha_0 > \alpha^*$ because $d_0 < d_* \equiv N(N-1) - \beta(N^2 - N - q + 1)$ (since $d_* - d_0 = \beta(N-1)^2(r-1) > 0$). Hence $\alpha < \alpha^* < \bar\alpha_0$ implies $D_I(0) > 0$, and therefore $D_{\text{base}}(0) > 0$.

*Conclusion:* $D_{\text{base}}$ is affine with positive values at both endpoints, hence $D_{\text{base}}(\mu) > 0$ for all $\mu \in [0,1]$.

**Step 3: Assembly across regions.**

*Region I* ($\mu < \mu_s^{R2}$, aggressive R2 and aggressive R1): $D = D_{\text{base}} + \delta_{R2}$, affine between $D_I(0) > 0$ (at $\mu = 0$) and $D_{\text{base}}(\mu_s^{R2}) > 0$ (at $\mu_s^{R2}$, since $\delta_{R2}(\mu_s^{R2}) = 0$). Hence $D > 0$.

*Region II* ($\mu_s^{R2} < \mu < \mu_s^{R1}$, conservative R2 and aggressive R1): $D = D_{\text{base}} > 0$ directly from Step 2.

*Region III* ($\mu > \mu_s^{R1}$, conservative R1): $D = D_{\text{base}} + \delta_{R1} \geq D_{\text{base}} > 0$, since $\delta_{R1} \geq 0$.

*Jump at $\mu_s^{R1}$:* the R1 correction $\delta_{R1}$ activates, adding $(N-1)\beta(r-1)(1-\mu_s^{R1})/N^2 > 0$. The left limit is positive (Region II) and the right limit is strictly larger.

Combining all regions: $D(\mu) > 0$ for every $\mu \in (0,1)$. At $\mu = 1$: $D(1) = D_{\text{base}}(1) = r[C_{\text{buy}} - C_{\text{out}}(1-\beta)]/N^2 > 0$ (Step 2), so the strict inequality extends to the closed interval $(0,1]$. This establishes sufficiency.

**Step 4: Necessity.** Suppose $\alpha \geq \alpha^*$. At $\mu = 1$, the R1 correction vanishes: $\delta_{R1}(1) = (N-1)\beta(r-1)(1-1)/N^2 = 0$. Since $\mu = 1 > \mu_s^{R2}$, the R2 correction is also absent. Therefore $D(1) = D_{\text{base}}(1) = r[C_{\text{buy}} - C_{\text{out}}(1-\beta)]/N^2$. The threshold $\alpha^*$ is defined as the root of $C_{\text{buy}} - C_{\text{out}}(1-\beta) = 0$, so $\alpha \geq \alpha^*$ implies $D(1) \leq 0$. Hence the strict inequality $D(\mu) > 0$ for all $\mu \in (0,1]$ fails. $\square$

## B.6 Proof of Theorem 1 (Threshold prior) {-}

To state the proof, define the majority payoff coefficient $\lambda_M \equiv [N[1+(N-1)\alpha]-\beta(q-1)(1-\alpha)]/N^2$, so that $v(\mu, M) = \lambda_M V_e(\mu)$ whenever entry occurs. Let $\kappa_M \equiv (1-\alpha)[N(N-1)+\beta(q-1)]/(N^2(N-1))$ and $\tau(M)=\max\{0,\;(c/\kappa_M-1)/(r-1)\}$. Under unanimity, let $\tau(U)$ denote the entry threshold and define $S_U \equiv \max_{\mu \geq \tau(U)} v(\mu, U)/\mu$. If $\tau(M) > 0$, let $S_M \equiv v(\tau(M), M)/\tau(M)$.

*Step 1: High priors ($p \geq \tau(U)$).* Entry occurs under both rules. By Lemma \ref{lem:conditional}, $v(p, U) > v(p, M)$. Under majority, $v(\mu, M)$ is affine above $\tau(M)$, hence already concave, so $\operatorname{cav} v(p, M) = v(p, M)$. Therefore $\operatorname{cav} v(p, U) \geq v(p, U) > v(p, M) = \operatorname{cav} v(p, M)$. This proves part (a) when $\tau(U) = 0$.

*Step 2: Concavification below $\tau(U)$.* Assume $\tau(U) > 0$. Since $v(\mu, U) = 0$ for $\mu < \tau(U)$ and $v(0, U) = 0$, the concavification on $[0, \mu_U^*]$ (where $\mu_U^*$ achieves $S_U$) is $\operatorname{cav} v(p, U) = S_U \cdot p$. Under majority: if $\tau(M) = 0$, then $\operatorname{cav} v(p, M) = \lambda_M V_e(p)$; if $\tau(M) > 0$, then $\operatorname{cav} v(p, M) = S_M \cdot p$ for $p < \tau(M)$ and $\lambda_M V_e(p)$ for $p \geq \tau(M)$.

*Step 3: Case $\tau(M) = 0 < \tau(U)$.* For $p \in (0, \tau(U))$, $D(p) = S_U \cdot p - \lambda_M V_e(p) = [S_U - \lambda_M(r-1)]p - \lambda_M$, which is linear. $D(0) = -\lambda_M < 0$; by Step 1, $D(\tau(U)) > 0$. So $D$ has a unique zero $p^* = \lambda_M/[S_U - \lambda_M(r-1)] \in (0, \tau(U))$.

*Step 4: Case $0 < \tau(M) \leq \tau(U)$, $S_U > S_M$.* For $p < \tau(M)$: $\operatorname{cav} v(p, U) = S_U p > S_M p = \operatorname{cav} v(p, M)$. For $p \in [\tau(M), \tau(U))$: $D(\tau(M)) = (S_U - S_M)\tau(M) > 0$ and $D(\tau(U)) > 0$ by Step 1. Unanimity dominates globally.

*Step 5: Case $0 < \tau(M) \leq \tau(U)$, $S_U \leq S_M$.* For $p < \tau(M)$: $\operatorname{cav} v(p, U) \leq \operatorname{cav} v(p, M)$. For $p \in [\tau(M), \tau(U))$: $D(\tau(M)) \leq 0$ and $D(\tau(U)) > 0$, so $D$ crosses zero once at $p^* \in [\tau(M), \tau(U)]$.

Since each case admits at most one crossing, the comparison has the single-crossing property. $\square$


# Appendix C: Robustness to $K > 2$ Types {-}

This appendix shows that the screening mechanism and the majority-linearity result both generalize to $K > 2$ types.

## C.1 Setup {-}

The model is identical to Section 4 except that $\theta \in \{0, 1, 2\}$ with $V(0) = 1$, $V(1) = r_1$, $V(2) = r_2$, where $1 < r_1 < r_2$. The prior is $p = (p_0, p_1, p_2) \in \Delta^2$, the 2-simplex. The expected value of cooperation is $V_e(\mu) = \sum_k \mu_k V(\theta_k)$. All other primitives---$N$ players, discount factor $\beta$, random proposer, outside option $d_H = \alpha V(\theta)$---are unchanged. The constraint on the outside option becomes $\alpha \in (0, 1/r_2)$, ensuring that $d_H < V(\theta)$ for all types.

## C.2 R2 Screening under Unanimity {-}

With $K$ types, the R2 screening problem has a natural generalization. When $W$ makes an offer in R2, it targets a particular type by setting the reservation value equal to $\alpha V(\theta_k)$ for some $k$. All types $\theta_j$ with $j \leq k$ accept (since $\alpha V(\theta_j) \leq \alpha V(\theta_k)$), and all types $\theta_j$ with $j > k$ reject. The optimal target balances the probability of acceptance against the cost of overpaying low types.

\begin{proposition}[R2 Screening Boundaries for General $K$]\label{prop:k_general_screening}
For $K$ ordered types with values $v_1 < v_2 < \cdots < v_K$, the R2 screening boundary between targeting type $k$ and type $k+1$ is defined by the posterior ratio
\[
\frac{\mu_{k+1}}{\sum_{j \leq k} \mu_j} = \frac{\alpha(v_{k+1} - v_k)}{v_{k+1}(1-\alpha)}.
\]
These $K-1$ conditions define hyperplanes that partition the simplex $\Delta^{K-1}$ into $K$ screening regions.
\end{proposition}

\begin{proof}[Proof sketch]
The weak proposer's payoff from an offer targeting type $k$ is $\sum_{j \leq k} \mu_j [V(\theta_j) - \alpha V(\theta_k)]$. Equating the payoffs from targeting type $k$ and type $k+1$ yields the indifference condition. The ratio simplifies because the incremental gain from including type $k+1$ (at cost $\alpha V(\theta_{k+1})$) must equal the incremental cost of overpaying all types $j \leq k$ (the additional payment $\alpha[V(\theta_{k+1}) - V(\theta_k)]$ per accepting type). For $K=2$, this reduces to $\mu_s^{R2} = \alpha(r-1)/(r-\alpha)$, recovering the main model.
\end{proof}

For $K = 3$ with values $(1, r_1, r_2)$, the R2 screening produces three regions separated by two boundaries. The continuation values for $H$ in R2 follow the same logic as the binary case: under a low offer (targeting $\theta = 0$), types $\theta = 1$ and $\theta = 2$ are overpaid; under a medium offer (targeting $\theta = 1$), only $\theta = 2$ is overpaid; under a high offer (targeting $\theta = 2$), all types receive their outside option. The key pattern persists: low types receive strictly more than their outside option whenever $W$ targets a higher type, and the highest type $\theta = 2$ always receives $r_2(1 + (N-1)\alpha)/N$ regardless of the screening region.

## C.3 Majority: No Screening {-}

\begin{proposition}[Linearity of $H$'s Payoff under Majority for General $K$]\label{prop:k_majority_linear}
Under majority rule with $K$ types, $E_\theta[V_H^{R1}(\theta, \mu, M)] = \lambda_M V_e(\mu)$, where $\lambda_M$ is defined as in Theorem \ref{thm:main}. No screening cutoff exists for any $K$.
\end{proposition}

\begin{proof}
The proof is identical to the $K = 2$ case. Under majority, $W$ never needs $H$'s vote; $H$ receives $\alpha V(\theta)$ in every coalition that forms and $\alpha V(\theta)$ as its outside option. Since $H$'s payoff is $\alpha V(\theta)$ type by type, the expectation is $\alpha V_e(\mu)$ scaled by the bargaining multiplier $\lambda_M / \alpha$. The belief $\mu$ enters only through $V_e(\mu)$, so the payoff is affine in $\mu$ (and hence in $V_e$) regardless of $K$.
\end{proof}

Together, Propositions \ref{prop:k_general_screening} and \ref{prop:k_majority_linear} establish the key asymmetry for general $K$: unanimity generates $K-1$ screening boundaries (each a potential source of non-convexity), while majority remains linear regardless of $K$. The binary model is the minimal case that captures this mechanism.


# References {-}
