---
title: "FormalizaĂ§ĂŁo do Modelo v2"
author: "Manoel Galdino"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    number_sections: TRUE
    toc: FALSE
    latex_engine: xelatex
header-includes:
  - \usepackage{setspace}
  - \setstretch{1.3}
  - \usepackage{parskip}
  - \setlength{\parskip}{6pt}
  - \setlength{\parindent}{0pt}
  - \usepackage{float}
  - \usepackage{amsmath,amssymb,amsthm}
  - \newtheorem{definition}{Definition}
  - \newtheorem{lemma}{Lemma}
  - \newtheorem{proposition}{Proposition}
  - \newtheorem{theorem}{Theorem}
  - \newtheorem{corollary}{Corollary}
  - \newtheorem{remark}{Remark}
  - \newtheorem{example}{Example}
  - \usepackage{tikz}
  - \usetikzlibrary{trees,arrows.meta,positioning,calc,shapes.geometric}
  - \usepackage{pdflscape}
fontsize: 12pt
geometry: margin=2.5cm
mainfont: Times New Roman
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

# Environment

\begin{definition}[Institutional Design Game]\label{def:game}
The game $\Gamma(N, p, r, \alpha, \beta, c)$ has the following primitives.

\begin{itemize}
\item There are $N \geq 3$ players: one hegemon $H$ and $n = N-1$ weak states $W_1, \dots, W_n$.
\item Nature draws $\theta \in \{0, 1\}$ with prior $\Pr(\theta = 1) = p \in (0,1)$.
\item The state determines the value of multilateral cooperation: $V(0) = 1$ and $V(1) = r > 1$.
\item The hegemon observes $\theta$; the weak states do not.
\item If bargaining fails (disagreement):
  \begin{itemize}
  \item Each weak state receives $0$.
  \item The hegemon receives $\alpha V(\theta)$, where $\alpha \in (0, 1/r)$, representing bilateral alternatives proportional to the value of cooperation.
  \end{itemize}
\item The common discount factor is $\beta \in (0,1)$.
\item Each weak state pays an entry cost $c > 0$ to join the institution.
\end{itemize}
\end{definition}

Let
\[
V_e(\mu) \equiv 1 + \mu(r-1)
\]
denote the expected value of cooperation at posterior belief $\mu$.

\begin{remark}[Normalization]\label{rem:normalization}
The assumption $d_W = 0$ is without loss of generality. It requires only that (i) weak states' outside options are strictly lower than the hegemon's, and (ii) outside options are symmetric across weak states. Given these conditions, we redefine the pie as the surplus above the weak states' common outside option, which normalizes their disagreement payoff to zero.
\end{remark}

\begin{remark}[Bilateral alternatives]\label{rem:bilateral}
The hegemon's disagreement payoff $\alpha V(\theta)$ captures bilateral alternatives that are proportional to the value of multilateral cooperation. The same factors that make multilateral cooperation valuable (favorable trade environment, complementary economies) also improve bilateral alternatives. The parameter $\alpha < 1/r$ ensures that bilateral alternatives are strictly dominated by multilateral cooperation for all $\theta$.
\end{remark}

# Institutional Packages

\begin{definition}[Institutional Packages]\label{def:packages}
The analysis compares two institutional packages that differ only in the voting rule. Both packages have symmetric proposal rights (random proposer with probability $1/N$ each).

\begin{itemize}
\item \textbf{Package M (Majority)}: decisions pass with strict majority support ($q = \lfloor N/2 \rfloor + 1$ votes).
\item \textbf{Package U (Unanimity/Consensus)}: decisions require unanimous support ($q = N$ votes).
\end{itemize}
\end{definition}

\begin{remark}[Why not agenda control?]\label{rem:agenda}
By agenda control we mean here monopoly over proposal rights (exclusive proposer). Under either voting rule, if the hegemon were the exclusive proposer and $d_W = 0$, weak states' expected bargaining payoff would be zero. No weak state would pay the entry cost $c > 0$, and the institution would never form. Monopoly proposal power is therefore self-defeating. The relevant institutional choice is between voting rules, holding proposal rights symmetric. Other forms of agenda influence---such as shaping the set of feasible proposals---are discussed in the extension.
\end{remark}

# Timing

The game has three stages.

\begin{enumerate}
\item \textbf{Stage 0 (Institutional choice).} The hegemon chooses the voting rule $R \in \{M, U\}$.

\item \textbf{Stage 1 (Information and entry).} Nature draws $\theta$. The hegemon observes $\theta$ and commits to a public signal structure $\pi: \{0,1\} \to \Delta(S)$. Weak states observe the realized signal $s \in S$, update to posterior $\mu = \Pr(\theta = 1 \mid s)$, and simultaneously decide whether to enter (paying cost $c$).

\item \textbf{Stage 2 (Bargaining).} If the institution forms, members play a two-round Baron-Ferejohn bargaining game over a pie of size $V(\theta)$:
\begin{itemize}
\item \textbf{Round 1}: a proposer is selected uniformly at random (probability $1/N$). The proposer offers a division. All members vote according to rule $R$. If accepted, the pie is divided as proposed. If rejected, the game proceeds to Round 2.
\item \textbf{Round 2 (terminal)}: a new proposer is selected uniformly at random. The proposer offers a division. All members vote. If accepted, the pie is divided. If rejected, each player receives their disagreement payoff ($0$ for each $W$, $\alpha V(\theta)$ for $H$).
\end{itemize}
\end{enumerate}

# Solution Concept

The solution concept is Perfect Bayesian Equilibrium (PBE). Strategies are sequentially rational given beliefs, and beliefs are updated by Bayes' rule on the equilibrium path. Off-path beliefs satisfy sequential rationality.

# Game Tree

The game tree is presented for the unanimity subgame (Package $U$). Under Package $M$ (majority), the structure is identical except that $W$ proposers exclude $H$ from the coalition, eliminating the screening problem.

\begin{landscape}

% ---- Figure 1: Main game tree (Stage 0 â†’ Stage 1 â†’ R1) ----
\begin{figure}[H]
\centering
\begin{tikzpicture}[
  grow=right,
  level distance=2.8cm,
  edge from parent/.style={draw,-{Latex[length=1.5mm]}},
  every node/.style={font=\scriptsize},
  player/.style={draw, circle, fill=white, inner sep=1pt, minimum size=5mm},
  nature/.style={draw, circle, fill=yellow!15, inner sep=1pt, minimum size=5mm},
  terminal/.style={draw, rectangle, fill=green!8, inner sep=2pt},
  subgame/.style={draw, rectangle, fill=orange!15, inner sep=3pt, rounded corners},
  infoset/.style={dashed, thick, gray},
  level 1/.style={sibling distance=8cm},
  level 2/.style={sibling distance=4cm},
  level 3/.style={sibling distance=2cm},
  level 4/.style={sibling distance=3cm},
  level 5/.style={sibling distance=1.5cm}
]

\node [player, label=above:{Stage 0}] (root) {$H$}
child {
  node [nature] (nat0) {$N$}
  child {
    node [player] (w0) {$W$}
    child {
      node [nature, label=above:{\tiny R1}] (r1p0) {$N$}
      child {
        % R1, W proposes, theta=0 â†’ H decides
        node [player] (r1h0) {$H$}
        child { node [subgame] {R2 subgame} edge from parent node[above] {rej} }
        child { node [terminal] {$x_H^{1W},\; x_W^{1W}{-c}$} edge from parent node[below] {acc} }
        edge from parent node[above] {\tiny $W$ pr.\ $y_H$}
      }
      child {
        % R1, H proposes, theta=0 â†’ W votes
        node [player] (r1wv0) {$W$}
        child { node [subgame] {R2 subgame} edge from parent node[above] {rej} }
        child { node [terminal] {$x_H^{1H},\; x_W^{1H}{-c}$} edge from parent node[below] {acc} }
        edge from parent node[below] {\tiny $H$ pr.}
      }
      edge from parent node[above] {enter}
    }
    child {
      node [terminal] {$0,\; 0$}
      edge from parent node[below] {not}
    }
    edge from parent node[above] {$\theta=0$}
  }
  child {
    node [player] (w1) {$W$}
    child {
      node [nature, label=above:{\tiny R1}] (r1p1) {$N$}
      child {
        % R1, W proposes, theta=1 â†’ H decides
        node [player] (r1h1) {$H$}
        child { node [subgame] {R2 subgame} edge from parent node[above] {rej} }
        child { node [terminal] {$x_H^{1W},\; x_W^{1W}{-c}$} edge from parent node[below] {acc} }
        edge from parent node[above] {\tiny $W$ pr.\ $y_H$}
      }
      child {
        % R1, H proposes, theta=1 â†’ W votes
        node [player] (r1wv1) {$W$}
        child { node [subgame] {R2 subgame} edge from parent node[above] {rej} }
        child { node [terminal] {$x_H^{1H},\; x_W^{1H}{-c}$} edge from parent node[below] {acc} }
        edge from parent node[below] {\tiny $H$ pr.}
      }
      edge from parent node[above] {enter}
    }
    child {
      node [terminal] {$0,\; 0$}
      edge from parent node[below] {not}
    }
    edge from parent node[below] {$\theta=1$}
  }
  edge from parent node[above] {$R = U$}
};

% Information sets
\draw [infoset] (w0) -- node[midway, above, text=gray] {\tiny $W$'s info set} (w1);
\draw [infoset] (r1wv0) -- node[midway, below, text=gray] {\tiny $W$} (r1wv1);

\end{tikzpicture}
\caption{Main game tree under Package $U$ (unanimity): Stage~0 through Round~1. $H$ chooses $R$; Nature draws $\theta$; $H$ sends signal (not shown); $W$ forms posterior $\mu$ and decides entry (paying $c$ only if entering). In R1, Nature selects proposer ($1/N$). When $H$ proposes, $W$ votes to accept or reject. When $W$ proposes, $W$ offers $y_H$ to $H$, who accepts or rejects. Any rejection in R1 leads to the R2 subgame (Figure~\ref{fig:r2}). Dashed lines: $W$'s information sets ($W$ does not observe $\theta$). Under Package $M$, $W$ proposers exclude $H$, eliminating the screening branch.}\label{fig:r1}
\end{figure}

% ---- Figure 2: R2 subgame ----
\begin{figure}[H]
\centering
\begin{tikzpicture}[
  grow=right,
  level distance=3cm,
  edge from parent/.style={draw,-{Latex[length=1.5mm]}},
  every node/.style={font=\scriptsize},
  player/.style={draw, circle, fill=white, inner sep=1pt, minimum size=5mm},
  nature/.style={draw, circle, fill=yellow!15, inner sep=1pt, minimum size=5mm},
  terminal/.style={draw, rectangle, fill=green!8, inner sep=2pt},
  infoset/.style={dashed, thick, gray},
  level 1/.style={sibling distance=4cm},
  level 2/.style={sibling distance=2cm},
  level 3/.style={sibling distance=1.5cm}
]

\node [nature, label=above:{R2 (terminal)}] (r2root) {$N$}
child {
  % W proposes in R2 â†’ H decides
  node [player] (r2h) {$H$}
  child { node [terminal] {$\alpha V(\theta),\; {-c}$} edge from parent node[above] {rej $\to$ disagr.} }
  child { node [terminal] {$x_H^{2W},\; x_W^{2W}{-c}$} edge from parent node[below] {acc} }
  edge from parent node[above] {\tiny $W$ pr.\ $y_H$}
}
child {
  % H proposes in R2
  node [player] (r2wv) {$W$}
  child { node [terminal] {$\alpha V(\theta),\; {-c}$} edge from parent node[above] {rej $\to$ disagr.} }
  child { node [terminal] {$x_H^{2H},\; x_W^{2H}{-c}$} edge from parent node[below] {acc} }
  edge from parent node[below] {\tiny $H$ proposes}
};

\end{tikzpicture}
\caption{R2 subgame (terminal round) under unanimity. This subgame is reached only when $W$ plays aggressive in R1 and $\theta=1$ rejects. Therefore, $W$ \textbf{knows} $\theta=1$ (complete information --- no screening in R2). Nature selects proposer. When $W$ proposes, $W$ offers $\alpha r$ to $H$ (exact reservation of $\theta=1$). When $H$ proposes, $H$ offers $0$ to $W$. Rejection leads to disagreement: $\alpha V(1)$ for $H$, $-c$ for $W$.}\label{fig:r2}
\end{figure}

\end{landscape}

# Equilibrium Analysis (to be developed)

## Terminal Round (R2) under Unanimity

\textbf{When $H$ proposes} (prob $1/N$): offers $0$ to each $W$ ($d_W = 0$), all $W$'s accept (indifferent: $0 \geq 0$). $H$ keeps $V(\theta)$. This is the same for both types and does not involve screening. When $W$ decides to accept or reject, $W$ also doesn't know $\theta$, but since the offer is $0$ and $d_W = 0$, $W$ is indifferent and accepts by convention.

\textbf{When $W_j$ proposes} (prob $1/N$): offers $0$ to other $W$'s (who accept by the same indifference argument). $W_j$ must decide what to offer $H$. This is the screening subgame, which we now analyze via the normal form.

### R2 Screening Subgame: Normal Form

\textbf{Players}: $W$ (proposer, uninformed) and $H$ (responder, type $\theta \in \{0,1\}$).

\textbf{Strategy spaces}. $W$ chooses an offer $y_H$. By dominance, only two offers are relevant:
\begin{itemize}
\item \textbf{Aggressive (A)}: $y_H = \alpha = \alpha V(0)$
\item \textbf{Conservative (C)}: $y_H = \alpha r = \alpha V(1)$
\end{itemize}
Any $y_H < \alpha$ is rejected by both types; any $y_H \in (\alpha, \alpha r)$ is accepted by $\theta=0$ and rejected by $\theta=1$ (same as $A$ but costlier); any $y_H > \alpha r$ is accepted by both (same as $C$ but costlier).

$H$ of type $\theta$ chooses accept/reject for each possible offer. Writing $H$'s strategy as $(a_\theta^A, a_\theta^C) \in \{a, r\}^2$:

\textbf{Dominance elimination}:
\begin{itemize}
\item \textbf{Type $\theta=1$ facing $A$}: reject strictly dominates ($\alpha r > \alpha$). So $a_1^A = r$ always.
\item \textbf{Type $\theta=0$ facing $C$}: accept strictly dominates ($\alpha r > \alpha$). So $a_0^C = a$ always.
\end{itemize}

After elimination, the remaining choices are:
\begin{itemize}
\item Type $\theta=0$: $a_0^A \in \{a, r\}$ (accept or reject the aggressive offer)
\item Type $\theta=1$: $a_1^C \in \{a, r\}$ (accept or reject the conservative offer)
\end{itemize}

$H$'s reduced strategy set: $(a_0^A, a_1^C) \in \{a,r\}^2$, giving 4 strategies. $W$'s strategy set: $\{A, C\}$.

\textbf{Key observation}: $H$ is indifferent at the non-dominated margins:
\begin{itemize}
\item Type $\theta=0$ facing $A$: accept gives $\alpha$; reject gives $\alpha V(0) = \alpha$. \textbf{Indifferent}.
\item Type $\theta=1$ facing $C$: accept gives $\alpha r$; reject gives $\alpha V(1) = \alpha r$. \textbf{Indifferent}.
\end{itemize}

Therefore $H$'s payoff does not depend on $H$'s own strategy:
\begin{align}
u_H(\theta=0) &= \alpha \text{ if } W \text{ plays } A, \quad \alpha r \text{ if } W \text{ plays } C \\
u_H(\theta=1) &= \alpha r \text{ regardless of } W\text{'s action}
\end{align}

\textbf{$W$'s payoff matrix} (expected over $\theta$ given beliefs $\mu = \Pr(\theta=1)$):

\begin{center}
\begin{tabular}{c|cccc}
 & $(a,a)$ & $(a,r)$ & $(r,a)$ & $(r,r)$ \\
\hline
$A$ & $(1-\mu)(1-\alpha)$ & $(1-\mu)(1-\alpha)$ & $0$ & $0$ \\
$C$ & $V_e(\mu)-\alpha r$ & $(1-\mu)(1-\alpha r)$ & $V_e(\mu)-\alpha r$ & $(1-\mu)(1-\alpha r)$ \\
\end{tabular}
\end{center}

Columns are $H$'s strategies $(a_0^A, a_1^C)$. Note: $H$'s payoff is the same across all columns (indifferent).

\textbf{BNE analysis}. Since $H$ is indifferent, any strategy is a best response for $H$. The equilibrium is determined by $W$'s best response:
\begin{itemize}
\item Under $(a,a)$: $W$ compares $(1-\mu)(1-\alpha)$ vs $V_e(\mu) - \alpha r$. Cutoff at $\mu_s$.
\item Under $(a,r)$: $W$ compares $(1-\mu)(1-\alpha)$ vs $(1-\mu)(1-\alpha r)$. Since $1-\alpha > 1-\alpha r$, $A$ dominates.
\item Under $(r,a)$: $W$ compares $0$ vs $V_e(\mu) - \alpha r > 0$. $C$ dominates.
\item Under $(r,r)$: $W$ compares $0$ vs $(1-\mu)(1-\alpha r) > 0$. $C$ dominates.
\end{itemize}

Multiple BNE exist (because $H$ is indifferent). The standard convention in bargaining games is: \textbf{accept when indifferent}. This selects $(a,a)$: type $\theta=0$ accepts $A$, type $\theta=1$ accepts $C$.

Under $(a,a)$, $W$'s screening cutoff is:
\begin{equation}\label{eq:cutoff}
(1-\mu)(1-\alpha) = V_e(\mu) - \alpha r \quad \Longrightarrow \quad \mu_s = \frac{\alpha(r-1)}{r - \alpha}
\end{equation}

$W$ plays $A$ if $\mu < \mu_s$, $C$ if $\mu > \mu_s$.

### R2 Continuation Values (combining H-proposes and W-proposes)

Averaging over proposer identity (prob $1/N$ each):

$V_W^{R2}(\mu)$:
\begin{align}
\mu < \mu_s: &\quad \frac{(1-\mu)(1-\alpha)}{N} \\
\mu > \mu_s: &\quad \frac{V_e(\mu) - \alpha r}{N}
\end{align}

\textbf{Continuity check at $\mu_s$}: at $\mu_s = \alpha(r-1)/(r-\alpha)$, we have $1 - \mu_s = r(1-\alpha)/(r-\alpha)$, so $(1-\mu_s)(1-\alpha) = r(1-\alpha)^2/(r-\alpha)$. And $V_e(\mu_s) - \alpha r = 1 + \mu_s(r-1) - \alpha r = 1 + \alpha(r-1)^2/(r-\alpha) - \alpha r$. Simplifying: $= [r - \alpha + \alpha(r-1)^2 - \alpha r(r-\alpha)]/(r-\alpha) = [r - \alpha + \alpha r^2 - 2\alpha r + \alpha - \alpha r^2 + \alpha^2 r]/(r-\alpha) = [r - 2\alpha r + \alpha^2 r]/(r-\alpha) = r(1-\alpha)^2/(r-\alpha)$. Equal. $\checkmark$

$V_H^{R2}(\theta, \mu)$:
\begin{align}
\theta = 1: &\quad \frac{r[1+(N-1)\alpha]}{N} \quad \text{(constant in $\mu$)} \\
\theta = 0, \; \mu < \mu_s: &\quad \frac{1 + (N-1)\alpha}{N} \\
\theta = 0, \; \mu > \mu_s: &\quad \frac{1 + (N-1)\alpha r}{N} \quad \text{(overpaid)}
\end{align}

\textbf{Jump in $E_\theta[V_H^{R2}(\mu)]$ at cutoff $\mu_s$:}
\begin{equation}\label{eq:jump}
\text{Jump} = \frac{(N-1)\alpha(r-1)(1-\mu_s)}{N} > 0
\end{equation}

### Source of the difference between unanimity and majority

\begin{proposition}[Overpayment under unanimity]\label{prop:overpayment}
Under unanimity on the conservative branch ($\mu > \mu_s$), type $\theta=0$ receives strictly more than under majority:
\[
V_H^{R2}(\theta=0, U, \mu > \mu_s) = \frac{1+(N-1)\alpha r}{N} > \frac{1+(N-1)\alpha}{N} = V_H^{R2}(\theta=0, M).
\]
On the aggressive branch ($\mu < \mu_s$), payoffs are identical across rules. Type $\theta=1$ receives the same under both rules for all $\mu$.
\end{proposition}

\begin{proof}
Under majority, $W$ excludes $H$; $H$ receives the disagreement payoff $\alpha V(\theta)$, which is type-specific: $\alpha$ for $\theta=0$, $\alpha r$ for $\theta=1$. Under unanimity, $W$ must include $H$ but does not know $\theta$. On the conservative branch, $W$ offers $\alpha r$ to \emph{both} types. Type $\theta=1$ receives $\alpha r$ (same as majority). Type $\theta=0$ receives $\alpha r > \alpha$ --- overpaid relative to the type-specific disagreement payoff. This overpayment arises because $W$ cannot tailor the offer to the realized type and must offer enough for the high type.

On the aggressive branch, $W$ offers $\alpha$. Type $\theta=0$ accepts ($\alpha$), type $\theta=1$ rejects and receives $\alpha r$ from disagreement. These are the same type-specific payoffs as under majority. $\qed$
\end{proof}

\begin{remark}[Where the mechanism lives]\label{rem:mechanism}
The difference between unanimity and majority at R2 is \emph{not} in the voting rule per se. It is in the information structure of the offer. Under majority, $H$ is excluded and receives $\alpha V(\theta)$ --- a type-specific payoff determined by nature, not by any strategic interaction. Under unanimity, $W$ must make a single offer to $H$ without knowing $\theta$. When $W$ plays conservative, this single offer necessarily overpays the low type. The screening channel and the jump in $E[V_H]$ are consequences of this forced pooling in $W$'s offer.
\end{remark}

### Numerical Examples (R2 under Unanimity, $N=5$, $\mu = 0.5$)

\textbf{Example 1}: $r = 1.1$, $\alpha = 0.5$ (bilateral captures 50\% of multilateral value; $\alpha < 1/r = 0.909$ $\checkmark$).

Screening cutoff: $\mu_s = 0.5 \times 0.1 / (1.1 - 0.5) = 0.05/0.6 \approx 0.083$.

Since $\mu = 0.5 > \mu_s$, $W$ plays \textbf{conservative} (offers $y_H = \alpha r = 0.55$).

\textit{Payoff matrix for $W$ as proposer} (expected over $\theta$, given $\mu = 0.5$):

\begin{center}
\begin{tabular}{c|cccc}
 & $(a,a)$ & $(a,r)$ & $(r,a)$ & $(r,r)$ \\
\hline
$A$ & $0.25$ & $0.25$ & $0$ & $0$ \\
$C$ & $0.50$ & $0.225$ & $0.50$ & $0.225$ \\
\end{tabular}
\end{center}

Under $(a,a)$: $W$ prefers $C$ ($0.50 > 0.25$). $\checkmark$ Consistent with $\mu > \mu_s$.

\textit{Equilibrium payoffs and budget check}:

The expected pie is $V_e(0.5) = 1.05$.

$H$ expects to receive: $0.65$ (62\% of expected pie).
\begin{itemize}
\item If $\theta=0$: $H$ receives $[1 + 4 \times 0.55]/5 = 0.64$ (from a pie of 1.0)
\item If $\theta=1$: $H$ receives $[1.1 \times 3]/5 = 0.66$ (from a pie of 1.1)
\item Average: $0.5 \times 0.66 + 0.5 \times 0.64 = 0.65$
\end{itemize}

Each $W$ expects to receive: $0.10$ (9.5\% of expected pie).
\begin{itemize}
\item $W$ only gains when proposer (prob $1/5$).
\item As proposer, $W$ keeps $V(\theta) - \alpha r = V(\theta) - 0.55$:
  \begin{itemize}
  \item If $\theta=0$: $1 - 0.55 = 0.45$
  \item If $\theta=1$: $1.1 - 0.55 = 0.55$
  \end{itemize}
\item Expected as proposer: $0.5 \times 0.45 + 0.5 \times 0.55 = 0.50$
\item Times $1/5$: $0.10$
\end{itemize}

Budget check: $H(0.65) + 4 \times W(0.10) = 0.65 + 0.40 = 1.05 = V_e(0.5)$. $\checkmark$

Jump magnitude at $\mu_s$: $4 \times 0.5 \times 0.1 \times (1 - 0.083)/5 = 0.037$.

Interpretation: $r$ close to 1 means the states are similar, $\alpha$ high means $H$ has strong bilateral alternatives. The screening cutoff $\mu_s$ is very low (8.3\%): $W$ almost always plays conservative. The jump is small.

\bigskip

\textbf{Example 2}: $r = 1.5$, $\alpha = 0.3$ (bilateral captures 30\%; $\alpha < 1/r = 0.667$ $\checkmark$).

Screening cutoff: $\mu_s = 0.3 \times 0.5 / (1.5 - 0.3) = 0.15/1.2 = 0.125$.

Since $\mu = 0.5 > \mu_s$, $W$ plays \textbf{conservative} (offers $y_H = \alpha r = 0.45$).

\textit{Payoff matrix for $W$ as proposer} (expected over $\theta$, given $\mu = 0.5$):

\begin{center}
\begin{tabular}{c|cccc}
 & $(a,a)$ & $(a,r)$ & $(r,a)$ & $(r,r)$ \\
\hline
$A$ & $0.35$ & $0.35$ & $0$ & $0$ \\
$C$ & $0.80$ & $0.275$ & $0.80$ & $0.275$ \\
\end{tabular}
\end{center}

Under $(a,a)$: $W$ prefers $C$ ($0.80 > 0.35$). $\checkmark$

\textit{Equilibrium payoffs and budget check}:

The expected pie is $V_e(0.5) = 1.25$.

$H$ expects to receive: $0.61$ (49\% of expected pie).
\begin{itemize}
\item If $\theta=0$: $H$ receives $[1 + 4 \times 0.45]/5 = 0.56$ (from a pie of 1.0)
\item If $\theta=1$: $H$ receives $[1.5 \times 2.2]/5 = 0.66$ (from a pie of 1.5)
\item Average: $0.5 \times 0.66 + 0.5 \times 0.56 = 0.61$
\end{itemize}

Each $W$ expects to receive: $0.16$ (12.8\% of expected pie).
\begin{itemize}
\item As proposer (prob $1/5$), keeps $V(\theta) - 0.45$:
  \begin{itemize}
  \item If $\theta=0$: $1 - 0.45 = 0.55$
  \item If $\theta=1$: $1.5 - 0.45 = 1.05$
  \end{itemize}
\item Expected as proposer: $0.5 \times 0.55 + 0.5 \times 1.05 = 0.80$
\item Times $1/5$: $0.16$
\end{itemize}

Budget check: $H(0.61) + 4 \times W(0.16) = 0.61 + 0.64 = 1.25 = V_e(0.5)$. $\checkmark$

Jump magnitude at $\mu_s$: $4 \times 0.3 \times 0.5 \times (1 - 0.125)/5 = 0.105$.

\bigskip

\textbf{Comparison}: In Example 1 ($\alpha$ high, $r$ low), $H$ captures 62\% of the expected pie --- strong bilateral alternatives give bargaining power. In Example 2 ($\alpha$ low, $r$ high), $H$ captures 49\% --- weaker outside option. But the jump is larger in Example 2 ($0.105$ vs $0.037$): the screening channel is most powerful precisely when $H$ is ``weaker'' in terms of outside option.

\bigskip

\textbf{Scaling with $N$}: the screening cutoff $\mu_s = \alpha(r-1)/(r-\alpha)$ does not depend on $N$. The jump $(N-1)\alpha(r-1)(1-\mu_s)/N$ converges to $\alpha(r-1)(1-\mu_s)$ as $N \to \infty$. Individual payoffs scale as $1/N$. The table below shows how equilibrium shares evolve.

```{r scaling, echo=FALSE, results='asis'}
library(knitr)

compute_payoffs <- function(r, alpha, mu, N) {
  mu_s <- alpha * (r - 1) / (r - alpha)
  Ve <- 1 + mu * (r - 1)

  # Conservative branch (mu > mu_s for both examples at mu=0.5)
  VW <- (Ve - alpha * r) / N
  VH_0 <- (1 + (N - 1) * alpha * r) / N
  VH_1 <- (r * (1 + (N - 1) * alpha)) / N
  EVH <- mu * VH_1 + (1 - mu) * VH_0
  jump <- (N - 1) * alpha * (r - 1) * (1 - mu_s) / N
  H_share <- EVH / Ve
  W_share <- VW / Ve
  budget <- EVH + (N - 1) * VW

  data.frame(N = N, V_W = round(VW, 4), V_H_0 = round(VH_0, 4),
             V_H_1 = round(VH_1, 4), E_V_H = round(EVH, 4),
             H_pct = paste0(round(H_share * 100, 1), "\\%"),
             W_pct = paste0(round(W_share * 100, 1), "\\%"),
             Jump = round(jump, 4),
             Budget = round(budget, 4))
}

Ns <- c(5, 51, 101, 151, 201)

cat("\\textit{Example 1} ($r=1.1$, $\\alpha=0.5$, $\\mu=0.5$, $\\mu_s \\approx 0.083$):\n\n")

df1 <- do.call(rbind, lapply(Ns, function(n) compute_payoffs(1.1, 0.5, 0.5, n)))
kable(df1, format = "latex", booktabs = TRUE, escape = FALSE,
      col.names = c("$N$", "$V_W$", "$V_H(\\theta=0)$", "$V_H(\\theta=1)$",
                     "$E[V_H]$", "$H$ \\%", "$W$ \\%", "Jump", "Budget"),
      caption = NULL) |> cat()

cat("\n\n\\bigskip\n\n")
cat("\\textit{Example 2} ($r=1.5$, $\\alpha=0.3$, $\\mu=0.5$, $\\mu_s = 0.125$):\n\n")

df2 <- do.call(rbind, lapply(Ns, function(n) compute_payoffs(1.5, 0.3, 0.5, n)))
kable(df2, format = "latex", booktabs = TRUE, escape = FALSE,
      col.names = c("$N$", "$V_W$", "$V_H(\\theta=0)$", "$V_H(\\theta=1)$",
                     "$E[V_H]$", "$H$ \\%", "$W$ \\%", "Jump", "Budget"),
      caption = NULL) |> cat()
```

As $N$ grows, each player's individual payoff shrinks ($\sim 1/N$), but $H$'s \emph{share} of the pie converges to $\alpha$ (the bilateral fraction) and each $W$'s share vanishes. The jump converges to $\alpha(r-1)(1-\mu_s)$, a positive constant independent of $N$. The screening channel does not disappear with large $N$.

## Terminal Round (R2) under Majority

\textbf{When $H$ proposes} (prob $1/N$): buys $q-1$ votes (offers $0$, $d_W = 0$). Keeps $V(\theta)$.

\textbf{When $W$ proposes} (prob $(N-1)/N$): $W$ is indifferent between excluding $H$ (coalition gets $(1-\alpha)V(\theta)$, $H$ captures $\alpha V(\theta)$ bilaterally) and including $H$ at reservation $\alpha V(\theta)$ ($W$ keeps $(1-\alpha)V(\theta)$, $H$ gets $\alpha V(\theta)$ from the deal). Payoffs are identical. \textbf{Convention}: $W$ includes $H$ in the proposal (WLOG). This ensures the full pie $V(\theta)$ is divided among all $N$ players under both voting rules, simplifying comparison. $W$ offers $\alpha V(\theta)$ to $H$, offers $0$ to $q-2$ other $W$'s in the coalition, and keeps $(1-\alpha)V(\theta)$.

Note: under majority, the proposal passes regardless of $H$'s vote ($W$ has $q-1$ other votes). $H$ accepts the offer $\alpha V(\theta)$ (weakly: it equals $H$'s outside option). There is no screening because $H$'s acceptance has no consequence for whether the deal passes --- unlike unanimity, where $H$'s rejection kills the proposal.

\begin{align}
V_H^{R2}(\theta, M) &= \frac{V(\theta)[1 + (N-1)\alpha]}{N} \quad \text{--- linear in $\mu$, no screening.} \\
V_W^{R2}(\mu, M) &= \frac{(1-\alpha)V_e(\mu)}{N}
\end{align}

\textit{Verification of $V_H^{R2}(\theta, M)$}: $H$ proposes with prob $1/N$ and keeps $V(\theta)$. $W$ proposes with prob $(N-1)/N$ and offers $H$ exactly $\alpha V(\theta)$. So $V_H = (1/N)V(\theta) + ((N-1)/N)\alpha V(\theta) = V(\theta)[1+(N-1)\alpha]/N$. Taking expectations: $E_\theta[V_H^{R2}(\mu, M)] = V_e(\mu)[1+(N-1)\alpha]/N$, affine in $\mu$. $\checkmark$

\textit{Verification of $V_W^{R2}(\mu, M)$}: $W_i$ proposes with prob $1/N$ and keeps $(1-\alpha)V(\theta)$ (offers $\alpha V(\theta)$ to $H$, $0$ to other $W$'s). When $H$ proposes (prob $1/N$), $W_i$ receives $0$. When another $W_j$ proposes (prob $(N-2)/N$), $W_i$ receives $0$. So $V_W = (1/N)(1-\alpha)E[V(\theta)] = (1-\alpha)V_e(\mu)/N$. $\checkmark$

\textit{Budget check}: $E[V_H] + (N-1)V_W = V_e(\mu)[1+(N-1)\alpha]/N + (N-1)(1-\alpha)V_e(\mu)/N = V_e(\mu) \cdot N/N = V_e(\mu)$. $\checkmark$

### Numerical Examples and Comparison (R2, Majority vs Unanimity, $\mu = 0.5$)

Under majority there is no screening: $W$ excludes $H$, so $H$ always receives $\alpha V(\theta)$ when $W$ proposes (type-specific, no overpayment).

```{r majority_scaling, echo=FALSE, results='asis'}
library(knitr)

compute_majority <- function(r, alpha, mu, N) {
  Ve <- 1 + mu * (r - 1)
  VW <- (1 - alpha) * Ve / N
  VH_0 <- (1 + (N - 1) * alpha) / N  # V(0)[1+(N-1)alpha]/N
  VH_1 <- (r * (1 + (N - 1) * alpha)) / N
  EVH <- mu * VH_1 + (1 - mu) * VH_0
  H_share <- EVH / Ve
  W_share <- VW / Ve
  budget <- EVH + (N - 1) * VW
  data.frame(N = N, V_W = round(VW, 4), V_H_0 = round(VH_0, 4),
             V_H_1 = round(VH_1, 4), E_V_H = round(EVH, 4),
             H_pct = paste0(round(H_share * 100, 1), "\\%"),
             W_pct = paste0(round(W_share * 100, 1), "\\%"),
             Budget = round(budget, 4))
}

compute_unanimity <- function(r, alpha, mu, N) {
  mu_s <- alpha * (r - 1) / (r - alpha)
  Ve <- 1 + mu * (r - 1)
  VW <- (Ve - alpha * r) / N  # conservative branch
  VH_0 <- (1 + (N - 1) * alpha * r) / N
  VH_1 <- (r * (1 + (N - 1) * alpha)) / N
  EVH <- mu * VH_1 + (1 - mu) * VH_0
  H_share <- EVH / Ve
  W_share <- VW / Ve
  jump <- (N - 1) * alpha * (r - 1) * (1 - mu_s) / N
  budget <- EVH + (N - 1) * VW
  data.frame(N = N, V_W = round(VW, 4), V_H_0 = round(VH_0, 4),
             V_H_1 = round(VH_1, 4), E_V_H = round(EVH, 4),
             H_pct = paste0(round(H_share * 100, 1), "\\%"),
             W_pct = paste0(round(W_share * 100, 1), "\\%"),
             Jump = round(jump, 4),
             Budget = round(budget, 4))
}

Ns <- c(5, 51, 101, 151, 201)

# Example 1
cat("\\textbf{Example 1} ($r=1.1$, $\\alpha=0.5$, $\\mu=0.5$):\n\n")
cat("\\textit{Majority}:\n\n")
df1m <- do.call(rbind, lapply(Ns, function(n) compute_majority(1.1, 0.5, 0.5, n)))
kable(df1m, format = "latex", booktabs = TRUE, escape = FALSE,
      col.names = c("$N$", "$V_W$", "$V_H(\\theta\\!=\\!0)$", "$V_H(\\theta\\!=\\!1)$",
                     "$E[V_H]$", "$H$ \\%", "$W$ \\%", "Budget")) |> cat()

cat("\n\n\\textit{Unanimity (conservative branch)}:\n\n")
df1u <- do.call(rbind, lapply(Ns, function(n) compute_unanimity(1.1, 0.5, 0.5, n)))
kable(df1u, format = "latex", booktabs = TRUE, escape = FALSE,
      col.names = c("$N$", "$V_W$", "$V_H(\\theta\\!=\\!0)$", "$V_H(\\theta\\!=\\!1)$",
                     "$E[V_H]$", "$H$ \\%", "$W$ \\%", "Jump", "Budget")) |> cat()

cat("\n\n\\bigskip\n\n")

# Example 2
cat("\\textbf{Example 2} ($r=1.5$, $\\alpha=0.3$, $\\mu=0.5$):\n\n")
cat("\\textit{Majority}:\n\n")
df2m <- do.call(rbind, lapply(Ns, function(n) compute_majority(1.5, 0.3, 0.5, n)))
kable(df2m, format = "latex", booktabs = TRUE, escape = FALSE,
      col.names = c("$N$", "$V_W$", "$V_H(\\theta\\!=\\!0)$", "$V_H(\\theta\\!=\\!1)$",
                     "$E[V_H]$", "$H$ \\%", "$W$ \\%", "Budget")) |> cat()

cat("\n\n\\textit{Unanimity (conservative branch)}:\n\n")
df2u <- do.call(rbind, lapply(Ns, function(n) compute_unanimity(1.5, 0.3, 0.5, n)))
kable(df2u, format = "latex", booktabs = TRUE, escape = FALSE,
      col.names = c("$N$", "$V_W$", "$V_H(\\theta\\!=\\!0)$", "$V_H(\\theta\\!=\\!1)$",
                     "$E[V_H]$", "$H$ \\%", "$W$ \\%", "Jump", "Budget")) |> cat()
```

### H's gain from unanimity over majority

The gain $\Delta(\mu) = E_\theta[V_H^{R2}(U,\mu)] - E_\theta[V_H^{R2}(M,\mu)]$ has a simple closed form:
\[
\Delta(\mu) = \begin{cases} 0 & \mu \leq \mu_s \\ \displaystyle\frac{(1-\mu)(N-1)\alpha(r-1)}{N} & \mu > \mu_s \end{cases}
\]

On the aggressive branch, payoffs are identical across rules. On the conservative branch, the gain comes entirely from the overpayment of $\theta=0$ (Proposition \ref{prop:overpayment}).

\textit{Comparative statics} (for $\mu > \mu_s$):
\begin{itemize}
\item $\partial \Delta / \partial \mu = -(N-1)\alpha(r-1)/N < 0$: gain is \textbf{maximized at $\mu_s$} (just above the cutoff) and decreasing.
\item $\partial \Delta / \partial r = (1-\mu)(N-1)\alpha/N > 0$: greater asymmetry between states $\Rightarrow$ larger gain.
\item $\partial \Delta / \partial \alpha = (1-\mu)(N-1)(r-1)/N > 0$: stronger bilateral alternatives $\Rightarrow$ larger gain at a given $\mu$.
\item $\Delta(\mu_s) = \text{Jump} = (N-1)\alpha(r-1)(1-\mu_s)/N$: the maximum gain equals the jump magnitude.
\item As $N \to \infty$: $\Delta \to \alpha(r-1)(1-\mu)$, a positive constant independent of $N$.
\end{itemize}

\textit{Note}: $\partial \Delta / \partial \alpha > 0$ at a given $\mu$ does not mean the \emph{maximum} gain is increasing in $\alpha$. Higher $\alpha$ raises $\mu_s$ (shrinking the conservative region where the gain exists). The maximum gain $\Delta(\mu_s)$ depends on $\alpha$ through both the numerator and $(1-\mu_s)$.

```{r gain_table, echo=FALSE, results='asis'}
library(knitr)

compute_gain <- function(r, alpha, mu, N) {
  mu_s <- alpha * (r - 1) / (r - alpha)
  Ve <- 1 + mu * (r - 1)
  if (mu <= mu_s) {
    delta <- 0
  } else {
    delta <- (1 - mu) * (N - 1) * alpha * (r - 1) / N
  }
  delta_pct <- delta / Ve * 100
  # Max gain (at mu_s)
  max_gain <- (N - 1) * alpha * (r - 1) * (1 - mu_s) / N
  max_pct <- max_gain / (1 + mu_s * (r - 1)) * 100
  data.frame(mu = mu, Delta = round(delta, 4),
             Pct_pie = paste0(round(delta_pct, 1), "\\%"),
             Max_gain = round(max_gain, 4),
             Max_pct = paste0(round(max_pct, 1), "\\%"))
}

mus <- c(0.05, 0.1, 0.2, 0.3, 0.5, 0.7, 0.9)

cat("\\textit{Example 1} ($r=1.1$, $\\alpha=0.5$, $N=5$, $\\mu_s \\approx 0.083$):\n\n")
df1 <- do.call(rbind, lapply(mus, function(m) compute_gain(1.1, 0.5, m, 5)))
kable(df1, format = "latex", booktabs = TRUE, escape = FALSE,
      col.names = c("$\\mu$", "$\\Delta$", "\\% of pie",
                     "Max $\\Delta$", "Max \\%")) |> cat()

cat("\n\n\\bigskip\n\n")
cat("\\textit{Example 2} ($r=1.5$, $\\alpha=0.3$, $N=5$, $\\mu_s = 0.125$):\n\n")
df2 <- do.call(rbind, lapply(mus, function(m) compute_gain(1.5, 0.3, m, 5)))
kable(df2, format = "latex", booktabs = TRUE, escape = FALSE,
      col.names = c("$\\mu$", "$\\Delta$", "\\% of pie",
                     "Max $\\Delta$", "Max \\%")) |> cat()
```

The gain is modest in absolute terms but persistent: it does not vanish with $N$ and is largest at low $\mu$ (just above $\mu_s$), which is precisely the region where BP operates.

Key observations from the comparison:
\begin{itemize}
\item $V_H(\theta=1)$ is \textbf{identical} across rules (always $r[1+(N-1)\alpha]/N$).
\item $V_H(\theta=0)$ is \textbf{higher under unanimity} (conservative): $[1+(N-1)\alpha r]/N > [1+(N-1)\alpha]/N$. This is the overpayment (Proposition \ref{prop:overpayment}).
\item $V_W$ is \textbf{higher under majority}: $V_e(\mu)/N > [V_e(\mu) - \alpha r]/N$. $W$ keeps the whole pie under majority (excludes $H$) vs sharing with $H$ under unanimity.
\item As $N \to \infty$: $H$'s share converges to $\alpha$ under majority and to $\alpha r$ under unanimity (conservative). $W$'s share vanishes under both.
\item The jump under unanimity converges to $\alpha(r-1)(1-\mu_s) > 0$, independent of $N$.
\end{itemize}

## Round 1 (R1) under Unanimity

Let $x \equiv (N-1)\alpha r$ throughout this section.

\begin{remark}[Where screening lives]\label{rem:screening_R1}
On the equilibrium path, R2 is reached only when $W$ plays aggressive in R1 and $\theta=1$ rejects. In that case, $W$ knows $\theta=1$ with certainty ($\mu' = 1$). There is no screening in R2 on the equilibrium path. The screening of the model occurs in R1: $W$ proposer chooses between aggressive and conservative offers to $H$ without knowing $\theta$. R2 serves as the off-path threat whose continuation values determine R1 reservation values.
\end{remark}

### When $H$ proposes in R1 (prob $1/N$)

$H$ offers $\beta V_W^{R2}(\mu)$ to each $W$ (their discounted R2 continuation value). Both types of $H$ make the same offer (pooling: the offer depends on $\mu$, not on $\theta$, since $V_W^{R2}$ is a function of $\mu$ only). All $W$'s accept (indifferent). $H$ keeps $V(\theta) - (N-1)\beta V_W^{R2}(\mu)$.

No information about $\theta$ is revealed by $H$'s proposal (same offer for both types).

### When $W$ proposes in R1 (prob $(N-1)/N$)

$W$ offers 0 to other $W$'s (their reservation from R2 is $\beta V_W^{R2}(\mu) > 0$... but other $W$'s accept 0 only if 0 $\geq$ their continuation from rejection).

\textit{Note on other $W$'s offers}: $W_j$ must offer each $W_k$ at least $\beta V_W^{R2}(\mu')$ where $\mu'$ depends on rejection beliefs. On the equilibrium path where the proposal is accepted, the exact value matters only for budget accounting. We set $x_k = \beta V_W^{R2}(\mu)$ for each $W_k$ (using on-path beliefs). All $W_k$ accept.

$W_j$ must decide what to offer $H$. The two relevant offers are:

\textbf{Conservative offer}: $y_H^{con} = \beta(r + x)/N$

This is $\beta \times V_H^{R2}(\theta=1)$, the discounted R2 continuation of the more demanding type. Both types accept:
\begin{itemize}
\item $\theta=1$: $\beta(r+x)/N \geq \beta(r+x)/N$. Indifferent, accepts by convention.
\item $\theta=0$: $\beta(r+x)/N > \beta(1+x)/N$. Strictly prefers to accept.
\end{itemize}
Game ends in R1.

\textbf{Aggressive offer}: $y_H^{agg} = \beta(1 + x)/N$

This is $\beta \times V_H^{R2}(\theta=0, \mu'=1)$, the discounted R2 continuation of type $\theta=0$ under the belief $\mu'=1$ (which is what $W$ would believe after rejection, since only $\theta=1$ rejects on path).

\begin{itemize}
\item $\theta=0$: if $\theta=0$ accepts, gets $\beta(1+x)/N$. If $\theta=0$ deviates and rejects, $W$ believes $\theta=1$ ($\mu'=1$). In R2 with $\mu'=1$, $W$ offers $\alpha r$ to $H$ (knows $\theta=1$). Type $\theta=0$ accepts $\alpha r$ (since $\alpha r > \alpha$). Full R2 payoff for $\theta=0$: $(1+x)/N$. Discounted: $\beta(1+x)/N$. Exactly equal to the aggressive offer. \textbf{Indifferent, accepts by convention.}
\item $\theta=1$: $\beta(1+x)/N < \beta(r+x)/N$ (since $1 < r$). \textbf{Strictly rejects.} Game goes to R2 with $\mu'=1$ (complete information).
\end{itemize}

\textbf{Gap}: $\beta(r+x)/N - \beta(1+x)/N = \beta(r-1)/N > 0$. Screening is feasible.

### R2 after R1 rejection (complete information)

When $\theta=1$ rejects the aggressive offer in R1, the game goes to R2 with $W$ knowing $\theta=1$:
\begin{itemize}
\item $H$ proposes (prob $1/N$): keeps $V(1) = r$.
\item $W$ proposes (prob $(N-1)/N$): offers $\alpha r$ to $H$ (exact reservation under $\theta=1$). Keeps $r - \alpha r = r(1-\alpha)$.
\end{itemize}

R2 payoffs (under complete information, $\theta=1$):
\begin{align}
V_H^{R2}(\theta=1, \text{known}) &= \frac{r + x}{N} \\
V_W^{R2}(\theta=1, \text{known}) &= \frac{r(1-\alpha)}{N}
\end{align}

### $W$'s expected payoff from each strategy in R1

Let $\omega(\mu) \equiv (N-2)\beta V_W^{R2}(\mu)$ denote the total payment to other $W$'s in R1.

\textbf{Conservative} (both types accept, game ends in R1):
\begin{align}
F_1^{con}(\mu) &= E[V(\theta)] - \frac{\beta(r+x)}{N} - \omega(\mu) \nonumber \\
&= V_e(\mu) - \frac{\beta(r+x)}{N} - \omega(\mu) \label{eq:F1con}
\end{align}

\textbf{Aggressive} ($\theta=0$ accepts in R1; $\theta=1$ rejects, goes to R2):
\begin{align}
F_1^{agg}(\mu) &= (1-\mu)\left[V(0) - \frac{\beta(1+x)}{N} - \omega(\mu)\right] + \mu \cdot \frac{\beta r(1-\alpha)}{N} \label{eq:F1agg}
\end{align}

The first term: with prob $(1-\mu)$, $\theta=0$ accepts. $W$ keeps $V(0) - y_H^{agg} - \omega(\mu) = 1 - \beta(1+x)/N - \omega(\mu)$.

The second term: with prob $\mu$, $\theta=1$ rejects. Game goes to R2 with complete information. $W$'s expected R2 payoff (discounted): $\beta V_W^{R2}(\theta=1, \text{known}) = \beta r(1-\alpha)/N$.

### R1 screening cutoff

$W$ prefers aggressive when $F_1^{agg}(\mu) > F_1^{con}(\mu)$:
\begin{align}
&(1-\mu)\left[1 - \frac{\beta(1+x)}{N} - \omega\right] + \frac{\mu \beta r(1-\alpha)}{N} > V_e(\mu) - \frac{\beta(r+x)}{N} - \omega \nonumber
\end{align}

Expanding and simplifying (the $\omega(\mu)$ terms partially cancel):

Left side:
\[
(1-\mu) - \frac{(1-\mu)\beta(1+x)}{N} - (1-\mu)\omega + \frac{\mu\beta r(1-\alpha)}{N}
\]

Right side:
\[
V_e(\mu) - \frac{\beta(r+x)}{N} - \omega
\]

Difference (Left $-$ Right):
\begin{align}
&(1-\mu) - V_e(\mu) - \frac{(1-\mu)\beta(1+x)}{N} + \frac{\beta(r+x)}{N} + \frac{\mu\beta r(1-\alpha)}{N} + \mu\omega \nonumber
\end{align}

Note: $V_e(\mu) = 1 + \mu(r-1)$, so $(1-\mu) - V_e(\mu) = -\mu r$.

\begin{align}
&= -\mu r + \frac{\beta}{N}\left[-(1-\mu)(1+x) + (r+x) + \mu r(1-\alpha)\right] + \mu\omega \nonumber
\end{align}

The bracket:
\begin{align}
&-(1+x) + \mu(1+x) + r + x + \mu r(1-\alpha) \nonumber \\
&= r - 1 + \mu(1 + x + r - r\alpha) \nonumber \\
&= (r-1) + \mu(1 + r + x - r\alpha) \nonumber
\end{align}

Since $x = (N-1)\alpha r$, we have $x - r\alpha = (N-2)\alpha r$. So:
\[
= (r-1) + \mu(1 + r + (N-2)\alpha r)
\]

The full difference:
\[
\Delta_1(\mu) = -\mu r + \frac{\beta}{N}\left[(r-1) + \mu(1 + r + (N-2)\alpha r)\right] + \mu\omega(\mu)
\]

This depends on $\omega(\mu) = (N-2)\beta V_W^{R2}(\mu)$, which is piecewise (depends on which branch of $V_W^{R2}$ applies). The cutoff $\mu_s^{R1}$ solves $\Delta_1(\mu) = 0$.

Rather than solving in full generality (the algebra is involved because $V_W^{R2}$ is piecewise), we proceed numerically for the examples and note the closed-form structure.

### Continuation values in R1 under unanimity

Averaging over proposer identity:

$V_H^{R1}(\theta, \mu)$:
\begin{align}
&= \frac{1}{N}\left[V(\theta) - (N-1)\beta V_W^{R2}(\mu)\right] \quad \text{(H proposes)} \nonumber \\
&+ \frac{N-1}{N} \times \begin{cases}
\frac{\beta(1+x)}{N} & \text{if } \mu < \mu_s^{R1}, \; \theta=0 \text{ (aggressive, accepts)} \\
\frac{\beta(r+x)}{N} & \text{if } \mu > \mu_s^{R1} \text{ (conservative, both accept)} \\
\beta V_H^{R2}(\theta=1, \text{known}) = \frac{\beta(r+x)}{N} & \text{if } \mu < \mu_s^{R1}, \; \theta=1 \text{ (aggressive, rejects $\to$ R2)}
\end{cases}
\end{align}

Simplifying for each type:

\textbf{Type $\theta=1$}:
\begin{align}
V_H^{R1}(\theta=1, \mu) &= \frac{1}{N}\left[r - (N-1)\beta V_W^{R2}(\mu)\right] + \frac{(N-1)\beta(r+x)}{N^2}
\end{align}
This does not depend on whether $W$ plays aggressive or conservative (type $\theta=1$ gets $\beta(r+x)/N$ either way --- accepts conservative, rejects aggressive and gets R2 payoff $(r+x)/N$). Constant across screening branches.

\textbf{Type $\theta=0$, aggressive ($\mu < \mu_s^{R1}$)}:
\begin{align}
V_H^{R1}(\theta=0, \mu < \mu_s^{R1}) &= \frac{1}{N}\left[1 - (N-1)\beta V_W^{R2}(\mu)\right] + \frac{(N-1)\beta(1+x)}{N^2}
\end{align}

\textbf{Type $\theta=0$, conservative ($\mu > \mu_s^{R1}$)}:
\begin{align}
V_H^{R1}(\theta=0, \mu > \mu_s^{R1}) &= \frac{1}{N}\left[1 - (N-1)\beta V_W^{R2}(\mu)\right] + \frac{(N-1)\beta(r+x)}{N^2}
\end{align}

\textbf{Jump in $V_H^{R1}(\theta=0)$ at $\mu_s^{R1}$}:
\[
\text{Jump}_{\theta=0} = \frac{(N-1)\beta(r-1)}{N^2}
\]
This comes from $\theta=0$ receiving $\beta(r+x)/N$ (conservative) vs $\beta(1+x)/N$ (aggressive) when $W$ proposes, weighted by prob $(N-1)/N$ of $W$ proposing.

\textbf{Jump in $E_\theta[V_H^{R1}(\mu)]$ at $\mu_s^{R1}$}:
\[
\text{Jump} = (1-\mu_s^{R1}) \times \frac{(N-1)\beta(r-1)}{N^2}
\]
(Only $\theta=0$ is affected; $\theta=1$'s payoff is the same on both sides.)

### Numerical examples (R1 under Unanimity)

```{r R1_examples, echo=FALSE, results='asis'}
library(knitr)

R1_analysis <- function(r, alpha, mu, N, beta) {
  x <- (N - 1) * alpha * r
  mu_s_R2 <- alpha * (r - 1) / (r - alpha)
  Ve <- 1 + mu * (r - 1)

  # R2 continuation values (for the on-path beliefs)
  if (mu < mu_s_R2) {
    VW_R2 <- (1 - mu) * (1 - alpha) / N
  } else {
    VW_R2 <- (Ve - alpha * r) / N
  }

  # R2 under complete information (theta=1 known)
  VW_R2_known1 <- r * (1 - alpha) / N
  VH_R2_known1 <- (r + x) / N

  # Omega: payment to other W's in R1
  omega <- (N - 2) * beta * VW_R2

  # W's payoff from conservative in R1
  F1_con <- Ve - beta * (r + x) / N - omega

  # W's payoff from aggressive in R1
  F1_agg <- (1 - mu) * (1 - beta * (1 + x) / N - omega) + mu * beta * r * (1 - alpha) / N

  # Which strategy?
  strategy <- ifelse(F1_agg > F1_con, "Aggressive", "Conservative")

  # H's R1 payoffs (when H proposes + when W proposes)
  H_proposes_0 <- (1 - (N - 1) * beta * VW_R2) / N  # theta=0, H proposes: keeps V(0)=1
  H_proposes_1 <- (r - (N - 1) * beta * VW_R2) / N   # theta=1, H proposes: keeps V(1)=r

  if (F1_agg > F1_con) {
    # Aggressive: theta=0 gets beta(1+x)/N from W, theta=1 gets beta(r+x)/N from R2
    W_offer_0 <- beta * (1 + x) / N
    W_offer_1 <- beta * (r + x) / N  # R2 continuation
    VH1_0 <- H_proposes_0 + (N - 1) * W_offer_0 / N
    VH1_1 <- H_proposes_1 + (N - 1) * W_offer_1 / N
    # W's payoff
    VW1 <- F1_agg / N  # prob 1/N of being proposer
  } else {
    # Conservative: both get beta(r+x)/N from W
    W_offer <- beta * (r + x) / N
    VH1_0 <- H_proposes_0 + (N - 1) * W_offer / N
    VH1_1 <- H_proposes_1 + (N - 1) * W_offer / N
    VW1 <- F1_con / N
  }

  EVH1 <- mu * VH1_1 + (1 - mu) * VH1_0
  jump_R1 <- (1 - mu) * (N - 1) * beta * (r - 1) / N^2  # at cutoff
  budget <- EVH1 + (N - 1) * VW1

  data.frame(
    mu = mu,
    Strategy = strategy,
    F1_agg = round(F1_agg, 4),
    F1_con = round(F1_con, 4),
    VH1_0 = round(VH1_0, 4),
    VH1_1 = round(VH1_1, 4),
    EVH1 = round(EVH1, 4),
    VW1 = round(VW1, 4),
    Jump = round(jump_R1, 4),
    Budget = round(budget, 4)
  )
}

mus <- c(0.05, 0.1, 0.2, 0.3, 0.5, 0.7, 0.9)
beta_val <- 0.9

cat(paste0("\\textbf{Example 1}: $r=1.1$, $\\alpha=0.5$, $N=5$, $\\beta=", beta_val, "$\n\n"))
cat(paste0("$\\mu_s^{R2} = ", round(0.5*0.1/(1.1-0.5), 3), "$, $x = (N-1)\\alpha r = ", 4*0.5*1.1, "$\n\n"))

df1 <- do.call(rbind, lapply(mus, function(m) R1_analysis(1.1, 0.5, m, 5, beta_val)))
kable(df1, format = "latex", booktabs = TRUE, escape = FALSE,
      col.names = c("$\\mu$", "Strat.", "$F_1^{agg}$", "$F_1^{con}$",
                     "$V_H^{R1}(0)$", "$V_H^{R1}(1)$", "$E[V_H^{R1}]$",
                     "$V_W^{R1}$", "Jump", "Budget"),
      caption = NULL) |> cat()

cat("\n\n\\bigskip\n\n")

cat(paste0("\\textbf{Example 2}: $r=1.5$, $\\alpha=0.3$, $N=5$, $\\beta=", beta_val, "$\n\n"))
cat(paste0("$\\mu_s^{R2} = ", round(0.3*0.5/(1.5-0.3), 3), "$, $x = (N-1)\\alpha r = ", 4*0.3*1.5, "$\n\n"))

df2 <- do.call(rbind, lapply(mus, function(m) R1_analysis(1.5, 0.3, m, 5, beta_val)))
kable(df2, format = "latex", booktabs = TRUE, escape = FALSE,
      col.names = c("$\\mu$", "Strat.", "$F_1^{agg}$", "$F_1^{con}$",
                     "$V_H^{R1}(0)$", "$V_H^{R1}(1)$", "$E[V_H^{R1}]$",
                     "$V_W^{R1}$", "Jump", "Budget"),
      caption = NULL) |> cat()
```

## Round 1 (R1) under Majority

Under majority, $W$ includes $H$ at reservation $\alpha V(\theta)$ (WLOG convention). The proposal passes regardless of $H$'s vote. No screening.

\textbf{When $H$ proposes (prob $1/N$)}:
$H$ offers $\beta V_W^{R2}(\mu, M) = \beta(1-\alpha)V_e(\mu)/N$ to each $W$ in the coalition. $W$ accepts. $H$ keeps $V(\theta) - (q-1)\beta(1-\alpha)V_e(\mu)/N$.

\textbf{When $W$ proposes (prob $(N-1)/N$)}:
$W$ offers $\alpha V(\theta)$ to $H$ (reservation, type-specific). But $W$ doesn't know $\theta$. Under majority, this doesn't matter: the proposal passes regardless. $W$ offers $\beta\alpha V_e(\mu)$ (expected reservation) to $H$ and keeps the rest. Actually, since the proposal passes by majority without $H$'s vote, $W$ can offer any amount. By the WLOG convention (include $H$), $W$ offers $\alpha V(\theta)$. Since $W$ doesn't know $V(\theta)$, $W$ offers $\beta \cdot \alpha V_e(\mu)$ and $H$ gets $\alpha V(\theta)$ regardless (either from the deal or outside option).

In practice, the R1 payoffs under majority are just discounted R2 values:
\begin{align}
V_H^{R1}(\theta, M) &= \frac{V(\theta)[1 + (N-1)\alpha \cdot \phi(\beta)]}{N} \\
V_W^{R1}(\mu, M) &= \frac{(1-\alpha)V_e(\mu) \cdot \psi(\beta)}{N}
\end{align}
where $\phi$ and $\psi$ are functions of $\beta$ from the standard BF backward induction. The key point: everything remains \textbf{linear in $\mu$, no screening, no jump}.

A full derivation of R1 under majority parallels the unanimity case but without the aggressive/conservative distinction. Since there is no screening, the continuation values are affine in $\mu$ and do not contribute to BP. We defer the exact expressions to focus on the unanimity mechanism.

## Entry

A weak state enters if its expected payoff from participating exceeds the entry cost:
\[
V_W^{R1}(\mu, R) \geq c
\]

The entry threshold $\tau(R)$ is the value of $\mu$ at which $W$ is indifferent.

### Entry under unanimity

$V_W^{R1}(\mu, U)$ is piecewise: it depends on whether $W$ plays aggressive or conservative in R1 (cutoff $\mu_s^{R1}$), and within the aggressive branch, on $V_W^{R2}(\mu)$ (cutoff $\mu_s$). So $V_W^{R1}$ can have up to three regimes.

The entry threshold $\tau(U)$ solves $V_W^{R1}(\tau, U) = c$. Since $V_W^{R1}$ is not monotone in general (it depends on the piecewise structure), the threshold may not be unique. We characterize it numerically for the examples.

```{r entry_unanimity, echo=FALSE, results='asis'}
library(knitr)

VW_R1_unanimity <- function(r, alpha, mu, N, beta) {
  x <- (N - 1) * alpha * r
  mu_s_R2 <- alpha * (r - 1) / (r - alpha)
  Ve <- 1 + mu * (r - 1)

  # V_W^R2(mu)
  if (mu < mu_s_R2) {
    VW_R2 <- (1 - mu) * (1 - alpha) / N
  } else {
    VW_R2 <- (Ve - alpha * r) / N
  }

  omega <- (N - 2) * beta * VW_R2

  # F1 aggressive and conservative
  F1_con <- Ve - beta * (r + x) / N - omega
  F1_agg <- (1 - mu) * (1 - beta * (1 + x) / N - omega) + mu * beta * r * (1 - alpha) / N

  # W's R1 payoff = (1/N) * max(F1_agg, F1_con) + 0 (non-proposer gets beta*VW_R2 when H proposes, 0 when other W proposes... )

  # Actually V_W^R1 = (1/N) * F_proposer + ((N-1)/N) * reservation_from_R1
  # When H proposes: W gets beta*VW_R2 (offered by H)
  # When other W proposes: W gets beta*VW_R2 (offered by other W)
  # When W_i proposes: W gets F (proposer surplus)

  # Wait, I need to be more careful.
  # V_W^R1 = (1/N)*F_proposer + (1/N)*beta*VW_R2 + ((N-2)/N)*beta*VW_R2
  # = (1/N)*F_proposer + ((N-1)/N)*beta*VW_R2

  F_proposer <- max(F1_agg, F1_con)
  VW_R1 <- F_proposer / N + (N - 1) * beta * VW_R2 / N

  return(VW_R1)
}

VW_R1_majority <- function(r, alpha, mu, N, beta) {
  Ve <- 1 + mu * (r - 1)
  # R2 majority: VW_R2 = (1-alpha)*Ve/N
  VW_R2_M <- (1 - alpha) * Ve / N

  # R1 majority: W proposes and keeps (1-alpha)*V(theta) - offers to other W's
  # When W proposes: keeps (1-alpha)*V(theta) - (N-2)*beta*VW_R2_M... wait

  # Under majority, same BF logic:
  # W proposes (1/N): offers alpha*V(theta) to H, beta*VW_R2_M to other W's, keeps residual
  # H proposes (1/N): offers beta*VW_R2_M to W, keeps rest
  # Other W proposes ((N-2)/N): W gets beta*VW_R2_M

  # F_proposer for W under majority:
  omega_M <- (N - 2) * beta * VW_R2_M
  F1_M <- Ve * (1 - alpha) - omega_M  # expected residual after paying H and other W's
  # Wait, that's not right. W offers alpha*V(theta) to H and beta*VW_R2_M to each other W
  # F = V(theta) - alpha*V(theta) - (N-2)*beta*VW_R2_M = (1-alpha)*V(theta) - omega_M
  # Expected: (1-alpha)*Ve - omega_M

  F_proposer_M <- (1 - alpha) * Ve - omega_M
  VW_R1_M <- F_proposer_M / N + (N - 1) * beta * VW_R2_M / N

  return(VW_R1_M)
}

mus <- seq(0.01, 0.99, by = 0.01)
beta_val <- 0.9

# Example 1
vw_u1 <- sapply(mus, function(m) VW_R1_unanimity(1.1, 0.5, m, 5, beta_val))
vw_m1 <- sapply(mus, function(m) VW_R1_majority(1.1, 0.5, m, 5, beta_val))

# Example 2
vw_u2 <- sapply(mus, function(m) VW_R1_unanimity(1.5, 0.3, m, 5, beta_val))
vw_m2 <- sapply(mus, function(m) VW_R1_majority(1.5, 0.3, m, 5, beta_val))

# Entry thresholds for c = 0.1
c_val <- 0.1

# Find thresholds (first crossing from below/above)
find_threshold <- function(vw, mus, c) {
  above <- vw >= c
  if (all(above)) return(0)
  if (all(!above)) return(1)
  # Find first mu where VW >= c
  idx <- which(above)[1]
  if (is.na(idx)) return(1)
  return(mus[idx])
}

tau_U1 <- find_threshold(vw_u1, mus, c_val)
tau_M1 <- find_threshold(vw_m1, mus, c_val)
tau_U2 <- find_threshold(vw_u2, mus, c_val)
tau_M2 <- find_threshold(vw_m2, mus, c_val)

cat(paste0("With $c = ", c_val, "$, $\\beta = ", beta_val, "$:\n\n"))

cat("\\begin{center}\n")
cat("\\begin{tabular}{lcccc}\n")
cat(" & $\\tau(U)$ & $\\tau(M)$ & $V_W^{R1}(0.5, U)$ & $V_W^{R1}(0.5, M)$ \\\\\n")
cat("\\hline\n")
cat(paste0("Example 1 ($r=1.1, \\alpha=0.5$) & ",
           round(tau_U1, 3), " & ", round(tau_M1, 3), " & ",
           round(VW_R1_unanimity(1.1, 0.5, 0.5, 5, beta_val), 4), " & ",
           round(VW_R1_majority(1.1, 0.5, 0.5, 5, beta_val), 4), " \\\\\n"))
cat(paste0("Example 2 ($r=1.5, \\alpha=0.3$) & ",
           round(tau_U2, 3), " & ", round(tau_M2, 3), " & ",
           round(VW_R1_unanimity(1.5, 0.3, 0.5, 5, beta_val), 4), " & ",
           round(VW_R1_majority(1.5, 0.3, 0.5, 5, beta_val), 4), " \\\\\n"))
cat("\\end{tabular}\n")
cat("\\end{center}\n\n")
```

### Entry under majority

Under majority, $V_W^{R1}(\mu, M)$ is affine in $\mu$ (no screening, no piecewise structure). The entry threshold $\tau(M)$ solves $V_W^{R1}(\tau, M) = c$, which has a unique solution.

### Entry comparison

```{r entry_plot, echo=FALSE, fig.width=8, fig.height=4, fig.cap="$V_W^{R1}(\\mu)$ under unanimity and majority for both examples. Horizontal line: entry cost $c = 0.1$. Entry occurs where the curve exceeds $c$."}
par(mfrow = c(1, 2), mar = c(4, 4, 2, 1))

# Example 1
plot(mus, vw_u1, type = "l", col = "blue", lwd = 2,
     xlab = expression(mu), ylab = expression(V[W]^{R1}),
     main = expression(paste("Example 1: r=1.1, ", alpha, "=0.5")),
     ylim = range(c(vw_u1, vw_m1, c_val)))
lines(mus, vw_m1, col = "red", lwd = 2)
abline(h = c_val, lty = 2, col = "gray40")
legend("topright", c("Unanimity", "Majority", paste0("c = ", c_val)),
       col = c("blue", "red", "gray40"), lwd = c(2, 2, 1), lty = c(1, 1, 2), cex = 0.7)

# Example 2
plot(mus, vw_u2, type = "l", col = "blue", lwd = 2,
     xlab = expression(mu), ylab = expression(V[W]^{R1}),
     main = expression(paste("Example 2: r=1.5, ", alpha, "=0.3")),
     ylim = range(c(vw_u2, vw_m2, c_val)))
lines(mus, vw_m2, col = "red", lwd = 2)
abline(h = c_val, lty = 2, col = "gray40")
legend("topright", c("Unanimity", "Majority", paste0("c = ", c_val)),
       col = c("blue", "red", "gray40"), lwd = c(2, 2, 1), lty = c(1, 1, 2), cex = 0.7)
```

## Bayesian Persuasion

$H$ chooses the institutional rule $R$ at Stage 0 and the signal structure $\pi$ at Stage 1 to maximize expected payoff. By the standard Bayesian Persuasion result (Kamenica and Gentzkow 2011), $H$'s optimal payoff under rule $R$ equals the concavification of $H$'s value function evaluated at the prior $p$:
\[
\Pi_H^*(R) = \operatorname{cav} v(p, R)
\]

### $H$'s value function

$H$'s value function maps posteriors $\mu$ to $H$'s expected payoff:
\[
v(\mu, R) = \begin{cases}
0 & \text{if } V_W^{R1}(\mu, R) < c \quad \text{(no entry)} \\
E_\theta[V_H^{R1}(\theta, \mu, R)] & \text{if } V_W^{R1}(\mu, R) \geq c \quad \text{(entry)}
\end{cases}
\]

When $W$ does not enter, $H$ gets 0 (no institution). When $W$ enters, $H$ gets the expected R1 payoff.

\textbf{Under majority}: $v(\mu, M) = \mathbf{1}\{V_W^{R1}(\mu, M) \geq c\} \cdot E_\theta[V_H^{R1}(\theta, \mu, M)]$. Both the indicator and the continuation value are affine (or piecewise affine) in $\mu$. The value function has a jump at $\tau(M)$ (entry threshold) but is linear above it. BP exploits only the entry threshold.

\textbf{Under unanimity}: $v(\mu, U)$ has TWO sources of non-convexity:
\begin{enumerate}
\item \textbf{Entry threshold} at $\tau(U)$: $v$ jumps from 0 to positive.
\item \textbf{Screening cutoff} at $\mu_s^{R1}$: $E_\theta[V_H^{R1}]$ has an upward jump (from the overpayment of $\theta=0$ when $W$ switches from aggressive to conservative).
\end{enumerate}

BP can exploit both non-convexities. This is the screening channel that is absent under majority.

### Analytical concavification

\textbf{Structure of $v(\mu, U)$}. The value function under unanimity is:
\[
v(\mu, U) = \begin{cases}
0 & \mu < \tau(U) \\
v_L(\mu) & \tau(U) \leq \mu < \mu_s^{R1} \\
v_H(\mu) & \mu \geq \mu_s^{R1}
\end{cases}
\]
where $v_L(\mu) = E_\theta[V_H^{R1}(\theta, \mu)]$ on the aggressive branch (pre-screening) and $v_H(\mu) = E_\theta[V_H^{R1}(\theta, \mu)]$ on the conservative branch (post-screening). There is an upward jump at $\mu_s^{R1}$:
\[
v_H(\mu_s^{R1}) - v_L(\mu_s^{R1}) = (1 - \mu_s^{R1}) \cdot \frac{(N-1)\beta(r-1)}{N^2} > 0.
\]

Both $v_L$ and $v_H$ depend on $V_W^{R2}(\mu)$ and are continuous on their respective domains.

\textbf{Structure of $v(\mu, M)$}. Under majority:
\[
v(\mu, M) = \begin{cases}
0 & \mu < \tau(M) \\
E_\theta[V_H^{R1}(\theta, \mu, M)] & \mu \geq \tau(M)
\end{cases}
\]
This is affine in $\mu$ for $\mu \geq \tau(M)$. No jump, no non-convexity beyond the entry threshold.

\bigskip
\textbf{Case 1: $\tau(U) > 0$, prior $p < \tau(U)$ (entry does not occur at the prior).}

The concavification from the origin creates a line with slope:
\[
S(R) = \sup_{\mu \in [\tau(R), 1]} \frac{v(\mu, R)}{\mu}
\]
and $\operatorname{cav} v(p, R) = S(R) \cdot p$ for $p < \tau(R)$.

Under majority: $v(\mu, M) = A_M + B_M \mu$ is affine for $\mu \geq \tau(M)$, so $v/\mu$ is decreasing. The supremum is at $\mu = \tau(M)$:
\[
S_M = \frac{v(\tau(M), M)}{\tau(M)}
\]

Under unanimity: $v/\mu$ can attain a higher value at the jump. If $v_H(\mu_s^{R1})/\mu_s^{R1} > v_L(\mu)/\mu$ for all $\mu \in [\tau(U), \mu_s^{R1})$, the supremum is at $\mu_s^{R1}$ (from the right):
\[
S_U = \frac{v_H(\mu_s^{R1})}{\mu_s^{R1}}
\]

\begin{proposition}[BP dominance at low priors]\label{prop:bp_low}
If $S_U > S_M$, then $H$ strictly prefers unanimity for all priors $p \in (0, \min(\tau(U), \tau(M)))$:
\[
\operatorname{cav} v(p, U) = S_U \cdot p > S_M \cdot p = \operatorname{cav} v(p, M).
\]
The condition $S_U > S_M$ is equivalent to:
\[
\frac{v_H(\mu_s^{R1})}{\mu_s^{R1}} > \frac{v(\tau(M), M)}{\tau(M)}.
\]
The left side is the ``return per unit of belief'' at the screening cutoff under unanimity. The right side is the return at the entry threshold under majority. Unanimity dominates when the screening jump creates a sufficiently high return relative to majority's entry-threshold return.
\end{proposition}

\textbf{Case 2: $\tau(U) \approx 0$, prior $p > \tau(U)$ (entry occurs at the prior).}

In this case, $v(p, U) > 0$. The BP gain comes from exploiting the screening jump at $\mu_s^{R1}$. The concavification bridges the non-convexity: for priors near $\mu_s^{R1}$, $H$ creates posteriors straddling the jump.

Formally, there exist $\mu_1 < \mu_s^{R1} < \mu_2$ such that the line connecting $({\mu_1, v_L(\mu_1)})$ to $({\mu_2, v_H(\mu_2)})$ is tangent to $v_L$ at $\mu_1$ and tangent to $v_H$ at $\mu_2$. For $p \in (\mu_1, \mu_2)$:
\[
\operatorname{cav} v(p, U) = v_L(\mu_1) + \frac{v_H(\mu_2) - v_L(\mu_1)}{\mu_2 - \mu_1}(p - \mu_1) > v(p, U).
\]

The optimal signal creates two posteriors: $\mu_1$ (with prob $(\mu_2 - p)/(\mu_2 - \mu_1)$) and $\mu_2$ (with prob $(p - \mu_1)/(\mu_2 - \mu_1)$). At $\mu_1$, $W$ plays aggressive. At $\mu_2$, $W$ plays conservative. $H$ benefits from the switch.

Under majority, $v(\mu, M)$ is concave for $\mu \geq \tau(M)$, so $\operatorname{cav} v(p, M) = v(p, M)$. No BP gain beyond entry.

\begin{proposition}[BP gain from screening]\label{prop:bp_screening}
Under unanimity, $\operatorname{cav} v(p, U) > v(p, U)$ for all $p \in (\mu_1, \mu_2)$, with BP gain:
\[
\operatorname{cav} v(p, U) - v(p, U) > 0.
\]
Under majority, $\operatorname{cav} v(p, M) = v(p, M)$ for all $p \geq \tau(M)$: no BP gain beyond entry.

The screening channel is therefore the unique source of differential BP advantage for unanimity over majority.
\end{proposition}

\textbf{Summary of the two channels}:
\begin{enumerate}
\item \textbf{Entry channel}: if $\tau(U) < \tau(M)$, unanimity has a lower entry threshold, giving BP a longer ``reach'' at low priors. This channel exists regardless of screening.
\item \textbf{Screening channel}: the jump at $\mu_s^{R1}$ creates a non-convexity in $v(\mu, U)$ that BP exploits. This channel is absent under majority (no jump, $v$ is concave). This is the paper's distinctive contribution.
\end{enumerate}

### Numerical illustration

```{r bp_plot, echo=FALSE, fig.width=8, fig.height=5, fig.cap="$H$'s value function $v(\\mu, R)$ and its concavification under both rules. Dashed: concave envelope. The screening jump under unanimity creates additional concavification gain."}

VH_R1_unanimity <- function(r, alpha, mu, N, beta) {
  x <- (N - 1) * alpha * r
  mu_s_R2 <- alpha * (r - 1) / (r - alpha)
  Ve <- 1 + mu * (r - 1)

  if (mu < mu_s_R2) {
    VW_R2 <- (1 - mu) * (1 - alpha) / N
  } else {
    VW_R2 <- (Ve - alpha * r) / N
  }

  omega <- (N - 2) * beta * VW_R2

  F1_con <- Ve - beta * (r + x) / N - omega
  F1_agg <- (1 - mu) * (1 - beta * (1 + x) / N - omega) + mu * beta * r * (1 - alpha) / N

  # H proposes: keeps V(theta) - (N-1)*beta*VW_R2
  H_prop_0 <- (1 - (N - 1) * beta * VW_R2) / N
  H_prop_1 <- (r - (N - 1) * beta * VW_R2) / N

  if (F1_agg > F1_con) {
    # Aggressive
    VH_0 <- H_prop_0 + (N - 1) * beta * (1 + x) / (N^2)
    VH_1 <- H_prop_1 + (N - 1) * beta * (r + x) / (N^2)
  } else {
    # Conservative
    VH_0 <- H_prop_0 + (N - 1) * beta * (r + x) / (N^2)
    VH_1 <- H_prop_1 + (N - 1) * beta * (r + x) / (N^2)
  }

  EVH <- mu * VH_1 + (1 - mu) * VH_0
  return(EVH)
}

VH_R1_majority <- function(r, alpha, mu, N, beta) {
  Ve <- 1 + mu * (r - 1)
  VW_R2_M <- (1 - alpha) * Ve / N
  omega_M <- (N - 2) * beta * VW_R2_M

  # H proposes: keeps V(theta) - (q-1)*beta*VW_R2_M ... simplify: same structure
  # Actually for majority, H offers beta*VW_R2_M to (q-1) W's
  # But for expected payoff, we average over proposer:
  # H proposes (1/N): keeps V(theta) - (N-1)*beta*VW_R2_M (offers to all W's... wait, majority only needs q-1)
  # Let's use the simpler: V_H^R1(M) follows standard BF discounting
  # V_H^R2(theta, M) = V(theta)[1+(N-1)alpha]/N
  # V_H^R1(theta, M) = V(theta) - (N-1)*beta*VW_R2_M ... when H proposes
  # + (N-1)/N * beta*alpha*V(theta) ... when W proposes, H gets alpha*V(theta)
  # Hmm, this needs care. Let me use the R2 approach:
  # R1 majority: same structure as R2 but with beta*R2 as continuation
  # When H proposes: offers beta*VW_R2_M to (N-1) W's, keeps V(theta) - (N-1)*beta*VW_R2_M
  # When W proposes: offers beta*V_H^R2(theta,M) to H... but doesn't know theta.
  # Under majority: proposal passes without H's vote. W offers beta*alpha*V_e(mu) or just alpha*V(theta).
  # Since no screening, W effectively pays alpha*V(theta) to H. Expected: alpha*Ve.
  # W keeps (1-alpha)*V(theta) - (N-2)*beta*VW_R2_M

  # H's payoff:
  # H proposes (1/N): V(theta) - (N-1)*beta*VW_R2_M
  # W proposes ((N-1)/N): beta * V_H^R2(theta, M) = beta * V(theta)[1+(N-1)*alpha]/N
  # Wait, H's reservation from R2 is beta * V_H^R2. W offers this amount.

  VH_R2_theta <- function(theta_val) {
    V_theta <- ifelse(theta_val == 0, 1, r)
    V_theta * (1 + (N - 1) * alpha) / N
  }

  # H proposes in R1: keeps V(theta) - (N-1)*beta*VW_R2_M
  # W proposes in R1: H gets beta*VH_R2(theta)
  VH1_0 <- (1 - (N-1)*beta*VW_R2_M)/N + (N-1)/N * beta * VH_R2_theta(0)
  VH1_1 <- (r - (N-1)*beta*VW_R2_M)/N + (N-1)/N * beta * VH_R2_theta(1)

  EVH <- mu * VH1_1 + (1 - mu) * VH1_0
  return(EVH)
}

# Value functions v(mu, R) = indicator(entry) * E[VH]
c_val <- 0.1
beta_val <- 0.9

v_unanimity <- function(r, alpha, mu, N, beta, c) {
  vw <- VW_R1_unanimity(r, alpha, mu, N, beta)
  if (vw >= c) {
    return(VH_R1_unanimity(r, alpha, mu, N, beta))
  } else {
    return(0)
  }
}

v_majority <- function(r, alpha, mu, N, beta, c) {
  vw <- VW_R1_majority(r, alpha, mu, N, beta)
  if (vw >= c) {
    return(VH_R1_majority(r, alpha, mu, N, beta))
  } else {
    return(0)
  }
}

mus_fine <- seq(0.001, 0.999, by = 0.001)

# Compute concave envelope (simple algorithm)
concavify <- function(mus, vals) {
  n <- length(mus)
  cav <- vals
  # Upper concave envelope: for each point, check if it's below the line
  # connecting any two other points
  changed <- TRUE
  while (changed) {
    changed <- FALSE
    for (i in 2:(n-1)) {
      # Linear interpolation between left and right
      for (j in 1:(i-1)) {
        for (k in (i+1):n) {
          interp <- vals[j] + (vals[k] - vals[j]) * (mus[i] - mus[j]) / (mus[k] - mus[j])
          if (interp > cav[i] + 1e-10) {
            cav[i] <- interp
            changed <- TRUE
          }
        }
      }
    }
  }
  # Simpler approach: use convex hull
  # Points: (mu, v). Concave envelope is upper boundary of convex hull
  pts <- cbind(mus, vals)
  # Add origin and endpoint
  hull_pts <- rbind(c(0, 0), pts, c(1, vals[n]))
  h <- chull(hull_pts[,1], hull_pts[,2])
  # Upper hull: points where y is maximal
  hull_x <- hull_pts[h, 1]
  hull_y <- hull_pts[h, 2]
  # Sort by x
  ord <- order(hull_x)
  hull_x <- hull_x[ord]
  hull_y <- hull_y[ord]
  # Keep only upper hull (above the line from first to last)
  # Use approx to interpolate
  cav_vals <- approx(hull_x, hull_y, xout = mus, rule = 2)$y
  # Ensure cav >= v
  cav_vals <- pmax(cav_vals, vals)
  return(cav_vals)
}

par(mfrow = c(1, 2), mar = c(4, 4, 2, 1))

for (ex in 1:2) {
  if (ex == 1) { r <- 1.1; alpha <- 0.5; ttl <- "Example 1: r=1.1, a=0.5" }
  else { r <- 1.5; alpha <- 0.3; ttl <- "Example 2: r=1.5, a=0.3" }

  v_U <- sapply(mus_fine, function(m) v_unanimity(r, alpha, m, 5, beta_val, c_val))
  v_M <- sapply(mus_fine, function(m) v_majority(r, alpha, m, 5, beta_val, c_val))

  cav_U <- concavify(mus_fine, v_U)
  cav_M <- concavify(mus_fine, v_M)

  ylim <- range(c(v_U, v_M, cav_U, cav_M))

  plot(mus_fine, v_U, type = "l", col = "blue", lwd = 2,
       xlab = expression(mu), ylab = expression(v(mu)),
       main = ttl, ylim = ylim)
  lines(mus_fine, v_M, col = "red", lwd = 2)
  lines(mus_fine, cav_U, col = "blue", lwd = 1, lty = 2)
  lines(mus_fine, cav_M, col = "red", lwd = 1, lty = 2)
  legend("topleft", c("v(U)", "v(M)", "cav v(U)", "cav v(M)"),
         col = c("blue", "red", "blue", "red"),
         lwd = c(2, 2, 1, 1), lty = c(1, 1, 2, 2), cex = 0.6)
}
```

\textbf{Key observation}: The concavification under unanimity exploits both the entry threshold and the screening jump. Under majority, only the entry threshold is available. For priors $p$ where $\operatorname{cav} v(p, U) > \operatorname{cav} v(p, M)$, $H$ prefers unanimity.

### $H$'s optimal payoff comparison

```{r bp_comparison, echo=FALSE, results='asis'}
# Reuse cav_U and cav_M already computed in bp_plot chunk
# (They use mus_fine grid)

priors <- seq(0.05, 0.95, by = 0.05)

# Compute once for both examples
v_U1_all <- sapply(mus_fine, function(m) v_unanimity(1.1, 0.5, m, 5, beta_val, c_val))
v_M1_all <- sapply(mus_fine, function(m) v_majority(1.1, 0.5, m, 5, beta_val, c_val))
cav_U1_all <- concavify(mus_fine, v_U1_all)
cav_M1_all <- concavify(mus_fine, v_M1_all)

v_U2_all <- sapply(mus_fine, function(m) v_unanimity(1.5, 0.3, m, 5, beta_val, c_val))
v_M2_all <- sapply(mus_fine, function(m) v_majority(1.5, 0.3, m, 5, beta_val, c_val))
cav_U2_all <- concavify(mus_fine, v_U2_all)
cav_M2_all <- concavify(mus_fine, v_M2_all)

cat("\\begin{center}\n")
cat("\\begin{tabular}{c|cc|cc}\n")
cat(" & \\multicolumn{2}{c|}{Example 1} & \\multicolumn{2}{c}{Example 2} \\\\\n")
cat("$p$ & $\\operatorname{cav}v(p,U)$ & $\\operatorname{cav}v(p,M)$ & $\\operatorname{cav}v(p,U)$ & $\\operatorname{cav}v(p,M)$ \\\\\n")
cat("\\hline\n")

for (p in priors) {
  idx <- which.min(abs(mus_fine - p))
  cat(paste0(p, " & ",
             round(cav_U1_all[idx], 4), " & ", round(cav_M1_all[idx], 4), " & ",
             round(cav_U2_all[idx], 4), " & ", round(cav_M2_all[idx], 4), " \\\\\n"))
}

cat("\\end{tabular}\n")
cat("\\end{center}\n")
```

## Institutional Comparison without BP

Before deriving the full theorem with BP, we characterize the comparison \emph{without} Bayesian Persuasion (i.e., $W$ uses the prior $p$ directly, no signal).

\begin{proposition}[Comparison without BP]\label{prop:no_bp}
Fix parameters $(N, r, \alpha, \beta, c)$ in the standard range ($r > 1$, $\alpha < 1/r$, $\beta$ not too low).

\begin{enumerate}
\item $W$ always prefers majority: $V_W^{R1}(\mu, U) < V_W^{R1}(\mu, M)$ for all $\mu \in [0,1]$.
\item Conditional on entry, $H$ prefers unanimity: $E_\theta[V_H^{R1}(\theta, \mu, U)] > E_\theta[V_H^{R1}(\theta, \mu, M)]$ for all $\mu$ above both entry thresholds. The overpayment of $\theta=0$ under unanimity dominates the coalition-size advantage under majority.
\item Entry is harder under unanimity: $\tau(U) \geq \tau(M)$ (since $V_W(U) < V_W(M)$, $W$ requires a higher $\mu$ to justify entry).
\end{enumerate}
\end{proposition}

\begin{remark}[Pareto ranking without BP]\label{rem:pareto}
The comparison depends on the prior $p$:
\begin{center}
\begin{tabular}{lccc}
Prior range & $W$ prefers & $H$ prefers & Pareto ranking \\
\hline
$p < \tau(M)$ & indifferent & indifferent & both get 0 \\
$\tau(M) < p < \tau(U)$ & $M$ & $M$ & \textbf{$M$ Pareto dominates} \\
$p > \tau(U)$ & $M$ & $U$ & not comparable \\
\end{tabular}
\end{center}
For priors in $(\tau(M), \tau(U))$, entry occurs only under majority: both players prefer $M$ (something vs nothing). Without BP, unanimity is \textbf{not} Pareto superior.
\end{remark}

\begin{remark}[The role of BP --- revised]\label{rem:bp_role}
$H$ already prefers unanimity conditional on entry (even without BP). The problem is that entry is harder under unanimity ($\tau(U) > \tau(M)$). BP solves this in two ways:
\begin{enumerate}
\item \textbf{Entry inducement}: BP can induce entry at priors $p < \tau(U)$ where $W$ would not enter at the prior.
\item \textbf{Screening exploitation}: the jump in $v(\mu, U)$ at $\mu_s^{R1}$ creates a non-convexity that concavification exploits, giving $H$ additional gain.
\end{enumerate}
BP does not invert $H$'s preference (H already wants unanimity). BP \emph{viabilizes} unanimity by making entry feasible at priors where it would otherwise fail.
\end{remark}

\begin{remark}[Parameter conditions]\label{rem:params}
The ranking $V_W(U) < V_W(M)$ and $E[V_H(U)] > E[V_H(M)]$ (conditional on entry) holds for standard parameters ($\beta \geq 0.7$, $\alpha \leq 0.5$). For extreme parameters ($\beta \leq 0.5$ or $\alpha \geq 0.6$), the $H$ comparison may reverse. The theorem should state the parametric conditions explicitly.
\end{remark}

## Institutional Comparison with BP --- to derive (Theorem)

# Open Questions

\begin{enumerate}
\item \textbf{Direction of entry without direct benefit}: under unanimity (aggressive branch), $V_W$ is decreasing in $\mu$ --- entry easier at low $\mu$. Under majority, $V_W$ is increasing --- entry easier at high $\mu$. Opposite directions. Derive consequences for BP.

\item \textbf{Feasibility of absolute offers}: $W$ proposes amounts without knowing $V(\theta)$. Conservative offer $y_H = \alpha r$: feasible when $\theta=0$ ($\alpha r < 1$ since $\alpha < 1/r$)? Yes. $\checkmark$

\item \textbf{$V_W$ under majority}: when $W$ proposes and excludes $H$, $W$ keeps $V(\theta)$ but does not know $V(\theta)$. $W$ proposes amounts and receives residual $V(\theta) - \text{offers}$. Residual depends on $\theta$.

\item \textbf{Continuity of $V_W^{R2}$ at $\mu_s$}: verify algebraically.

\item \textbf{Condition for $H$ to prefer unanimity}: when does screening channel + superior extraction outweigh entry disadvantage?
\end{enumerate}
