---
title: "Pivotal-Response Rederivation: Gate 0 Contract"
subtitle: "Clean Two-Round Baseline Under Simultaneous Ballots"
author: "Formal model workspace"
date: "2026-08-11"
output:
  bookdown::html_document2:
    toc: true
    number_sections: true
    theme: readable
    code_folding: hide
  bookdown::pdf_document2:
    toc: true
    number_sections: true
    latex_engine: xelatex
geometry: margin=0.8in
fontsize: 10pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
transition_registry <- utils::read.csv(
  "../tables/pivotal_response_gate0_transitions_v1.csv",
  check.names = FALSE,
  na.strings = character()
)
information_registry <- utils::read.csv(
  "../tables/pivotal_response_information_sets_v1.csv",
  check.names = FALSE,
  na.strings = character()
)
state_registry <- utils::read.csv(
  "../tables/pivotal_response_sufficient_states_v1.csv",
  check.names = FALSE,
  na.strings = character()
)
relevance_registry <- utils::read.csv(
  "../tables/pivotal_response_relevance_registry_v1.csv",
  check.names = FALSE,
  na.strings = character()
)
proof_ledger <- utils::read.delim(
  "../quality_reports/2026-08-11_pivotal_response_proof_ledger.tsv",
  check.names = FALSE,
  quote = "",
  comment.char = ""
)
```

# Status and readback

This is a new, autonomous Gate 0 workspace. Its objective is to freeze the
extensive form needed to rederive the clean two-round model by dependency-safe
backward reasoning under **Perfect Bayesian equilibrium**. It contains no
Round-2 or Round-1 equilibrium result.

The conceptual method is:

1. reconstruct primitives and histories without importing candidate formulas;
2. define information sets, vote relevance, transitions, payoff dates, and
   assessment-level completion discipline;
3. validate the continuation DAG and obtain an independent read-only Gate 0
   review;
4. solve the three independent terminal Round-2 problems in native Round-2
   units and freeze their interfaces;
5. only then solve Round 1, entry, and the institutional comparison in reverse
   dependency order.

The new artifacts are this document, the JSON contract interface, a
dependency-complete JSON bundle, the independent-review interface and report,
the JSON DAG, four machine-readable Gate 0 registries, a protected-artifact
hash manifest, a new proof ledger and status report, and a new R verifier. The
protected manuscripts, prior derivation, quarantined workspace and its
reports, reviews, scripts, and outputs are read-only. Their initial SHA-256
values are recorded in
`quality_reports/2026-08-11_pivotal_response_protected_hash_manifest.tsv`.

The remaining work has four controlled phases: terminal Round 2 plus cold
review; two parallel Round-1 nodes plus review; entry and common-domain
comparison; and a non-migrating survival matrix plus two final read-only
audits. Each phase may reopen an upstream interface and invalidate all
descendants.

**Current status:** Gate 0 is **PASS** at `passed_order=10`. After an
initial REPAIR, independent formal-model and adversarial rereviews returned
PASS on the exact dependency-complete bundle, with zero findings. The three
Round-2 nodes are authorized and graph-ready but remain `pending`, with
no `started_order`, interface, or result. Rendered layout remains not
tested.

# Provenance and quarantine

The implementation began from:

| Field | Recorded value |
|:--|:--|
| Git root | `/Users/manoelgaldino/Documents/DCP/Papers/PowerBayesianPersuasion` |
| Branch | `main` |
| HEAD | `df3a0f2990ad7b86559ffc3ef05015e2250653a7` |
| Worktree at start | clean |
| Primitive source | `model_redesign/power_architecture_derivations.Rmd` |
| Quarantined history | `model_redesign/undominated_voting_rederivation.Rmd` |
| Contract interface | `model_redesign/pivotal_response_interfaces/gate0_contract_v1.json` |

The primitive source is used only to reconstruct the players, timing,
contract space, budgets, information, payoffs, and continuation protocol. Its
equilibrium formulas and classifications are not inputs. The quarantined file
is provenance only and supplies no result, threshold, branch, or selection to
this workspace.

Neither manuscript is edited or compiled here. No protected historical file
is edited or compiled here.

# Extensive-form contract

## Players, type, prior, and primitive domain

There are $N=m+1\geq3$ states: one hegemon $H$ and $m$ weak states
$W_1,\ldots,W_m$. Nature draws a persistent type

\[
\theta\in\{0,1\}
\]

before formation. Only $H$ observes the realization. Weak states share the
common prior

\[
\mu=\Pr(\theta=1)\in[0,1].
\]

The weak institutional surplus is fixed, public, and normalized to one. Weak
outside payoffs are zero. The clean primitive domain is

\[
0\leq o_0<o_1\leq \bar y\leq1,
\qquad
\beta\in(0,1].
\]

An agreement that includes $H$ gives it $y$; hence $b_0=b_1=0$. The
outside payoff $o_\theta$ is external to the unit institutional surplus.
The rule $R\in\{U,M\}$ is fixed exogenously; there is no rule-choice move.

All primitives, the prior, the recognition rule, the contract space, the
ballot, the transition map, payoffs, and the two tie rules below are common
knowledge.

## Formation, recognition, and horizon

After Nature draws the type, but before Round-1 recognition, the weak states
make one collective all-or-nothing formation decision without observing
$\theta$. If they form, each pays an external sunk cost $\chi\geq0$. If
they do not form, every weak state receives zero and $H$ receives
$o_\theta$.

Let $\alpha$ index a complete equilibrium assessment or an authorized
selection from the downstream payoff correspondence. The gross collective
value under rule $R$ is

\[
V_W^R(\alpha,\mu)
=
\frac{1}{m}
\mathbb E_\alpha\!\left[
  \sum_{j=1}^{m}u_j^R
  \mathrel{\big|}\text{formation}
\right].
\]

The net value is $V_W^R(\alpha,\mu)-\chi$. Formation occurs if and only if
this quantity is weakly positive; equality leads to formation. If bargaining
payoffs form a correspondence, entry is evaluated assessment by assessment.
Payoff bounds may support a selection-free claim, but they never choose an
assessment or define the formation rule.

There are exactly two bargaining rounds. Round 2 is terminal. In every reached
round, exactly one weak proposer is drawn uniformly. The draws are independent
across rounds and independent of type and public history. The hegemon is never
recognized in this baseline: $\pi_H=0$ in both rounds.

## Proposal space and budgets

When $H$ is active and weak state $i$ is recognized, it proposes

\[
s_i=\left(y,(x_j)_{j\ne i}\right),
\qquad
0\leq y\leq\bar y,
\qquad
x_j\geq0.
\]

If the current agreement includes $H$, the proposer's residual is

\[
x_i^H=1-y-\sum_{j\ne i}x_j\geq0.
\]

If a majority agreement passes without $H$, the participation component is
not burned. It is reabsorbed by the proposer:

\[
x_i^{WO}=1-\sum_{j\ne i}x_j=x_i^H+y.
\]

A named $x_j$ is paid whenever the current proposal passes, even if $j$
voted no. The participation component $y$ is paid only when the implemented
agreement includes $H$. After $H$ has opted out, a weak-only proposal
fixes $y=0$ and allocates the unit weak surplus.

The proposal space contains gifts to voters whose ballots may not be needed
for passage. They are not removed from the action space or from deviation
checks.

## Simultaneous ballot and quotas

The proposer is counted as voting yes. All other active players observe the
public history and current proposal, then vote yes or no simultaneously. $H$
also observes its private type. No voter observes another current vote when
acting. The complete vector and outcome become public only after every vote
has been cast.

In the absence of an authorized correlating device, behavioral randomizations
are independent conditional on the player's information set. A derivation may
not insert correlated ballots.

The quotas are

\[
Q_U=N,
\qquad
Q_M=q=\left\lfloor\frac N2\right\rfloor+1.
\]

After $H$ opts out, the original quota is retained rather than recalculated.
Because only $m=N-1$ players then remain, the unanimity quota is impossible.
That branch is a terminal outcome, not an artificial continuation game.

## Immediate opt-out and continuation

A no vote by type $\theta$ of $H$ is an immediate and irreversible
opt-out. No current or future agreement can include it after that action. In
Round 1 the action pays $o_\theta$ immediately, without discount. In Round
2 it pays $o_\theta$ at the Round-2 date.

If $H$ votes yes in Round 1 but weak votes make the current ballot fail,
$H$ remains active and the appropriate active-$H$ Round-2 state is reached.
If $H$ votes no under majority, a current weak-only agreement passes when
the weak yes votes reach $q$; otherwise the game reaches the terminal
weak-only Round-2 majority problem. A Round-1 no vote by $H$ under
unanimity ends at the quota-impossible terminal outcome.

If an active-$H$ Round-2 ballot fails, $H$ receives $o_\theta$ in
Round-2 units and weak states receive zero. If $H$ opted out in Round 1, it
receives no additional Round-2 flow.

## Payoff dates and discounting

Round 2 is solved entirely in its native current units, with no internal
discount factor. In the backward step, **beta is applied exactly once only
when a frozen Round-2 value enters Round 1**. Immediate Round-1 opt-out is not
discounted. A predecessor must import a literal frozen continuation interface;
it may not rederive or repair that continuation locally.

## Solution concept and beliefs

The solution concept is Perfect Bayesian equilibrium. An assessment specifies
formation, proposal strategies, behavioral ballot strategies at every public
proposal history, $H$'s type-contingent ballots, and beliefs wherever the
type remains payoff-relevant. Strategies must be sequentially rational and
beliefs must follow Bayes' rule wherever possible.

The full public history and the posterior are separate state objects.
Continuation strategies may depend on the proposal and the complete revealed
vote vector even when two histories induce the same posterior. At a
zero-probability history, the required belief must be stated as part of the
assessment; it may not be introduced inside a proof to suppress a branch.

The maintained weak-vote-passive belief discipline reflects that weak voters
do not observe the type: a unilateral weak vote is not direct evidence about
$\theta$. A type-contingent vote by $H$ may update beliefs once the ballot
closes. After completed opt-out, the type is payoff-irrelevant to the
weak-only continuation.

## Direct response convention and proposal ties

When a player's vote changes a payoff-relevant event with positive
probability, it votes yes when the expected payoff from yes is greater than or
equal to the expected payoff from no. It votes no when the expected payoff
from no is strictly larger. At equality with positive probability of
relevance, the player votes yes.

When relevance has probability zero, the two actions are locally
outcome-equivalent holding the other strategies fixed. The assessment must
still record a complete behavioral profile, but that bookkeeping action is not
a substantive equilibrium class. Section \@ref(completion-discipline) explains
why local equivalence does not license simultaneous replacement of several
actions.

Only after maximizing the weak proposer's payoff, a proposal-payoff tie is
resolved by minimizing $H$'s expected payoff. If both the proposer and $H$
tie while downstream objects differ, the outcome remains a correspondence.

# Vote relevance

## Outcome signature

Fix player $i$'s ballot information set $I$, its action
$a_i\in\{Y,N\}$, the other simultaneous actions $a_{-i}$, and the type
when it remains relevant. Define the outcome signature

\[
\Phi_I(a_i,a_{-i},\theta)
=
\bigl(
A,
J_H,
O_H,
P,
S
\bigr),
\]

where:

- $A$ records approval or rejection;
- $J_H$ records inclusion or exclusion of $H$;
- $O_H$ records opt-out status;
- $P$ records the implemented payment vector and the destination of $y$;
- $S$ records the induced terminal outcome or the continuation
  outcome/interface under the frozen assessment.

If an action reaches a continuation, $S$ is the action-specific object
$\kappa_{C_2}(h_2)$, including its induced belief and payoff distribution.
It is not a raw successor label or a raw vote vector. Thus, in Round 1, two
actions that reach the same named $C_2$ class can remain relevant when they
produce different full histories and
$\kappa_{C_2}(h_2^Y)\ne\kappa_{C_2}(h_2^N)$.

The other-action profile is relevant for voter $i$ exactly when

\[
\Phi_I(Y,a_{-i},\theta)\ne
\Phi_I(N,a_{-i},\theta).
\]

This definition is conditional on the information set. It does not condition
on a simultaneous action the voter has not observed.

## Expected voting incentive

Let ${\cal R}_i(I)$ be the set of relevant pairs
$(\theta,a_{-i})$. The expected difference is

\[
\Delta_i(I)
=
\sum_{(\theta,a_{-i})\in{\cal R}_i(I)}
\Pr(\theta,a_{-i}\mid I)
\left[
u_i(Y,a_{-i};\theta,I)
-u_i(N,a_{-i};\theta,I)
\right].
\]

Every profile outside ${\cal R}_i(I)$ contributes exactly zero. If
$\Pr({\cal R}_i(I)\mid I)>0$, the prescribed response is yes when
$\Delta_i(I)\geq0$ and no when $\Delta_i(I)<0$. This is an expected
comparison over events the vote actually changes, not a blanket quota test.

For weak voters, relevant profiles normally occur at quota boundaries and can
change current payment versus continuation. For $H$, relevance must be
derived separately. Even when weak votes already meet a majority quota,
changing $H$'s ballot can change inclusion, payment $y$, opt-out, or the
continuation. Terminal opt-out status is part of the outcome signature even
when terminal monetary payoffs happen to coincide.

## Formal relevance registry

The authoritative registry is
`tables/pivotal_response_relevance_registry_v1.csv`. It separates weak and
hegemon ballots by round, rule, participation status, and information set.

```{r relevance-table, results='asis'}
knitr::kable(
  relevance_registry,
  caption = "Gate 0 vote-relevance registry."
)
```

For a weak nonproposer $j$, let $w_{-j}\in\{1,\ldots,m-1\}$ count weak
yes votes other than $j$, including the proposer's automatic yes. The
current-approval structural toggle occurs exactly in the following cases:

- active-$H$ unanimity: $H=Y$ and $w_{-j}=m-1$; no weak toggle can
  restore unanimity after $H=N$;
- active-$H$ majority with $H=Y$: $w_{-j}=q-2$, an empty set when
  $N=3$ because then $q-2=0<w_{-j}$;
- active-$H$ majority with $H=N$: $w_{-j}=q-1$;
- weak-only majority: $w_{-j}=q-1$.

These quota cases do not exhaust Round-1 relevance when both actions fail.
The action-specific full histories must still be passed through the frozen
continuation interface. The verifier enumerates the structural cases for
$N=3,\ldots,60$, including the empty $N=3$ majority-with-$H=Y$ case. The
hegemon is analyzed separately for every $z\in\{1,\ldots,m\}$; no weak-voter
quota shortcut is applied to its inclusion, opt-out, payment, or continuation
comparison.

## Product-safe completion discipline {#completion-discipline}

Local irrelevance is not closed under changing several actions jointly.
Consider terminal unanimity with $N=4$. The proposer and $H$ vote yes and
the two weak nonproposers vote no. Holding the other no fixed, either weak
voter can switch alone without producing passage, so each action is locally
irrelevant. If both switch to yes, the proposal passes. Terminal weak-only
majority with $N=4,q=3$ has the same structure.

Accordingly, this Gate 0 does not impose a universal player-by-player
completion. Every assessment records an explicit full behavioral profile at
each proposal history. Two full completions may be quotiented only if they
induce identical approval, $H$ inclusion and opt-out, transfers, successor
state, relevant belief distribution, and payoff distribution at every
affected proposal and history.

Any later invariance claim must be product-safe: it must compare the induced
joint distributions, not merely verify that each vote is individually
irrelevant at one profile. If this proof fails, the different completions
remain distinct outcome classes or a correspondence. Proposal optimality must
also remain unchanged before a result is called completion-invariant.

# Coalitions, gifts, and proposal optimality

No coalition-size restriction is a primitive or an additional equilibrium
concept. Offers required to obtain support must be derived from ballot
incentives. **Gifts to outsiders remain feasible deviations** and must be
included in the recognized proposer's best-response problem.

Nor does local irrelevance generally imply zero outsider gifts or a minimum
winning representation. In terminal weak-only majority with $N=5,q=3$, an
on-path proposal may have the proposer plus all three other weak voters voting
yes. Each nonproposer is locally irrelevant against the other two yes votes,
and the proposal passes. At a different, zero-gift proposal history, all three
nonproposers may vote no; each unilateral switch remains locally irrelevant
because proposer plus one yes still falls short of $q$, and the proposal
fails. Deleting a gift changes the proposal information set and need not
preserve its completion or the proposer's payoff.

This finding is not confined to a knife-edge. An open set of feasible positive
gift vectors can accompany the all-yes completion, while a distinct deviated
proposal receives the all-no completion. Therefore a general zero-gift or
minimum-coalition conclusion is rejected in
this contract. A compact minimum-winning representation may be used only for a
proved subclass in which passage is secured or full completion invariance is
product-safe and proposer optimality removes every redundant transfer.
Otherwise the proof must retain unrestricted proposal-contingent completions,
positive gifts, oversized support, and any resulting correspondence.

# Histories and information sets

The authoritative information-set registry is
`tables/pivotal_response_information_sets_v1.csv`. It records what every mover
observes, what remains unobserved, when actions become public, the feasible
actions, successors, and the relevant belief object.

```{r information-set-table, results='asis'}
knitr::kable(
  information_registry,
  caption = "Gate 0 information sets and observations."
)
```

The key safeguards are:

- weak ballot strategies cannot condition on the type or another current
  vote;
- $H$'s strategy can condition on its type but not on current weak votes;
- a Round-2 state retains the complete public Round-1 proposal and vote vector
  separately from its posterior;
- the weak-only Round-2 state requires no belief about the type for its payoff
  problem.

# Exhaustive transition registry

Let $z\in\{1,\ldots,m\}$ be the number of weak yes votes, including the
proposer's automatic yes. The authoritative transition registry is
`tables/pivotal_response_gate0_transitions_v1.csv`.

```{r transition-table, results='asis'}
knitr::kable(
  transition_registry,
  caption = "Gate 0 exhaustive ballot transitions in native payoff units."
)
```

The R verifier enumerates every admissible ballot tuple for
$N=3,\ldots,60$, both rules, both rounds, active-$H$ yes and no, and the
terminal weak-only majority state. Each tuple must map to exactly one of
PR01--PR16. The symbolic range $1,\ldots,q-2$ is empty at $q=2$, rather than
new histories.

No terminal outcome is promoted to a decision problem. In particular, after
an $H$-no under unanimity, the original quota is impossible and the branch
ends.

# Sufficient states

The authoritative sufficient-state registry is
`tables/pivotal_response_sufficient_states_v1.csv`.

```{r state-table, results='asis'}
knitr::kable(
  state_registry,
  caption = "Gate 0 payoff- and belief-sufficient states."
)
```

The registry separates pre-recognition, proposer-before-$s_i$, weak-ballot,
and $H$-ballot states. A weak-ballot key retains both voter $j$ and the
public proposal $s_i$; an $H$-ballot key retains $s_i$ and private
$\theta$. Each $C_2(h_2)$ imported by Round 1 is a pre-recognition
correspondence indexed by the full public $h_2$ and integrates the uniform
proposer draw. Histories may be compressed only after proving that feasible
actions, payoff consequences, beliefs, admissible strategies, and all future
continuations coincide.

# Continuation DAG and barriers

The machine-readable graph is
`model_redesign/pivotal_response_game_dag.json`. The Gate 0 contract is a
`pass` protocol node whose artifact is the dependency-complete
`model_redesign/pivotal_response_interfaces/gate0_bundle_v1.json`. The bundle
hash covers the contract plus the transition, information-set,
sufficient-state, and relevance registries. The separate
`gate0_review_v1.json` records both independent PASS decisions and the
event log through `passed_order=10`. Because the gate is closed, the
generic DAG checker reports exactly the three terminal Round-2 problems as the
ready antichain.

The dependency-first schedule is:

| Batch | Nodes | Required frozen inputs | Current authorization |
|:--|:--|:--|:--|
| Gate 0 | `gate0_contract` | none | PASS at event 10 |
| R2 antichain | `r2_unanimity_active_h`; `r2_majority_active_h`; `r2_majority_weak_only` | reviewed Gate 0 hash | authorized and ready; not started |
| R2 batch gate | `r2_batch_review` | all three R2 interfaces and independent review | blocked |
| R1 antichain | `r1_unanimity`; `r1_majority` | `r2_batch_review` plus local R2 hashes | blocked |
| R1 batch gate | `r1_batch_frozen` | both R1 interfaces frozen | blocked |
| Entry | `entry_unanimity`; `entry_majority` | `r1_batch_frozen` plus corresponding R1 hash | blocked |
| Comparison | `institutional_comparison` | both entry hashes and common existence domain | blocked |
| Survival inventory | `v6_survival_matrix` | comparison hash | blocked; no migration |

The three Round-2 nodes are graph-independent conditional on Gate 0. The two
Round-1 nodes remain blocked until `r2_batch_review` independently accepts and
freezes all three terminal interfaces. Entry remains blocked until
`r1_batch_frozen` freezes both Round-1 interfaces. Entry then consumes the
corresponding Round-1 payoff correspondence assessment by assessment; a valid
bound can support only a selection-free claim. Institutional comparison is
restricted to the common equilibrium-existence domain; no scalar selection is
silently inserted.

The following are terminal outcomes, not nodes: nonformation, current passage
with $H$, current weak-only passage, terminal active-$H$ failure, terminal
weak-only failure, and the quota-impossible unanimity outcome after opt-out.

# Continuation interfaces and native-time ledger

The versioning policy is in
`model_redesign/pivotal_response_interfaces/README.md`. Gate 0 has a
versioned contract, dependency-complete bundle, and independent-review
interface. Its contract and four authoritative registries are hash-closed by
`gate0_bundle_v1.json`. No Round-2 interface file exists.

Every future closed continuation interface must state strategies, beliefs,
payoff vector or correspondence, native payoff date, existence and
multiplicity, boundaries, equality cases, assumptions, checks, status,
provenance, and SHA-256.

| Imported object | Native date | Consuming date | Permitted transformation | Status |
|:--|:--|:--|:--|:--|
| $C_{2,U}^{A}(h_2)$ | R2 | R1 | $\beta C_{2,U}^{A}(h_2)$ exactly once | nonexistent; R2 not started |
| $C_{2,M}^{A}(h_2)$ | R2 | R1 | $\beta C_{2,M}^{A}(h_2)$ exactly once | nonexistent; R2 not started |
| $C_{2,M}^{WO}(h_2)$ | R2 | R1 weak payoffs only | $\beta C_{2,M}^{WO}(h_2)$ exactly once | nonexistent; R2 not started |
| immediate $o_\theta$ after $H$-no in R1 | R1 | R1 | none | primitive |

Changing any Gate 0 bundle component changes the bundle hash, reopens Gate 0,
and invalidates every descendant derivation and review. Likewise, changing any
later frozen interface hash reopens that node and invalidates its descendants.
A predecessor may not fill an interface gap locally.

# Proof ledger

```{r proof-ledger-table, results='asis'}
knitr::kable(
  proof_ledger,
  caption = "Pivotal-response proof and execution ledger."
)
```

The statuses `checked numerically` in Gate 0 refer only to hash, partition, or
graph checks. The extensive form, relevance discipline, sufficient states,
transitions, and DAG are `proved` in the ledger only in the governance
sense that both independent reviewers accepted the exact bundle and source
with zero findings. The general zero-gift or minimum-coalition conclusion
remains `rejected` by the proposal-contingent completion counterexample.
Every equilibrium node remains `pending` and has no `started_order` or
`passed_order`.

# Independent Gate 0 review closure

The formal-model and adversarial read-only reviewers independently answered
all of the following questions affirmatively for the exact frozen bundle:

1. PR01--PR16 exhaust and partition the feasible ballot profiles;
2. the information sets prevent conditioning on unobserved current votes;
3. vote relevance includes approval, participation, opt-out, payments, and
   the induced terminal or assessed continuation outcome without using a raw
   successor label or collapsing $H$'s problem to quota pivotality;
4. equality is resolved as yes only when relevance has positive probability;
5. assessment-level completions and independent behavioral mixing are defined
   without overclaiming invariance from individual local irrelevance;
6. the N=4 joint-completion and N=5 gift counterexamples are handled
   correctly;
7. the unrestricted proposal space preserves gifts as deviations and avoids a
   general minimum-coalition claim;
8. the DAG enforces the Gate 0, R2-batch-to-R1, and
   R1-batch-to-entry barriers;
9. all Round-2 payoffs are native-date objects and discounting occurs exactly
   once upon import to Round 1;
10. no equilibrium result has entered the Gate 0 contract.

The first review round returned REPAIR on contract hash
`37bd422df0bafd7c93162198594c49fd3114d304750d282318d7a705348bc9bb`.
After repair, both rereviews returned PASS with zero critical, major, or minor
findings on bundle
`sha256:6e28cb3faf3b70bc5ed990ce35a9e39326be15f93252e93e6663d771e2b0b7c1`.
The closure and conditions are recorded in
`model_redesign/pivotal_response_interfaces/gate0_review_v1.json` and
`quality_reports/2026-08-11_pivotal_response_gate0_independent_review.md`.
No unresolved Gate 0 protocol choice remains. Any bundle-component change
reopens the gate and invalidates all descendants.

# Stop boundary

`STOP_BEFORE_R2`

This file ends at the Gate 0 contract and work agenda. It does not solve any
terminal ballot, characterize any proposal, compute any continuation value,
or import any prior equilibrium result.
