---
title: "Pivotal-Response Rederivation"
subtitle: "Clean Two-Round Baseline Under Simultaneous Sealed Ballots"
author: "Formal model workspace"
date: "2026-08-12"
output:
  bookdown::html_document2:
    toc: true
    number_sections: true
    theme: readable
    code_folding: hide
  bookdown::pdf_document2:
    toc: true
    number_sections: true
    latex_engine: xelatex
    keep_tex: true
geometry: margin=0.8in
fontsize: 10pt
header-includes:
  - \setlength{\emergencystretch}{3em}
  - \raggedbottom
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
transition_registry <- utils::read.csv(
  "../tables/pivotal_response_gate0_transitions_v1.csv",
  check.names = FALSE,
  na.strings = character()
)
information_registry <- utils::read.csv(
  "../tables/pivotal_response_information_sets_v1.csv",
  check.names = FALSE,
  na.strings = character()
)
state_registry <- utils::read.csv(
  "../tables/pivotal_response_sufficient_states_v1.csv",
  check.names = FALSE,
  na.strings = character()
)
relevance_registry <- utils::read.csv(
  "../tables/pivotal_response_relevance_registry_v1.csv",
  check.names = FALSE,
  na.strings = character()
)
proof_ledger <- utils::read.delim(
  "../quality_reports/2026-08-11_pivotal_response_proof_ledger.tsv",
  check.names = FALSE,
  quote = "",
  comment.char = ""
)
survival_matrix <- utils::read.csv(
  "../tables/pivotal_response_v6_survival_matrix_v1.csv",
  check.names = FALSE,
  na.strings = character()
)

include_node <- function(path, heading_shift = 1L) {
  text <- paste(
    readLines(path, warn = FALSE, encoding = "UTF-8"),
    collapse = "\n"
  )
  snapshot_updates <- c(
    "- Status: candidate, pending independent read-only review." =
      "- Post-freeze status: exact interface accepted by two independent read-only reviews.",
    "- Status: repaired candidate, pending independent read-only rereview." =
      "- Post-freeze status: repaired exact interface accepted by two independent read-only rereviews.",
    "**Status.** Repaired candidate pending independent read-only rereview" =
      "**Post-freeze status.** Repaired exact interface accepted by two independent read-only rereviews.",
    "It remains pending independent rereview." =
      "Its repaired exact hash was accepted by two independent read-only rereviews.",
    "- Interface status: **candidate pending independent review**." =
      "- Post-freeze interface status: **PASS after two independent read-only reviews**.",
    "| Independent formal/adversarial acceptance | not yet performed | pending |" =
      "| Independent formal/adversarial acceptance | two exact-hash read-only reviews | proved by review bundle |",
    "| Rendered document layout | not attempted for this node note | pending |" =
      "| Inclusion in standalone output | post-freeze render | checked separately |",
    "| Rendered layout | not attempted | pending |" =
      "| Inclusion in standalone output | post-freeze render | checked separately |",
    "| Independent formal acceptance | pending |" =
      "| Independent read-only reviews | PASS on the exact repaired hash |",
    "`proved_consumable_pending_independent_rereview`" =
      "`proved_consumable_after_independent_rereview`",
    "The exact candidate interface is" =
      "The exact reviewed interface is",
    "invalidates this candidate" =
      "invalidates this frozen interface",
    "This implementer does not mark the node PASS." =
      "The independent review bundle later marked the exact node PASS.",
    "# v6 survival matrix: claim-level, nonmigratory classification" =
      paste0(
        "# v6 survival matrix: claim-level, nonmigratory classification\n\n",
        "This frozen inventory records how the completed rederivation bears on historical claims; ",
        "it neither selects an equilibrium nor authorizes manuscript migration."
      )
  )
  for (old in names(snapshot_updates)) {
    text <- gsub(old, snapshot_updates[[old]], text, fixed = TRUE)
  }
  text <- gsub(
    paste0(
      "This artifact is a candidate pending independent read-only review. Its R\n",
      "verifier checks"
    ),
    paste0(
      "The exact artifact was subsequently accepted by two independent read-only reviews. Its R\n",
      "verifier checks"
    ),
    text,
    fixed = TRUE
  )
  text <- gsub(
    paste0(
      "This artifact remains a candidate until two independent read-only\n",
      "reviews accept its exact hash. No survival-matrix or manuscript claim may use\n",
      "it as frozen before then."
    ),
    paste0(
      "The exact artifact was subsequently accepted by two independent read-only\n",
      "reviews. Its frozen review bundle, rather than this implementation snapshot,\n",
      "authorizes the nonmigrating survival inventory to consume it."
    ),
    text,
    fixed = TRUE
  )
  text <- gsub(
    paste0(
      "This candidate remains nonmigratory and pending two independent read-only\n",
      "reviews. `formal_model_v6.Rmd`, `formal_model_v5.Rmd`, the prior derivation,\n",
      "their outputs, the shared Rmd, DAG, and proof ledger were not edited or\n",
      "compiled. No protected-target migration is authorized."
    ),
    paste0(
      "The exact nonmigratory candidate was subsequently accepted by two independent\n",
      "read-only reviews and frozen by its review bundle. The protected manuscripts\n",
      "and prior derivation remained byte-identical. No protected-target migration is\n",
      "authorized."
    ),
    text,
    fixed = TRUE
  )
  text <- gsub(
    paste0(
      "This repaired export covers `N>=3`, including the attained N=3 construction\n",
      "in Proposition 4. Downstream entry and institutional comparison may consume\n",
      "it only after independent rereview accepts this exact candidate hash."
    ),
    paste0(
      "This repaired export covers `N>=3`, including the attained N=3 construction\n",
      "in Proposition 4. Its exact hash was accepted by two independent rereviews and\n",
      "is consumed downstream only through the frozen R1 batch."
    ),
    text,
    fixed = TRUE
  )
  text <- gsub(
    paste0(
      "Any byte change to Gate 0, the R1 batch, or C1-U invalidates this candidate.\n",
      "This implementer does not mark the node PASS."
    ),
    paste0(
      "Any byte change to Gate 0, the R1 batch, or C1-U invalidates this interface.\n",
      "The exact repaired hash was later marked PASS by its independent review bundle."
    ),
    text,
    fixed = TRUE
  )
  text <- gsub(
    paste0(
      "invalidates this node and every descendant. Reviewers must audit the exact\n",
      "candidate hash and remain read-only."
    ),
    paste0(
      "invalidates this node and every descendant. Two independent read-only reviews\n",
      "subsequently accepted the exact candidate hash."
    ),
    text,
    fixed = TRUE
  )
  text <- gsub(
    "sha256:([a-f0-9]{8})[a-f0-9]{56}",
    "sha256:\\1...",
    text,
    perl = TRUE
  )
  text <- gsub(
    "(^|[^a-f0-9])([a-f0-9]{8})[a-f0-9]{56}([^a-f0-9]|$)",
    "\\1\\2...\\3",
    text,
    perl = TRUE
  )
  text <- gsub(
    "model_redesign/pivotal_response_interfaces/",
    "interfaces/",
    text,
    fixed = TRUE
  )
  text <- gsub(
    "model_redesign/power_architecture_derivations.Rmd",
    "power_architecture_derivations.Rmd",
    text,
    fixed = TRUE
  )
  text <- gsub(
    "verify_pivotal_response_[A-Za-z0-9_]+[.]R",
    "node-specific R verifier",
    text,
    perl = TRUE
  )
  text <- gsub(
    "scripts/node-specific R verifier",
    "scripts/verify_pivotal_response_v6_survival_matrix.R",
    text,
    fixed = TRUE
  )
  text <- gsub(
    "`delta_j^d(theta,a_-j)`",
    "$\\delta_j^d(\\theta,a_{-j})$",
    text,
    fixed = TRUE
  )
  text <- gsub("formal_model_v5.Rmd", "v5 manuscript", text, fixed = TRUE)
  text <- gsub("formal_model_v6.Rmd", "v6 manuscript", text, fixed = TRUE)
  text <- gsub(
    "power_architecture_derivations.Rmd",
    "prior derivation",
    text,
    fixed = TRUE
  )
  lines <- strsplit(text, "\n", fixed = TRUE)[[1L]]
  flattened <- character()
  idx <- 1L
  while (idx <= length(lines)) {
    is_table <- idx < length(lines) &&
      grepl("^[[:space:]]*\\|.*\\|[[:space:]]*$", lines[[idx]]) &&
      grepl("^[[:space:]]*\\|[|: -]+\\|[[:space:]]*$", lines[[idx + 1L]])
    if (!is_table) {
      flattened <- c(flattened, lines[[idx]])
      idx <- idx + 1L
      next
    }
    split_cells <- function(line) {
      line <- sub("^[[:space:]]*\\|", "", line)
      line <- sub("\\|[[:space:]]*$", "", line)
      trimws(strsplit(line, "\\|", fixed = FALSE)[[1L]])
    }
    headers <- split_cells(lines[[idx]])
    idx <- idx + 2L
    flattened <- c(flattened, "")
    while (
      idx <= length(lines) &&
      grepl("^[[:space:]]*\\|.*\\|[[:space:]]*$", lines[[idx]])
    ) {
      cells <- split_cells(lines[[idx]])
      if (length(cells) < length(headers)) {
        cells <- c(cells, rep("", length(headers) - length(cells)))
      }
      flattened <- c(
        flattened,
        paste0("- **", cells[[1L]], ".**")
      )
      for (col_idx in seq.int(2L, length(headers))) {
        flattened <- c(
          flattened,
          paste0("  - **", headers[[col_idx]], ":** ", cells[[col_idx]])
        )
      }
      idx <- idx + 1L
    }
    flattened <- c(flattened, "")
  }
  lines <- flattened
  in_fence <- FALSE
  for (idx in seq_along(lines)) {
    if (grepl("^[[:space:]]*```", lines[[idx]])) {
      in_fence <- !in_fence
    } else if (!in_fence && grepl("^#{1,5}[[:space:]]", lines[[idx]])) {
      lines[[idx]] <- sub(
        "^(#{1,5})[[:space:]]+[0-9]+([.][0-9]+)*[.]?[[:space:]]+",
        "\\1 ",
        lines[[idx]],
        perl = TRUE
      )
      lines[[idx]] <- paste0(strrep("#", heading_shift), lines[[idx]])
    }
  }
  cat(paste(lines, collapse = "\n"), "\n")
  invisible(NULL)
}

compact_table <- function(x, caption, font_size = 6) {
  out <- knitr::kable(
    x,
    caption = caption,
    booktabs = knitr::is_latex_output()
  )
  if (knitr::is_latex_output()) {
    out <- kableExtra::kable_styling(
      out,
      latex_options = "scale_down",
      font_size = font_size
    )
  }
  out
}
```

# Status and readback

This autonomous workspace rederives the clean two-round model by
dependency-safe backward reasoning under **Perfect Bayesian equilibrium**. It
does not add a global deletion of weakly dominated actions. A voter chooses
yes at a payoff-relevant equality; when its action has zero probability of
changing any payoff-relevant outcome, the assessment records an explicit
completion without treating that completion as the substantive equilibrium
object.

The conceptual method is:

1. reconstruct primitives and histories without importing candidate formulas;
2. define information sets, vote relevance, transitions, payoff dates, and
   assessment-level completion discipline;
3. validate the continuation DAG and obtain an independent read-only Gate 0
   review;
4. solve and freeze the three independent terminal Round-2 problems in native
   Round-2 units;
5. import those interfaces literally into Round 1, applying `beta` exactly
   once where a terminal continuation enters a Round-1 comparison;
6. derive collective entry, institutional comparison, and a nonmigrating
   claim-level survival matrix only after their dependencies pass.

The artifacts include this document, versioned continuation interfaces,
dependency-complete review bundles, detailed node proofs, the JSON DAG, four
machine-readable Gate 0 registries, claim-level and verification tables, a
protected-artifact hash manifest, a proof ledger, status reports, and separate
R verifiers. The protected manuscripts, prior derivation, quarantined
workspace and its historical reports, reviews, scripts, and outputs remain
read-only. Their initial SHA-256 values are recorded in
the protected-artifact manifest.

**Analytic status:** the complete DAG is **PASS** through the nonmigrating
survival inventory at `passed_order=67`; the dependency checker reports
`Ready: none`. Every substantive node was implemented by an agent distinct
from its read-only reviewers. Rejected candidates remain in the ledger as
provenance and are not consumed. The integrated document and its rendered
outputs are a post-freeze presentation layer: they may summarize or reproduce
the frozen objects, but they cannot repair or reinterpret them.

# Provenance and quarantine

The implementation began from:

| Field | Recorded value |
|:--|:--|
| Git root | repository `PowerBayesianPersuasion` |
| Branch | `main` |
| HEAD | `df3a0f2...` |
| Worktree at start | clean |
| Primitive source | `power_architecture_derivations.Rmd` |
| Quarantined history | `undominated_voting_rederivation.Rmd` |
| Contract interface | `gate0_contract_v1.json` |

The primitive source is used only to reconstruct the players, timing,
contract space, budgets, information, payoffs, and continuation protocol. Its
equilibrium formulas and classifications are not inputs. The quarantined file
is provenance only and supplies no result, threshold, branch, or selection to
this workspace.

Neither manuscript is edited or compiled here. No protected historical file
is edited or compiled here.

# Extensive-form contract

## Players, type, prior, and primitive domain

There are $N=m+1\geq3$ states: one hegemon $H$ and $m$ weak states
$W_1,\ldots,W_m$. Nature draws a persistent type

\[
\theta\in\{0,1\}
\]

before formation. Only $H$ observes the realization. Weak states share the
common prior

\[
\mu=\Pr(\theta=1)\in[0,1].
\]

The weak institutional surplus is fixed, public, and normalized to one. Weak
outside payoffs are zero. The clean primitive domain is

\[
0\leq o_0<o_1\leq \bar y\leq1,
\qquad
\beta\in(0,1].
\]

An agreement that includes $H$ gives it $y$; hence $b_0=b_1=0$. The
outside payoff $o_\theta$ is external to the unit institutional surplus.
The rule $R\in\{U,M\}$ is fixed exogenously; there is no rule-choice move.

All primitives, the prior, the recognition rule, the contract space, the
ballot, the transition map, payoffs, and the two tie rules below are common
knowledge.

## Formation, recognition, and horizon

After Nature draws the type, but before Round-1 recognition, the weak states
make one collective all-or-nothing formation decision without observing
$\theta$. If they form, each pays an external sunk cost $\chi\geq0$. If
they do not form, every weak state receives zero and $H$ receives
$o_\theta$.

Let $\alpha$ index a complete equilibrium assessment or an authorized
selection from the downstream payoff correspondence. The gross collective
value under rule $R$ is

\[
V_W^R(\alpha,\mu)
=
\frac{1}{m}
\mathbb E_\alpha\!\left[
  \sum_{j=1}^{m}u_j^R
  \mathrel{\big|}\text{formation}
\right].
\]

The net value is $V_W^R(\alpha,\mu)-\chi$. Formation occurs if and only if
this quantity is weakly positive; equality leads to formation. If bargaining
payoffs form a correspondence, entry is evaluated assessment by assessment.
Payoff bounds may support a selection-free claim, but they never choose an
assessment or define the formation rule.

There are exactly two bargaining rounds. Round 2 is terminal. In every reached
round, exactly one weak proposer is drawn uniformly. The draws are independent
across rounds and independent of type and public history. The hegemon is never
recognized in this baseline: $\pi_H=0$ in both rounds.

## Proposal space and budgets

When $H$ is active and weak state $i$ is recognized, it proposes

\[
s_i=\left(y,(x_j)_{j\ne i}\right),
\qquad
0\leq y\leq\bar y,
\qquad
x_j\geq0.
\]

If the current agreement includes $H$, the proposer's residual is

\[
x_i^H=1-y-\sum_{j\ne i}x_j\geq0.
\]

If a majority agreement passes without $H$, the participation component is
not burned. It is reabsorbed by the proposer:

\[
x_i^{WO}=1-\sum_{j\ne i}x_j=x_i^H+y.
\]

A named $x_j$ is paid whenever the current proposal passes, even if $j$
voted no. The participation component $y$ is paid only when the implemented
agreement includes $H$. After $H$ has opted out, a weak-only proposal
fixes $y=0$ and allocates the unit weak surplus.

The proposal space contains gifts to voters whose ballots may not be needed
for passage. They are not removed from the action space or from deviation
checks.

## Simultaneous ballot and quotas

The proposer is counted as voting yes. All other active players observe the
public history and current proposal, then vote yes or no simultaneously. $H$
also observes its private type. No voter observes another current vote when
acting. The complete vector and outcome become public only after every vote
has been cast.

In the absence of an authorized correlating device, behavioral randomizations
are independent conditional on the player's information set. A derivation may
not insert correlated ballots.

The quotas are

\[
Q_U=N,
\qquad
Q_M=q=\left\lfloor\frac N2\right\rfloor+1.
\]

After $H$ opts out, the original quota is retained rather than recalculated.
Because only $m=N-1$ players then remain, the unanimity quota is impossible.
That branch is a terminal outcome, not an artificial continuation game.

## Immediate opt-out and continuation

A no vote by type $\theta$ of $H$ is an immediate and irreversible
opt-out. No current or future agreement can include it after that action. In
Round 1 the action pays $o_\theta$ immediately, without discount. In Round
2 it pays $o_\theta$ at the Round-2 date.

If $H$ votes yes in Round 1 but weak votes make the current ballot fail,
$H$ remains active and the appropriate active-$H$ Round-2 state is reached.
If $H$ votes no under majority, a current weak-only agreement passes when
the weak yes votes reach $q$; otherwise the game reaches the terminal
weak-only Round-2 majority problem. A Round-1 no vote by $H$ under
unanimity ends at the quota-impossible terminal outcome.

If an active-$H$ Round-2 ballot fails, $H$ receives $o_\theta$ in
Round-2 units and weak states receive zero. If $H$ opted out in Round 1, it
receives no additional Round-2 flow.

## Payoff dates and discounting

Round 2 is solved entirely in its native current units, with no internal
discount factor. In the backward step, **beta is applied exactly once only
when a frozen Round-2 value enters Round 1**. Immediate Round-1 opt-out is not
discounted. A predecessor must import a literal frozen continuation interface;
it may not rederive or repair that continuation locally.

## Solution concept and beliefs

The solution concept is Perfect Bayesian equilibrium. An assessment specifies
formation, proposal strategies, behavioral ballot strategies at every public
proposal history, $H$'s type-contingent ballots, and beliefs wherever the
type remains payoff-relevant. Strategies must be sequentially rational and
beliefs must follow Bayes' rule wherever possible.

The full public history and the posterior are separate state objects.
Continuation strategies may depend on the proposal and the complete revealed
vote vector even when two histories induce the same posterior. At a
zero-probability history, the required belief must be stated as part of the
assessment; it may not be introduced inside a proof to suppress a branch.

The maintained weak-vote-passive belief discipline reflects that weak voters
do not observe the type: a unilateral weak vote is not direct evidence about
$\theta$. A type-contingent vote by $H$ may update beliefs once the ballot
closes. After completed opt-out, the type is payoff-irrelevant to the
weak-only continuation.

## Direct response convention and proposal ties

When a player's vote changes a payoff-relevant event with positive
probability, it votes yes when the expected payoff from yes is greater than or
equal to the expected payoff from no. It votes no when the expected payoff
from no is strictly larger. At equality with positive probability of
relevance, the player votes yes.

When relevance has probability zero, the two actions are locally
outcome-equivalent holding the other strategies fixed. The assessment must
still record a complete behavioral profile, but that bookkeeping action is not
a substantive equilibrium class. Section \@ref(completion-discipline) explains
why local equivalence does not license simultaneous replacement of several
actions.

Only after maximizing the weak proposer's payoff, a proposal-payoff tie is
resolved by minimizing $H$'s expected payoff. If both the proposer and $H$
tie while downstream objects differ, the outcome remains a correspondence.

# Vote relevance

## Outcome signature

Fix player $i$'s ballot information set $I$, its action
$a_i\in\{Y,N\}$, the other simultaneous actions $a_{-i}$, and the type
when it remains relevant. Define the outcome signature

\[
\Phi_I(a_i,a_{-i},\theta)
=
\bigl(
A,
J_H,
O_H,
P,
S
\bigr),
\]

where:

- $A$ records approval or rejection;
- $J_H$ records inclusion or exclusion of $H$;
- $O_H$ records opt-out status;
- $P$ records the implemented payment vector and the destination of $y$;
- $S$ records the induced terminal outcome or the continuation
  outcome/interface under the frozen assessment.

If an action reaches a continuation, $S$ is the action-specific object
$\kappa_{C_2}(h_2)$, including its induced belief and payoff distribution.
It is not a raw successor label or a raw vote vector. Thus, in Round 1, two
actions that reach the same named $C_2$ class can remain relevant when they
produce different full histories and
$\kappa_{C_2}(h_2^Y)\ne\kappa_{C_2}(h_2^N)$.

The other-action profile is relevant for voter $i$ exactly when

\[
\Phi_I(Y,a_{-i},\theta)\ne
\Phi_I(N,a_{-i},\theta).
\]

This definition is conditional on the information set. It does not condition
on a simultaneous action the voter has not observed.

## Expected voting incentive

Let ${\cal R}_i(I)$ be the set of relevant pairs
$(\theta,a_{-i})$. The expected difference is

\[
\Delta_i(I)
=
\sum_{(\theta,a_{-i})\in{\cal R}_i(I)}
\Pr(\theta,a_{-i}\mid I)
\left[
u_i(Y,a_{-i};\theta,I)
-u_i(N,a_{-i};\theta,I)
\right].
\]

Every profile outside ${\cal R}_i(I)$ contributes exactly zero. If
$\Pr({\cal R}_i(I)\mid I)>0$, the prescribed response is yes when
$\Delta_i(I)\geq0$ and no when $\Delta_i(I)<0$. This is an expected
comparison over events the vote actually changes, not a blanket quota test.

For weak voters, relevant profiles normally occur at quota boundaries and can
change current payment versus continuation. For $H$, relevance must be
derived separately. Even when weak votes already meet a majority quota,
changing $H$'s ballot can change inclusion, payment $y$, opt-out, or the
continuation. Terminal opt-out status is part of the outcome signature even
when terminal monetary payoffs happen to coincide.

## Formal relevance registry

The authoritative registry is
`tables/pivotal_response_relevance_registry_v1.csv`. It separates weak and
hegemon ballots by round, rule, participation status, and information set.

```{r relevance-table, results='asis'}
compact_table(
  dplyr::select(
    relevance_registry,
    relevance_id,
    round,
    rule,
    h_status,
    player_class,
    structural_condition
  ),
  caption = "Gate 0 vote-relevance registry."
)
```

The complete 14-column registry remains authoritative in the CSV; the table
shows only the columns needed to audit the partition in print.

For a weak nonproposer $j$, let $w_{-j}\in\{1,\ldots,m-1\}$ count weak
yes votes other than $j$, including the proposer's automatic yes. The
current-approval structural toggle occurs exactly in the following cases:

- active-$H$ unanimity: $H=Y$ and $w_{-j}=m-1$; no weak toggle can
  restore unanimity after $H=N$;
- active-$H$ majority with $H=Y$: $w_{-j}=q-2$, an empty set when
  $N=3$ because then $q-2=0<w_{-j}$;
- active-$H$ majority with $H=N$: $w_{-j}=q-1$;
- weak-only majority: $w_{-j}=q-1$.

These quota cases do not exhaust Round-1 relevance when both actions fail.
The action-specific full histories must still be passed through the frozen
continuation interface. The verifier enumerates the structural cases for
$N=3,\ldots,60$, including the empty $N=3$ majority-with-$H=Y$ case. The
hegemon is analyzed separately for every $z\in\{1,\ldots,m\}$; no weak-voter
quota shortcut is applied to its inclusion, opt-out, payment, or continuation
comparison.

## Product-safe completion discipline {#completion-discipline}

Local irrelevance is not closed under changing several actions jointly.
Consider terminal unanimity with $N=4$. The proposer and $H$ vote yes and
the two weak nonproposers vote no. Holding the other no fixed, either weak
voter can switch alone without producing passage, so each action is locally
irrelevant. If both switch to yes, the proposal passes. Terminal weak-only
majority with $N=4,q=3$ has the same structure.

Accordingly, this Gate 0 does not impose a universal player-by-player
completion. Every assessment records an explicit full behavioral profile at
each proposal history. Two full completions may be quotiented only if they
induce identical approval, $H$ inclusion and opt-out, transfers, successor
state, relevant belief distribution, and payoff distribution at every
affected proposal and history.

Any later invariance claim must be product-safe: it must compare the induced
joint distributions, not merely verify that each vote is individually
irrelevant at one profile. If this proof fails, the different completions
remain distinct outcome classes or a correspondence. Proposal optimality must
also remain unchanged before a result is called completion-invariant.

# Coalitions, gifts, and proposal optimality

No coalition-size restriction is a primitive or an additional equilibrium
concept. Offers required to obtain support must be derived from ballot
incentives. **Gifts to outsiders remain feasible deviations** and must be
included in the recognized proposer's best-response problem.

Nor does local irrelevance generally imply zero outsider gifts or a minimum
winning representation. In terminal weak-only majority with $N=5,q=3$, an
on-path proposal may have the proposer plus all three other weak voters voting
yes. Each nonproposer is locally irrelevant against the other two yes votes,
and the proposal passes. At a different, zero-gift proposal history, all three
nonproposers may vote no; each unilateral switch remains locally irrelevant
because proposer plus one yes still falls short of $q$, and the proposal
fails. Deleting a gift changes the proposal information set and need not
preserve its completion or the proposer's payoff.

This finding is not confined to a knife-edge. An open set of feasible positive
gift vectors can accompany the all-yes completion, while a distinct deviated
proposal receives the all-no completion. Therefore a general zero-gift or
minimum-coalition conclusion is rejected in
this contract. A compact minimum-winning representation may be used only for a
proved subclass in which passage is secured or full completion invariance is
product-safe and proposer optimality removes every redundant transfer.
Otherwise the proof must retain unrestricted proposal-contingent completions,
positive gifts, oversized support, and any resulting correspondence.

# Histories and information sets

The authoritative information-set registry is
`tables/pivotal_response_information_sets_v1.csv`. It records what every mover
observes, what remains unobserved, when actions become public, the feasible
actions, successors, and the relevant belief object.

```{r information-set-table, results='asis'}
compact_table(
  dplyr::select(
    information_registry,
    information_set_id,
    stage,
    movers,
    unobserved,
    successors
  ),
  caption = "Gate 0 information sets and observations."
)
```

The complete 11-column information registry remains authoritative in the CSV.

The key safeguards are:

- weak ballot strategies cannot condition on the type or another current
  vote;
- $H$'s strategy can condition on its type but not on current weak votes;
- a Round-2 state retains the complete public Round-1 proposal and vote vector
  separately from its posterior;
- the weak-only Round-2 state requires no belief about the type for its payoff
  problem.

# Exhaustive transition registry

Let $z\in\{1,\ldots,m\}$ be the number of weak yes votes, including the
proposer's automatic yes. The authoritative transition registry is
`tables/pivotal_response_gate0_transitions_v1.csv`.

```{r transition-table, results='asis'}
compact_table(
  dplyr::select(
    transition_registry,
    transition_id,
    round,
    rule,
    h_status,
    h_vote,
    z_range,
    current_outcome,
    successor
  ),
  caption = "Gate 0 exhaustive ballot transitions in native payoff units."
)
```

The complete 16-column transition registry, including native payoffs and
discount notes, remains authoritative in the CSV.

The R verifier enumerates every admissible ballot tuple for
$N=3,\ldots,60$, both rules, both rounds, active-$H$ yes and no, and the
terminal weak-only majority state. Each tuple must map to exactly one of
PR01--PR16. The symbolic range $1,\ldots,q-2$ is empty at $q=2$, rather than
new histories.

No terminal outcome is promoted to a decision problem. In particular, after
an $H$-no under unanimity, the original quota is impossible and the branch
ends.

# Sufficient states

The authoritative sufficient-state registry is
`tables/pivotal_response_sufficient_states_v1.csv`.

```{r state-table, results='asis'}
compact_table(
  dplyr::select(
    state_registry,
    state_id,
    round,
    rule,
    h_status,
    stage,
    interface_role
  ),
  caption = "Gate 0 payoff- and belief-sufficient states."
)
```

The complete 12-column state registry remains authoritative in the CSV.

The registry separates pre-recognition, proposer-before-$s_i$, weak-ballot,
and $H$-ballot states. A weak-ballot key retains both voter $j$ and the
public proposal $s_i$; an $H$-ballot key retains $s_i$ and private
$\theta$. Each $C_2(h_2)$ imported by Round 1 is a pre-recognition
correspondence indexed by the full public $h_2$ and integrates the uniform
proposer draw. Histories may be compressed only after proving that feasible
actions, payoff consequences, beliefs, admissible strategies, and all future
continuations coincide.

# Continuation DAG and barriers

The machine-readable graph is
`model_redesign/pivotal_response_game_dag.json`. Its dependency direction is

```text
Gate 0
  -> {R2-U-active, R2-M-active, R2-M-weak-only}
  -> R2 batch review
  -> {R1-U, R1-M}
  -> R1 batch freeze
  -> {entry-U, entry-M}
  -> institutional comparison
  -> nonmigrating survival matrix.
```

Every brace is an antichain conditional on the preceding reviewed freeze.
The final graph has all nodes at `pass`; the generic checker validates the
execution orders and reports `Ready: none`.

The dependency-first schedule is:

| Batch | Nodes | Frozen closure | Status |
|:--|:--|:--|:--|
| Gate 0 | `gate0_contract` | bundle `6e28cb3...` | PASS, event 10 |
| R2 antichain | three terminal nodes | batch `00f84d42...` | PASS, event 25 |
| R1 antichain | `r1_unanimity`; `r1_majority` | batch `f4ac7b89...` | PASS, event 40 |
| Entry | `entry_unanimity`; `entry_majority` | batch `8817a9c5...` | PASS, event 53 |
| Comparison | comparison node | review `0acd9648...` | PASS, event 60 |
| Survival inventory | `v6_survival_matrix` | review `80c912ba...` | PASS, event 67; no migration |

Entry consumes its corresponding Round-1 payoff correspondence assessment by
assessment; a valid bound supports only a selection-free claim. Institutional
comparison uses the Cartesian product of complete entry assessments on the
full common existence domain. No scalar selection, rank matching, or
cross-rule randomization is inserted.

The following are terminal outcomes, not nodes: nonformation, current passage
with $H$, current weak-only passage, terminal active-$H$ failure, terminal
weak-only failure, and the quota-impossible unanimity outcome after opt-out.

# Continuation interfaces and native-time ledger

The versioning policy is in
`model_redesign/pivotal_response_interfaces/README.md`. Every closed
continuation states strategies, beliefs, payoff vectors or correspondences,
native payoff date, existence and multiplicity, boundaries, equality cases,
assumptions, checks, status, provenance, and dependency hashes. Candidate
interfaces are immutable snapshots; authority to consume them is supplied by
their reviewed batch bundle.

| Imported object | Native date | Consuming date | Permitted transformation | Status |
|:--|:--|:--|:--|:--|
| $C_{2,U}^{A}(h_2)$ | R2 | R1 | $\beta C_{2,U}^{A}(h_2)$ exactly once | frozen `f3ca4ebf...` |
| $C_{2,M}^{A}(h_2)$ | R2 | R1 | $\beta C_{2,M}^{A}(h_2)$ exactly once | frozen `a4ae73e9...` |
| $C_{2,M}^{WO}(h_2)$ | R2 | R1 weak payoffs only | $\beta C_{2,M}^{WO}(h_2)$ exactly once | frozen `e0ec6cd3...` |
| immediate $o_\theta$ after $H$-no in R1 | R1 | R1 | none | primitive |

Changing any frozen component changes its bundle hash, reopens that node, and
invalidates every descendant derivation and review. A predecessor may not fill
an interface gap locally.

The principal frozen objects and local check snapshots are:

| Object | SHA-256 prefix | Checks | Independent closure |
|:--|:--|--:|:--|
| Gate 0 | `6e28cb3f...` | 69/69 | two rereviews PASS |
| R2-U, active H | `f3ca4ebf...` | 18/18 | two reviews PASS |
| R2-M, active H | `a4ae73e9...` | 17/17 | two rereviews PASS |
| R2-M, weak-only | `e0ec6cd3...` | 13/13 | two reviews PASS |
| R2 batch | `00f84d42...` | 30/30 | frozen PASS |
| R1-U | `37aae1bf...` | 37/37 | two rereviews PASS |
| R1-M | `21c3a9dd...` | 23/23 | two rereviews PASS |
| R1 batch | `f4ac7b89...` | 24/24 | frozen PASS |
| Entry-U | `05e39ad2...` | 37/37 | two rereviews PASS |
| Entry-M | `4eb343ce...` | 24/24 | two reviews PASS |
| Entry batch | `8817a9c5...` | 27/27 | frozen PASS |
| Comparison | `cab69c5d...` | 42/42 | two reviews PASS |
| Comparison review | `0acd9648...` | 26/26 | frozen PASS |
| Survival candidate | `5278b14d...` | 43/43 | two reviews PASS |
| Survival review | `80c912ba...` | 24/24 | frozen PASS |

Full digests are machine-readable in the corresponding review bundles and in
the proof ledger. Prefixes are shown here only to keep the rendered table
legible.

# Proof ledger

```{r proof-ledger-table, results='asis'}
proof_ledger_display <- dplyr::select(
  proof_ledger,
  object,
  status,
  started_order,
  passed_order
)
proof_ledger_table <- knitr::kable(
  proof_ledger_display,
  caption = "Pivotal-response proof and execution ledger (display columns).",
  longtable = knitr::is_latex_output(),
  booktabs = knitr::is_latex_output()
)
if (knitr::is_latex_output()) {
  proof_ledger_table <- kableExtra::kable_styling(
    proof_ledger_table,
    font_size = 6
  )
}
proof_ledger_table
```

The ledger distinguishes analytic proof, numerical or mechanical checking,
rejected candidates, and final reviewed freezes. Numerical enumeration does
not substitute for the propositions below. The general zero-gift or
minimum-coalition conclusion remains `rejected` by the proposal-contingent
completion counterexample. Initial R2-M-active, R1-U, R1-M, and entry-U
candidates are also retained as rejected provenance; only their repaired,
independently approved hashes are consumed downstream.

# Independent Gate 0 review closure

The formal-model and adversarial read-only reviewers independently answered
all of the following questions affirmatively for the exact frozen bundle:

1. PR01--PR16 exhaust and partition the feasible ballot profiles;
2. the information sets prevent conditioning on unobserved current votes;
3. vote relevance includes approval, participation, opt-out, payments, and
   the induced terminal or assessed continuation outcome without using a raw
   successor label or collapsing $H$'s problem to quota pivotality;
4. equality is resolved as yes only when relevance has positive probability;
5. assessment-level completions and independent behavioral mixing are defined
   without overclaiming invariance from individual local irrelevance;
6. the N=4 joint-completion and N=5 gift counterexamples are handled
   correctly;
7. the unrestricted proposal space preserves gifts as deviations and avoids a
   general minimum-coalition claim;
8. the DAG enforces the Gate 0, R2-batch-to-R1, and
   R1-batch-to-entry barriers;
9. all Round-2 payoffs are native-date objects and discounting occurs exactly
   once upon import to Round 1;
10. no equilibrium result has entered the Gate 0 contract.

The first review round returned REPAIR on contract hash
`37bd422d...`.
After repair, both rereviews returned PASS with zero critical, major, or minor
findings on bundle
`sha256:6e28cb3f...`.
The closure and conditions are recorded in
`gate0_review_v1.json` and the independent Gate 0 review report.
No unresolved Gate 0 protocol choice remains. Any bundle-component change
reopens the gate and invalidates all descendants.

# How to read the frozen node proofs

The remaining chapters reproduce the detailed node proofs that generated the
reviewed interfaces. Proposition numbering is local to each named node. The
notes are immutable implementation snapshots, so a note may describe its own
interface as a review candidate; the batch and review bundles listed above are
the authoritative records that subsequently accepted those exact bytes. The
embedded notes do not select an equilibrium from a correspondence and do not
authorize manuscript migration.

# Round 2: terminal continuation correspondences

Each terminal problem is solved in Round-2 units. There is no `beta` inside a
terminal interface.

```{r include-r2-u, results='asis'}
include_node("pivotal_response_nodes/r2_unanimity_active_h_v1.md")
```

```{r include-r2-m-active, results='asis'}
include_node("pivotal_response_nodes/r2_majority_active_h_v1.md")
```

```{r include-r2-m-weak-only, results='asis'}
include_node("pivotal_response_nodes/r2_majority_weak_only_v1.md")
```

# Round 1: literal continuation import

The two Round-1 nodes import the frozen terminal elements by full public
history, type, identity, payoff vector, outcome distribution, belief, and
completion. A terminal payoff is multiplied by `beta` once, and only when it
enters a Round-1 continuation branch.

```{r include-r1-u, results='asis'}
include_node("pivotal_response_nodes/r1_unanimity_v1.md")
```

```{r include-r1-m, results='asis'}
include_node("pivotal_response_nodes/r1_majority_v1.md")
```

# Collective entry

Entry is evaluated assessment by assessment. The full bargaining assessment
is retained after the weak coalition forms or declines to form; a proposer
payoff projection never determines collective value by itself.

```{r include-entry-u, results='asis'}
include_node("pivotal_response_nodes/entry_unanimity_v1.md")
```

```{r include-entry-m, results='asis'}
include_node("pivotal_response_nodes/entry_majority_v1.md")
```

# Institutional comparison

The exact comparison object is the Cartesian product of whole entry
assessments under the two fixed rules. There is no primitive that couples,
rank-matches, or jointly selects assessments across rules.

```{r include-comparison, results='asis'}
include_node("pivotal_response_nodes/institutional_comparison_v1.md")
```

# Nonmigrating survival inventory

The claim-level matrix compares historical statements with the frozen current
objects only after the new derivation is complete. It is a provenance and
migration-warning artifact, not a source of equilibrium results and not an
authorization to edit a manuscript.

```{r include-survival, results='asis'}
include_node("pivotal_response_nodes/v6_survival_matrix_v1.md")
```

The authoritative matrix has the following final distribution:

```{r survival-status-summary, results='asis'}
survival_counts <- as.data.frame(table(survival_matrix$current_status))
names(survival_counts) <- c("Status", "Claims")
knitr::kable(
  survival_counts,
  caption = "Claim-level survival classifications after the independent rederivation."
)
```

# Consolidated results and limits

The rederivation changes the substantive object from a small list of scalar
candidate outcomes to a full PBE correspondence. Five conclusions organize
the result.

1. **Local voting discipline does not generate a global equilibrium
   selection.** When a vote changes a payoff-relevant outcome with positive
   probability, yes is prescribed exactly when its expected payoff is weakly
   higher. When relevance is zero, explicit completions remain part of the
   assessment. Individual irrelevance is not closed under joint changes, so
   different full-profile completions can sustain different outcomes.

2. **Terminal correspondences are generally broad.** Under unanimity with an
   active hegemon, coordinated failure is available from `N>=4`; under
   active-H majority, the recognized-proposer projection is `{1}` for
   `3<=N<=5` when `o_0>0`, becomes
   `[max{1-o_1,1-nu},1]` when `o_0=0`, and is `[0,1]` for `N>=6`.
   Weak-only majority is unique only at `N=3`; from `N>=4` it contains passage,
   failure, gifts, and identity-distinct allocations. None of these scalar
   projections replaces the full interface.

3. **Both Round-1 games have a PBE throughout the primitive domain.** The
   exact characterization is proposal- and history-indexed. Under majority at
   `N=3`, with `b=beta/2`,
   `A_3=1-b`, `C_3=(1-mu)(1-o_0)+mu*b`, and `D_3=1-o_1`, the proposer
   projection is `{max(A_3,C_3,D_3)}` if `o_0>0` and
   `[max(C_3,D_3),max(A_3,C_3,D_3)]` if `o_0=0`. A value-one majority PBE
   exists for every `N>=4`. Under unanimity, the familiar pooling, low-only,
   and continuation motifs survive only as verified subclasses; they do not
   exhaust the equilibrium correspondence.

4. **Entry is assessment-specific.** For either rule,

   \[
   G_R(\alpha,\mu)
   =\frac{(1-\mu)T_W^R(\alpha,0)+\mu T_W^R(\alpha,1)}{m},
   \]

   and the weak coalition forms exactly when `G_R>=chi`. In both rules,
   `0<=G_R<=1/m`. For `N>=4`, unanimity attains its lower endpoint zero and
   majority attains its upper endpoint `1/m`; the other general endpoints are
   not identified by the frozen interfaces.

5. **The institutional comparison is set-valued.** For a fixed assessment
   pair, formation costs are `[0,G_R]`, so pairwise U-in-M nesting holds
   exactly when `G_U<=G_M`. Possible-cost, guaranteed-cost, and universal
   cross-assessment nesting are different propositions. For `N>=4`, there is
   an M-only pair at every `0<chi<=1/m`, and universal M-in-U nesting is false.
   Universal U-in-M nesting remains pending. The hegemon's payoff is always at
   least its type-specific outside option, but this floor does not create a
   universal rule ranking; when both institutions form, the sign is generally
   indeterminate.

The proof therefore does **not** establish equilibrium uniqueness, a general
minimum winning coalition, zero gifts, an all-yes ballot, a global Round-1
cutoff, an exhaustive three-motif reduction, a universal formation nesting,
or universal hegemon dominance. Positive hegemon recognition, endogenous rule
choice, delayed or hybrid exit, and historical numerical applications are
outside this baseline. No result in this workspace has been migrated to the
target manuscript.
