---
title: "Clean Immediate-Opt-Out Baseline"
subtitle: "A Formal Laboratory for Informational Power Through Pivotality"
author: "Working formal appendix"
date: "2026-08-04"
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
    number_sections: true
    theme: readable
    code_folding: hide
  bookdown::pdf_document2:
    toc: true
    number_sections: true
    latex_engine: xelatex
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{pdflscape}
  - \usepackage{array}
geometry: margin=0.8in
fontsize: 10pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
history_path <- "../tables/clean_optout_gate0_histories_piH0.tsv"
gate0_histories <- utils::read.delim(
  history_path,
  check.names = FALSE,
  quote = "",
  comment.char = "",
  stringsAsFactors = FALSE
)
```

# Status and scope

This document is the autonomous formal laboratory for Goal 1. It replaces the
superseded delayed-continuation derivation with a clean benchmark in which

\[
\pi_H=0,\qquad b_0=b_1=0,
\]

and a no vote by type \(\theta\) of the hegemon is an immediate, irreversible
opt-out. In Round 1 that action pays \(o_\theta\) immediately and without
discounting. It does not also give the hegemon a discounted continuation
option.

The current version contains the complete Gate 0 contract. The independent
read-only audit returned PASS without substantive reservation on candidate
commit `fff4a35d1572c08bb2098cae9ad264a2eba80b41`. Analytical equilibrium
results are not presumed in the Gate 0 specification; they must be derived
separately by backward induction.

**Current formal status:** Gate 0 PASS; Phase 1 analytical rederivation
authorized.

**Manuscript boundary:** `formal_model_v6.Rmd` and `formal_model_v5.Rmd` are
reference objects only. This Goal does not edit either manuscript.

**Excluded from the baseline:** \(y+b_\theta\), a rejection payoff
\(\beta C_\theta\) chosen by a no vote in Round 1, the hybrid
\(\max\{o_\theta,\beta C_\theta\}\), the decomposition
\(t_\theta=d_\theta-b_\theta\), positive recognition probability for \(H\),
endogenous rule choice, sequential roll-call voting, and the archived
feasibility/C-B-R architecture. A discounted Round-2 payoff after \(H\) votes
yes and weak votes make the Round-1 ballot fail is part of the baseline; it is
not the payoff from opting out.

## Provenance

The pre-reset state is commit
`84a644128586d4f4d81c022cd7d3e09454ee8004`, recorded in the user-approved
annotated tag `pre-clean-optout-goal1-2026-08-03`. The initial SHA-256 of
`formal_model_v6.Rmd` is
`f18a999300c88e32bdc7542f3249cd258afd19a0c251c0b29c2c387a7131dbc1`.

# Gate 0: complete game contract

## Players, type, prior, and parameters

There are \(N\geq3\) states: one hegemon \(H\) and \(m=N-1\) weak states.
Nature draws

\[
\theta\in\{0,1\}
\]

before Round 1. The type persists. Only \(H\) observes \(\theta\). Weak states
share the prior

\[
\mu=\Pr(\theta=1)\in[0,1].
\]

The Bayesian derivations use \(\mu\in(0,1)\). The endpoints are treated as
degenerate-belief cases or limits, and not as positive-probability separating
histories containing both types.

The weak coalition has a fixed, known institutional surplus normalized to
one. The private type does not change this pie. Weak outside payoffs are zero.
The clean primitive domain is

\[
0\leq o_0<o_1\leq\bar y\leq1,
\qquad
\beta\in(0,1].
\tag{G0-domain}\label{G0-domain}
\]

If a current agreement includes \(H\), its payoff is \(y\). Thus
\(b_0=b_1=0\). The outside payoff \(o_\theta\) is external to the institutional
pie. It is never subtracted from weak surplus and is never paid together with
\(y\) in the same history.

The institutional rule \(R\in\{U,M\}\) is fixed exogenously for each
comparison. There is no endogenous rule-choice stage.

## Common knowledge and solution concept

The primitives, parameters, type distribution, recognition rule, contract
space, ballot protocol, quotas, implementation rule, payoff consequences, and
tie-breaking rules stated here are common knowledge.

The solution concept is perfect Bayesian equilibrium (PBE). An assessment
contains the weak states' collective formation strategy; every weak proposer's
proposal strategy at each public history where it is recognized; every active
non-proposer's behavioral voting strategy at each ballot information set;
\(H\)'s type-contingent behavioral voting strategy; and a belief system at
every information set where \(\theta\) remains payoff-relevant. Strategies are
sequentially rational given the belief system, and beliefs follow Bayes' rule
wherever possible. Every result below must state whether it concerns all PBEs,
pure-strategy PBEs, a constructed or selected PBE outcome, or a
selection-free bound. The solution concept does not select among multiple PBEs
unless one of the two stated tie rules applies.

## Horizon, recognition, and time units

The game has exactly two bargaining rounds, R1 and R2. There is no R3. In each
round that is reached, a new proposer is drawn publicly and independently of
the previous recognition, the type, and the public history. The baseline fixes

\[
\pi_H=0
\]

in every round. Each weak state is recognized with probability \(1/m\),
conditional on that round being reached.

Round 1 defines the payoff unit. A payoff realized in R2 is multiplied by
\(\beta\) in Round-1 units. A no vote by \(H\) in R1 pays \(o_\theta\) at once,
without \(\beta\). A no vote by \(H\) in R2 pays \(o_\theta\) at the R2 date,
or \(\beta o_\theta\) in Round-1 units.

R2 is terminal. If \(H\) is still active and its R2 ballot fails, then \(H\)
receives \(o_\theta\) at the R2 date, regardless of whether it voted yes or
no; weak states receive zero. This is the terminal outside payoff, not a
second payment. If \(H\) opted out in R1, it receives \(o_\theta\) exactly
once and receives no R2 flow.

## Formation

After Nature draws \(\theta\), but before the R1 recognition draw, the \(m\)
weak states make one collective, all-or-nothing formation decision. They do
not observe \(\theta\). The hegemon observes its type but takes no formation
action.

If the institution forms, every weak state pays an external sunk cost
\(\chi\geq0\). The cost does not reduce the unit pie. If the institution does
not form, every weak state receives zero and \(H\) receives \(o_\theta\).

For rule \(R\), the gross payoff relevant to collective formation is

\[
V_W^R(\mu)
=
\frac{1}{m}
\mathbb E\!\left[
  \sum_{j=1}^{m}u_j^R
  \;\middle|\;
  \text{formation}
\right].
\tag{G0-entry-value}\label{G0-entry-value}
\]

The expectation integrates over the type, recognition, public histories, and
any symmetric selection among weak coalition partners. Formation occurs iff
\(V_W^R(\mu)-\chi\geq0\); equality leads to formation. This collective
primitive is not an assertion that every realized weak state individually
prefers formation under an asymmetric coalition selection.

## Contract space and budget identities

Suppose weak state \(i\) is proposer. While \(H\) is active, a proposal is

\[
s_i=(y,(x_j)_{j\ne i}),
\]

where

\[
y\in[0,\bar y],\qquad x_j\geq0,
\]

and the proposer's residual if \(H\) participates is

\[
x_i^H
=1-y-\sum_{j\ne i}x_j\geq0.
\tag{G0-H-budget}\label{G0-H-budget}
\]

The same space applies in R1 and R2 and under both rules. The component \(y\)
is a public offer conditional on the observed participation of \(H\); there is
no hidden inclusion indicator. If \(H\) votes yes and the proposal passes, it
is included and receives \(y\). Every named weak-state offer \(x_j\) is paid
when the proposal passes, independently of that recipient's individual vote.

If \(H\) votes no under majority and the weak votes nevertheless reach the
quota, the agreement passes without \(H\). The conditional component \(y\) is
not burned and is not paid to \(H\). It is reabsorbed by the weak proposer, so

\[
x_i^{WO}
=1-\sum_{j\ne i}x_j
=x_i^H+y.
\tag{G0-WO-budget}\label{G0-WO-budget}
\]

The hegemon receives \(o_\theta\) externally. Hence both implemented budget
identities close at one:

\[
y+\sum_{j\ne i}x_j+x_i^H=1,
\qquad
\sum_{j\ne i}x_j+x_i^{WO}=1.
\tag{G0-budget-closure}\label{G0-budget-closure}
\]

After an earlier opt-out, only the \(m\) weak states remain active, every
proposal fixes \(y=0\), and the weak-only identity in
\eqref{G0-WO-budget} applies. If a ballot fails, no current \(x_j\) and no
current \(y\) is implemented.

These rules make \(y\) contingent only on publicly observed participation,
not on the private type.

## Common simultaneous ballot

The game is sequential and public across rounds: proposal, ballot, publication
of the complete vote vector and outcome, followed by any continuation. It is
not sequential within a ballot.

The proposer is counted as voting yes. All other active players observe the
proposal and public history and vote simultaneously. The individual votes are
revealed only after every ballot has been cast. The same ballot protocol is
used under both rules. Only the quota changes:

\[
Q_U=N,
\qquad
Q_M=q=\left\lfloor\frac{N}{2}\right\rfloor+1.
\tag{G0-quotas}\label{G0-quotas}
\]

Because \(N\geq3\), \(q\leq m\). While \(H\) remains active, the electorate
contains all \(N\) states. After opt-out, the active electorate contains only
the \(m\) weak states, but the majority quota remains the original \(q\); it
is not recalculated. The unanimity quota remains \(N\). Therefore an opt-out
makes future unanimity impossible and that branch ends immediately.

Players vote yes when indifferent. Strategies are functions of information
available before the ballot closes. No strategy may condition on a
simultaneous vote that has not yet been observed.

## Immediate opt-out

A no vote by type \(\theta\) of \(H\) is an immediate and irreversible
opt-out. No current or future agreement can then include \(H\). In R1 the
action pays \(o_\theta\) immediately. In R2 it pays \(o_\theta\) at the R2
date. There is no additional choice between opt-out and a discounted
continuation.

Under unanimity, an R1 no vote by \(H\) makes the quota impossible and ends the
branch. Under majority, if the weak yes votes already reach \(q\), their
current weak-only agreement passes. Otherwise the game proceeds to an R2
weak-only subgame with the original quota \(q\). A joint R1 rejection by \(H\)
and weak voters follows the same rules.

A yes vote by \(H\) is not an opt-out. If the R1 ballot then fails only because
weak votes do not reach the quota, \(H\) remains active and the game proceeds
to R2 under the same institutional rule.

## Exhaustive voting-history table

Let \(z\in\{1,\ldots,m\}\) denote the number of weak yes votes, including the
weak proposer's automatic yes vote. The tables in this section are generated
from the single versioned tab-separated source
`tables/clean_optout_gate0_histories_piH0.tsv`. HTML displays the complete
19-column table in a scroll box. To preserve legibility, PDF displays the same
rows and columns as four keyed panels. The concise entries use
`C_H2^R(theta,h2)` only for the legitimate continuation after an R1 yes vote by
\(H\) followed by a weak-caused failure. They never use that object as the
payoff from a no vote.

```{r gate0-history-table-html, eval=!knitr::is_latex_output(), results='asis'}
kableExtra::kbl(
  gate0_histories,
  format = "html",
  escape = TRUE,
  caption = "Gate 0 exhaustive voting histories, implementation rules, payoffs, beliefs, and continuations."
) |>
  kableExtra::kable_styling(font_size = 10) |>
  kableExtra::scroll_box(width = "100%", height = "650px")
```

```{r gate0-history-timing, eval=knitr::is_latex_output(), results='asis'}
history_timing <- dplyr::select(
  gate0_histories,
  history_id,
  round,
  rule,
  `active players`,
  quota,
  `quota reached?`,
  `H vote`
)
kableExtra::kbl(
  history_timing,
  format = "latex",
  booktabs = TRUE,
  longtable = TRUE,
  escape = TRUE,
  caption = "Gate 0 histories, panel A: timing, electorate, quota, and votes."
) |>
  kableExtra::kable_styling(
    latex_options = "repeat_header",
    font_size = 7
  ) |>
  kableExtra::column_spec(1, width = "0.55in") |>
  kableExtra::column_spec(2, width = "0.45in") |>
  kableExtra::column_spec(3, width = "0.35in") |>
  kableExtra::column_spec(4, width = "1.55in") |>
  kableExtra::column_spec(5, width = "0.75in") |>
  kableExtra::column_spec(6, width = "1.55in") |>
  kableExtra::column_spec(7, width = "0.85in") |>
  kableExtra::landscape()
```

```{r gate0-history-implementation, eval=knitr::is_latex_output(), results='asis'}
history_implementation <- dplyr::select(
  gate0_histories,
  history_id,
  `participation offer to H?`,
  `failure cause`,
  `H included?`,
  `H remains?`,
  `y implemented?`,
  `destination of y`
)
kableExtra::kbl(
  history_implementation,
  format = "latex",
  booktabs = TRUE,
  longtable = TRUE,
  escape = TRUE,
  caption = "Gate 0 histories, panel B: implementation and the disposition of the participation package."
) |>
  kableExtra::kable_styling(
    latex_options = "repeat_header",
    font_size = 7
  ) |>
  kableExtra::column_spec(1, width = "0.55in") |>
  kableExtra::column_spec(2, width = "1.25in") |>
  kableExtra::column_spec(3, width = "1.8in") |>
  kableExtra::column_spec(4, width = "0.8in") |>
  kableExtra::column_spec(5, width = "0.95in") |>
  kableExtra::column_spec(6, width = "1.15in") |>
  kableExtra::column_spec(7, width = "1.65in") |>
  kableExtra::landscape()
```

```{r gate0-history-payoffs, eval=knitr::is_latex_output(), results='asis'}
history_payoffs <- dplyr::select(
  gate0_histories,
  history_id,
  `H payoff`,
  `weak payoffs`,
  `discount factor`
)
kableExtra::kbl(
  history_payoffs,
  format = "latex",
  booktabs = TRUE,
  longtable = TRUE,
  escape = TRUE,
  caption = "Gate 0 histories, panel C: payoffs and time units."
) |>
  kableExtra::kable_styling(
    latex_options = "repeat_header",
    font_size = 8
  ) |>
  kableExtra::column_spec(1, width = "0.65in") |>
  kableExtra::column_spec(2, width = "1.75in") |>
  kableExtra::column_spec(3, width = "3.7in") |>
  kableExtra::column_spec(4, width = "2.1in") |>
  kableExtra::landscape()
```

```{r gate0-history-continuations, eval=knitr::is_latex_output(), results='asis'}
history_continuations <- dplyr::select(
  gate0_histories,
  history_id,
  `relevant posterior`,
  `terminal or continuation`,
  `next subgame`
)
kableExtra::kbl(
  history_continuations,
  format = "latex",
  booktabs = TRUE,
  longtable = TRUE,
  escape = TRUE,
  caption = "Gate 0 histories, panel D: beliefs and continuation subgames."
) |>
  kableExtra::kable_styling(
    latex_options = "repeat_header",
    font_size = 8
  ) |>
  kableExtra::column_spec(1, width = "0.65in") |>
  kableExtra::column_spec(2, width = "4.1in") |>
  kableExtra::column_spec(3, width = "2.05in") |>
  kableExtra::column_spec(4, width = "1.8in") |>
  kableExtra::landscape()
```

The R2-unanimity row with \(H\) already absent is included to document a
structurally unreachable class: the branch ends when the opt-out occurs because
the original quota \(N\) cannot be met by \(m\) active weak states. It is not a
latent R2 action.

# Information, beliefs, and incentive constraints

## Public histories and beliefs

Let \(h\) denote the public history before a proposal, let \(s_i\) be the
public proposal, and let

\[
I=(h,s_i)
\]

denote a ballot information set before simultaneous votes are cast. After the
ballot, the public history is

\[
h_2=(h,s_i,a_H,v_W,\text{quota outcome},\text{opt-out status}).
\]

Thus continuation strategies may condition on the full proposal and revealed
vote vector, not only on the posterior. Let \(\mu(h)\) be the belief before the
ballot. Whenever the denominator is positive, Bayes' rule requires

\[
\nu(h_2)
=
\frac{
  \mu(h)\Pr(h_2\mid\theta=1,h)
}{
  \mu(h)\Pr(h_2\mid\theta=1,h)
  +\{1-\mu(h)\}\Pr(h_2\mid\theta=0,h)
}.
\tag{G0-Bayes}\label{G0-Bayes}
\]

The posterior \(\nu(h_2)\) and a continuation payoff
\(C_{H,2}^R(\theta,h_2)\) are distinct objects. Histories with identical
posteriors can induce different continuation strategies and payoffs.

The baseline uses the expression **weak-vote-passive assessment**. It is a
maintained assessment of the simultaneous public ballot, not a refinement, D1,
the Intuitive Criterion, a sequential-equilibrium restriction, or a
characterization of all PBEs.

A unilateral vote deviation by a weak state is not treated as direct evidence
about \(\theta\), because weak states do not observe \(\theta\). Following a
weak-voter deviation, the posterior is the posterior implied by the public
proposal and \(H\)'s simultaneous type-contingent strategy and vote, not by the
action of an uninformed deviator. A separating vote by \(H\) can update beliefs
when the continuation keeps \(H\) active and the type remains payoff-relevant.

Every zero-probability continuation in which \(H\) remains active must state
the required off-path posterior and identify which result depends on it.
Off-path posteriors are components of an assessment, not primitives of the
game. They cannot be selected to delete an inconvenient branch. Once \(H\)
opts out, \(\theta\) is payoff-irrelevant for the weak-only continuation and no
posterior is needed to solve it.

## Hegemon's expected ballot IC

Fix an R1 ballot information set \(I=(h,s_i)\) and type \(\theta\). Let
\(v_W\) be the simultaneous vector of non-proposer weak votes, let
\(A_R^Y(v_W)\) indicate that the rule-\(R\) quota is reached when \(H\) votes
yes, and let \(\sigma_W(v_W\mid I)\) be the probability induced by the weak
strategies. The correct R1 payoff from voting yes is

\[
EU_H^R(Y\mid\theta,I)
=
\sum_{v_W}\sigma_W(v_W\mid I)
\left[
A_R^Y(v_W)y
+
\{1-A_R^Y(v_W)\}
\beta C_{H,2}^R\!\left(\theta,h_2(h,s_i,Y,v_W)\right)
\right].
\tag{G0-H-yes-R1}\label{G0-H-yes-R1}
\]

The payoff from voting no is always

\[
EU_H^R(N\mid\theta,I)=o_\theta.
\tag{G0-H-no-R1}\label{G0-H-no-R1}
\]

This equality holds whether a majority weak-only agreement passes immediately
or the weak states continue alone: \(H\) has opted out and receives no further
flow. Sequential rationality requires

\[
EU_H^R(Y\mid\theta,I)\geq o_\theta
\tag{G0-H-IC-R1}\label{G0-H-IC-R1}
\]

for a prescribed yes vote, with the reverse weak inequality for a prescribed
no vote.

At an R2 ballot information set \(I_2=(h_2,s_i)\), a failed ballot with \(H\)
still active pays the terminal outside payoff. Thus, in Round-1 units,

\[
EU_H^R(Y\mid\theta,I_2)
=
\beta\sum_{v_W}\sigma_W(v_W\mid I_2)
\left[
A_R^Y(v_W)y
+
\{1-A_R^Y(v_W)\}o_\theta
\right],
\qquad
EU_H^R(N\mid\theta,I_2)=\beta o_\theta.
\tag{G0-H-IC-R2}\label{G0-H-IC-R2}
\]

At an R1 information set where weak strategies make
\(A_R^Y(v_W)=1\) with probability one, the comparison directly reduces to

\[
y\geq o_\theta.
\tag{G0-direct-cutoff}\label{G0-direct-cutoff}
\]

Therefore the direct threshold \(y_\theta^*=o_\theta\) is derived at a
certain-implementation information set. It is not a global voting rule. If
weak-caused failure has positive probability, equation
\eqref{G0-H-IC-R1} must integrate current implementation and the legitimate
R2 continuation after a yes vote.

R2 has a special terminal algebra that does not change this scope statement.
If

\[
p_R(I_2)=\Pr\!\left(A_R^Y(v_W)=1\mid I_2\right),
\]

then equation \eqref{G0-H-IC-R2} implies

\[
EU_H^R(Y\mid\theta,I_2)-EU_H^R(N\mid\theta,I_2)
=
\beta p_R(I_2)(y-o_\theta).
\tag{G0-H-R2-difference}\label{G0-H-R2-difference}
\]

Consequently, every \(p_R(I_2)>0\) produces the same sign as
\(y-o_\theta\), while \(p_R(I_2)=0\) makes \(H\) indifferent for every \(y\)
and the voting tie rule prescribes yes. This is a consequence of terminal
failure payoffs, not a global primitive cutoff and not a license to replace
the R1 expected IC when failure has positive probability.

## Weak-voter approval IC

Fix a non-proposing weak voter \(j\) at ballot information set \(I=(h,s_i)\).
Its strategy is chosen before the simultaneous ballots are revealed. Let
\(u_j(a_j,v_{-j};I)\) be the payoff
specified by the history table: an implemented \(x_j\) if the current agreement
passes, the appropriate discounted continuation if R1 fails, or zero after
terminal failure. Its approval condition is

\[
\sum_{v_{-j}}\Pr(v_{-j}\mid I)
u_j(Y,v_{-j};I)
\geq
\sum_{v_{-j}}\Pr(v_{-j}\mid I)
u_j(N,v_{-j};I).
\tag{G0-W-IC}\label{G0-W-IC}
\]

The expectation includes \(H\)'s type-contingent simultaneous vote. A voter
cannot condition on that vote before the ballot closes. If voter \(j\) is
pivotal with probability one, equation \eqref{G0-W-IC} reduces to the usual
comparison between \(x_j\) and the continuation induced by a no vote. If its
vote never changes implementation or continuation, it is indifferent and votes
yes. Offers are implemented for named recipients whenever the agreement passes,
even if the recipient voted no.

## Tie-breaking

Every player votes yes when indifferent. Among proposals that maximize the
weak proposer's payoff, the selected proposal minimizes \(H\)'s expected
payoff. No third tie-break is available. If two proposals tie for both the weak
proposer and \(H\) but imply different downstream objects, the result must
remain set-valued or exclude that boundary by a stated strict theorem domain;
the proof may not select one silently.

# Gate 0 audit checklist

The implementer records the following mechanical assessment, which does not
substitute for the independent verdict.

| Requirement | Candidate evidence | Implementer status |
|---|---|---|
| Every required history has a payoff and continuation | Versioned history tables, rows G01--G21 | checked |
| No double payment to H | Contract identities and all H-no rows | checked |
| No reinclusion after opt-out | Rows G03, G04, G07--G09, G12, G13, and G16--G21 | checked |
| No discounted continuation as payoff from H-no in R1 | Equation \eqref{G0-H-no-R1} and table H-no rows | checked |
| Same simultaneous ballot under U and M | Equation \eqref{G0-quotas} | checked |
| Post-opt-out electorate and quotas defined | Common-ballot section and rows G19--G21 | checked |
| H-including and weak-only budgets close | Equations \eqref{G0-H-budget}--\eqref{G0-budget-closure} | checked |
| H IC integrates simultaneous weak votes | Equations \eqref{G0-H-yes-R1}--\eqref{G0-H-IC-R2} | checked |
| Beliefs stated whenever type remains relevant | Beliefs section and G02/G06 | checked |
| No equilibrium result presumed as a primitive | Status, primitives, and proof ledger | checked |

**Independent Gate 0 verdict:** **PASS without substantive reservation** on
commit `fff4a35d1572c08bb2098cae9ad264a2eba80b41`. The read-only reviewer found
no critical, major, or minor substantive issue after the repair round.

# Phase 1 placeholder

No equilibrium candidate, historical label, threshold ordering, No-Cheap-H
condition, entry nesting result, or institutional comparison has yet been
promoted under the clean contract. Gate 0 is closed, so the next derivation
must now proceed from terminal histories backward.

## Preliminary survival matrix

| Historical result | Current status before rederivation | Required evidence |
|---|---|---|
| Terminal low-only/pooling choice | pending | derive from R2 direct opt-out thresholds |
| R1 pooling label P | historical candidate only | derive H and weak-voter ICs from Gate 0 |
| R1 low-only label L | historical candidate only | derive H and weak-voter ICs from Gate 0 |
| R1 rejection/delay label R or D | historical candidate only | exhaust immediate-opt-out deviations; do not impose delay |
| Rejected-history reduction lemma | historical only | reprove or reject under the new H-no payoff |
| No-Cheap-H | historical only | rederive with primitive \(o_\theta\) |
| Majority no-screening | historical only | compare all payoff-relevant majority coalitions |
| Entry nesting | historical only | derive total weak payoffs under selected clean paths |
| H payoff comparison and classification | historical only | derive only after all preceding objects are proved |

# Proof ledger

| Object | Status | Notes |
|---|---|---|
| Gate -1 provenance | proved | Clean HEAD, user-approved annotated tag, and initial v6 hash recorded. |
| Gate 0 primitives and timing | specified and independently verified | PASS on candidate commit `fff4a35`. |
| Gate 0 contract space and budgets | specified and independently verified | PASS on candidate commit `fff4a35`. |
| Gate 0 voting histories | specified, mechanically checked, and independently verified | Source is the versioned TSV; 36/36 protocol checks pass. |
| Gate 0 beliefs and IC definitions | specified and independently verified | Expected IC explicitly integrates simultaneous weak votes; continuation uses full public history. |
| R2 unanimity | pending | No claim promoted. |
| R2 majority with H active | pending | No claim promoted. |
| R2 majority weak-only | pending | No claim promoted. |
| R1 unanimity | pending | Historical P/L/R labels are not primitives. |
| R1 majority and No-Cheap-H | pending | Must be rederived under primitive outside payoffs. |
| Entry, nesting, and classification | pending | Must follow the equilibrium derivation. |
| Delayed continuation and other architectures | extension/history only | Not part of this Goal's baseline. |
