---
title: "Clean Immediate-Opt-Out Baseline"
subtitle: "A Formal Laboratory for Informational Power Through Pivotality"
author: "Working formal appendix"
date: "2026-08-04"
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
    number_sections: true
    theme: readable
    code_folding: hide
  bookdown::pdf_document2:
    toc: true
    number_sections: true
    latex_engine: xelatex
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{pdflscape}
  - \usepackage{array}
geometry: margin=0.8in
fontsize: 10pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
history_path <- "../tables/clean_optout_gate0_histories_piH0.tsv"
gate0_histories <- utils::read.delim(
  history_path,
  check.names = FALSE,
  quote = "",
  comment.char = "",
  stringsAsFactors = FALSE
)
```

# Status and scope

This document is the autonomous formal laboratory for Goal 1. It replaces the
superseded delayed-continuation derivation with a clean benchmark in which

\[
\pi_H=0,\qquad b_0=b_1=0,
\]

and a no vote by type \(\theta\) of the hegemon is an immediate, irreversible
opt-out. In Round 1 that action pays \(o_\theta\) immediately and without
discounting. It does not also give the hegemon a discounted continuation
option.

The current version contains the complete Gate 0 contract. The independent
read-only audit returned PASS without substantive reservation on candidate
commit `fff4a35d1572c08bb2098cae9ad264a2eba80b41`. Analytical equilibrium
results are not presumed in the Gate 0 specification; they must be derived
separately by backward induction.

**Current formal status:** the extensive-form Gate 0 audit passed on the cited
commit. Phase 1 has rederived the clean game under weak PBE and found endogenous
existence restrictions, boundary multiplicity, and a conditional comparison
domain. The analytical candidate has passed the reproducible checks and is in
the independent repair-and-rereview loop. The formal-model, adversarial
game-theory, and R rereviews all returned PASS without substantive
reservation on commit `db52b2030b4c4e8e84c845a18ea04d4c2a27ab9c`; no
protocol change has been introduced.

**Manuscript boundary:** `formal_model_v6.Rmd` and `formal_model_v5.Rmd` are
reference objects only. This Goal does not edit either manuscript.

**Excluded from the baseline:** \(y+b_\theta\), a rejection payoff
\(\beta C_\theta\) chosen by a no vote in Round 1, the hybrid
\(\max\{o_\theta,\beta C_\theta\}\), the decomposition
\(t_\theta=d_\theta-b_\theta\), positive recognition probability for \(H\),
endogenous rule choice, sequential roll-call voting, and the archived
feasibility/C-B-R architecture. A discounted Round-2 payoff after \(H\) votes
yes and weak votes make the Round-1 ballot fail is part of the baseline; it is
not the payoff from opting out.

## Provenance

The pre-reset state is commit
`84a644128586d4f4d81c022cd7d3e09454ee8004`, recorded in the user-approved
annotated tag `pre-clean-optout-goal1-2026-08-03`. The initial SHA-256 of
`formal_model_v6.Rmd` is
`f18a999300c88e32bdc7542f3249cd258afd19a0c251c0b29c2c387a7131dbc1`.

# Gate 0: complete game contract

## Players, type, prior, and parameters

There are \(N\geq3\) states: one hegemon \(H\) and \(m=N-1\) weak states.
Nature draws

\[
\theta\in\{0,1\}
\]

before Round 1. The type persists. Only \(H\) observes \(\theta\). Weak states
share the prior

\[
\mu=\Pr(\theta=1)\in[0,1].
\]

The Bayesian derivations use \(\mu\in(0,1)\). The endpoints are treated below
as one-sided limits of that interior model, not as literal degenerate games.
This convention retains both type-contingent strategies while avoiding an
unstated restriction on beliefs about a zero-probability type.

The weak coalition has a fixed, known institutional surplus normalized to
one. The private type does not change this pie. Weak outside payoffs are zero.
The clean primitive domain is

\[
0\leq o_0<o_1\leq\bar y\leq1,
\qquad
\beta\in(0,1].
\tag{G0-domain}\label{G0-domain}
\]

If a current agreement includes \(H\), its payoff is \(y\). Thus
\(b_0=b_1=0\). The outside payoff \(o_\theta\) is external to the institutional
pie. It is never subtracted from weak surplus and is never paid together with
\(y\) in the same history.

The institutional rule \(R\in\{U,M\}\) is fixed exogenously for each
comparison. There is no endogenous rule-choice stage.

## Common knowledge and solution concept

The primitives, parameters, type distribution, recognition rule, contract
space, ballot protocol, quotas, implementation rule, payoff consequences, and
tie-breaking rules stated here are common knowledge.

The solution concept is perfect Bayesian equilibrium (PBE). An assessment
contains the weak states' collective formation strategy; every weak proposer's
proposal strategy at each public history where it is recognized; every active
non-proposer's behavioral voting strategy at each ballot information set;
\(H\)'s type-contingent behavioral voting strategy; and a belief system at
every information set where \(\theta\) remains payoff-relevant. Strategies are
sequentially rational given the belief system, and beliefs follow Bayes' rule
wherever possible. Every result below must state whether it concerns all PBEs,
pure-strategy PBEs, a constructed or selected PBE outcome, or a
selection-free bound. The solution concept does not select among multiple PBEs
unless one of the two stated tie rules applies.

## Horizon, recognition, and time units

The game has exactly two bargaining rounds, R1 and R2. There is no R3. In each
round that is reached, a new proposer is drawn publicly and independently of
the previous recognition, the type, and the public history. The baseline fixes

\[
\pi_H=0
\]

in every round. Each weak state is recognized with probability \(1/m\),
conditional on that round being reached.

Round 1 defines the payoff unit. A payoff realized in R2 is multiplied by
\(\beta\) in Round-1 units. A no vote by \(H\) in R1 pays \(o_\theta\) at once,
without \(\beta\). A no vote by \(H\) in R2 pays \(o_\theta\) at the R2 date,
or \(\beta o_\theta\) in Round-1 units.

R2 is terminal. If \(H\) is still active and its R2 ballot fails, then \(H\)
receives \(o_\theta\) at the R2 date, regardless of whether it voted yes or
no; weak states receive zero. This is the terminal outside payoff, not a
second payment. If \(H\) opted out in R1, it receives \(o_\theta\) exactly
once and receives no R2 flow.

## Formation

After Nature draws \(\theta\), but before the R1 recognition draw, the \(m\)
weak states make one collective, all-or-nothing formation decision. They do
not observe \(\theta\). The hegemon observes its type but takes no formation
action.

If the institution forms, every weak state pays an external sunk cost
\(\chi\geq0\). The cost does not reduce the unit pie. If the institution does
not form, every weak state receives zero and \(H\) receives \(o_\theta\).

For rule \(R\), the gross payoff relevant to collective formation is

\[
V_W^R(\mu)
=
\frac{1}{m}
\mathbb E\!\left[
  \sum_{j=1}^{m}u_j^R
  \;\middle|\;
  \text{formation}
\right].
\tag{G0-entry-value}\label{G0-entry-value}
\]

The expectation integrates over the type, recognition, public histories, and
any symmetric selection among weak coalition partners. Formation occurs iff
\(V_W^R(\mu)-\chi\geq0\); equality leads to formation. This collective
primitive is not an assertion that every realized weak state individually
prefers formation under an asymmetric coalition selection.

## Contract space and budget identities

Suppose weak state \(i\) is proposer. While \(H\) is active, a proposal is

\[
s_i=(y,(x_j)_{j\ne i}),
\]

where

\[
y\in[0,\bar y],\qquad x_j\geq0,
\]

and the proposer's residual if \(H\) participates is

\[
x_i^H
=1-y-\sum_{j\ne i}x_j\geq0.
\tag{G0-H-budget}\label{G0-H-budget}
\]

The same space applies in R1 and R2 and under both rules. The component \(y\)
is a public offer conditional on the observed participation of \(H\); there is
no hidden inclusion indicator. If \(H\) votes yes and the proposal passes, it
is included and receives \(y\). Every named weak-state offer \(x_j\) is paid
when the proposal passes, independently of that recipient's individual vote.

If \(H\) votes no under majority and the weak votes nevertheless reach the
quota, the agreement passes without \(H\). The conditional component \(y\) is
not burned and is not paid to \(H\). It is reabsorbed by the weak proposer, so

\[
x_i^{WO}
=1-\sum_{j\ne i}x_j
=x_i^H+y.
\tag{G0-WO-budget}\label{G0-WO-budget}
\]

The hegemon receives \(o_\theta\) externally. Hence both implemented budget
identities close at one:

\[
y+\sum_{j\ne i}x_j+x_i^H=1,
\qquad
\sum_{j\ne i}x_j+x_i^{WO}=1.
\tag{G0-budget-closure}\label{G0-budget-closure}
\]

After an earlier opt-out, only the \(m\) weak states remain active, every
proposal fixes \(y=0\), and the weak-only identity in
\eqref{G0-WO-budget} applies. If a ballot fails, no current \(x_j\) and no
current \(y\) is implemented.

These rules make \(y\) contingent only on publicly observed participation,
not on the private type.

## Common simultaneous ballot

The game is sequential and public across rounds: proposal, ballot, publication
of the complete vote vector and outcome, followed by any continuation. It is
not sequential within a ballot.

The proposer is counted as voting yes. All other active players observe the
proposal and public history and vote simultaneously. The individual votes are
revealed only after every ballot has been cast. The same ballot protocol is
used under both rules. Only the quota changes:

\[
Q_U=N,
\qquad
Q_M=q=\left\lfloor\frac{N}{2}\right\rfloor+1.
\tag{G0-quotas}\label{G0-quotas}
\]

Because \(N\geq3\), \(q\leq m\). While \(H\) remains active, the electorate
contains all \(N\) states. After opt-out, the active electorate contains only
the \(m\) weak states, but the majority quota remains the original \(q\); it
is not recalculated. The unanimity quota remains \(N\). Therefore an opt-out
makes future unanimity impossible and that branch ends immediately.

Players vote yes when indifferent. Strategies are functions of information
available before the ballot closes. No strategy may condition on a
simultaneous vote that has not yet been observed.

## Immediate opt-out

A no vote by type \(\theta\) of \(H\) is an immediate and irreversible
opt-out. No current or future agreement can then include \(H\). In R1 the
action pays \(o_\theta\) immediately. In R2 it pays \(o_\theta\) at the R2
date. There is no additional choice between opt-out and a discounted
continuation.

Under unanimity, an R1 no vote by \(H\) makes the quota impossible and ends the
branch. Under majority, if the weak yes votes already reach \(q\), their
current weak-only agreement passes. Otherwise the game proceeds to an R2
weak-only subgame with the original quota \(q\). A joint R1 rejection by \(H\)
and weak voters follows the same rules.

A yes vote by \(H\) is not an opt-out. If the R1 ballot then fails only because
weak votes do not reach the quota, \(H\) remains active and the game proceeds
to R2 under the same institutional rule.

## Exhaustive voting-history table

Let \(z\in\{1,\ldots,m\}\) denote the number of weak yes votes, including the
weak proposer's automatic yes vote. The tables in this section are generated
from the single versioned tab-separated source
`tables/clean_optout_gate0_histories_piH0.tsv`. HTML displays the complete
19-column table in a scroll box. To preserve legibility, PDF displays the same
rows and columns as four keyed panels. The concise entries use
`C_H2^R(theta,h2)` only for the legitimate continuation after an R1 yes vote by
\(H\) followed by a weak-caused failure. They never use that object as the
payoff from a no vote.

```{r gate0-history-table-html, eval=!knitr::is_latex_output(), results='asis'}
kableExtra::kbl(
  gate0_histories,
  format = "html",
  escape = TRUE,
  caption = "Gate 0 exhaustive voting histories, implementation rules, payoffs, beliefs, and continuations."
) |>
  kableExtra::kable_styling(font_size = 10) |>
  kableExtra::scroll_box(width = "100%", height = "650px")
```

```{r gate0-history-timing, eval=knitr::is_latex_output(), results='asis'}
history_timing <- dplyr::select(
  gate0_histories,
  history_id,
  round,
  rule,
  `active players`,
  quota,
  `quota reached?`,
  `H vote`
)
kableExtra::kbl(
  history_timing,
  format = "latex",
  booktabs = TRUE,
  longtable = TRUE,
  escape = TRUE,
  caption = "Gate 0 histories, panel A: timing, electorate, quota, and votes."
) |>
  kableExtra::kable_styling(
    latex_options = "repeat_header",
    font_size = 7
  ) |>
  kableExtra::column_spec(1, width = "0.55in") |>
  kableExtra::column_spec(2, width = "0.45in") |>
  kableExtra::column_spec(3, width = "0.35in") |>
  kableExtra::column_spec(4, width = "1.55in") |>
  kableExtra::column_spec(5, width = "0.75in") |>
  kableExtra::column_spec(6, width = "1.55in") |>
  kableExtra::column_spec(7, width = "0.85in") |>
  kableExtra::landscape()
```

```{r gate0-history-implementation, eval=knitr::is_latex_output(), results='asis'}
history_implementation <- dplyr::select(
  gate0_histories,
  history_id,
  `participation offer to H?`,
  `failure cause`,
  `H included?`,
  `H remains?`,
  `y implemented?`,
  `destination of y`
)
kableExtra::kbl(
  history_implementation,
  format = "latex",
  booktabs = TRUE,
  longtable = TRUE,
  escape = TRUE,
  caption = "Gate 0 histories, panel B: implementation and the disposition of the participation package."
) |>
  kableExtra::kable_styling(
    latex_options = "repeat_header",
    font_size = 7
  ) |>
  kableExtra::column_spec(1, width = "0.55in") |>
  kableExtra::column_spec(2, width = "1.25in") |>
  kableExtra::column_spec(3, width = "1.8in") |>
  kableExtra::column_spec(4, width = "0.8in") |>
  kableExtra::column_spec(5, width = "0.95in") |>
  kableExtra::column_spec(6, width = "1.15in") |>
  kableExtra::column_spec(7, width = "1.65in") |>
  kableExtra::landscape()
```

```{r gate0-history-payoffs, eval=knitr::is_latex_output(), results='asis'}
history_payoffs <- dplyr::select(
  gate0_histories,
  history_id,
  `H payoff`,
  `weak payoffs`,
  `discount factor`
)
kableExtra::kbl(
  history_payoffs,
  format = "latex",
  booktabs = TRUE,
  longtable = TRUE,
  escape = TRUE,
  caption = "Gate 0 histories, panel C: payoffs and time units."
) |>
  kableExtra::kable_styling(
    latex_options = "repeat_header",
    font_size = 8
  ) |>
  kableExtra::column_spec(1, width = "0.65in") |>
  kableExtra::column_spec(2, width = "1.75in") |>
  kableExtra::column_spec(3, width = "3.7in") |>
  kableExtra::column_spec(4, width = "2.1in") |>
  kableExtra::landscape()
```

```{r gate0-history-continuations, eval=knitr::is_latex_output(), results='asis'}
history_continuations <- dplyr::select(
  gate0_histories,
  history_id,
  `relevant posterior`,
  `terminal or continuation`,
  `next subgame`
)
kableExtra::kbl(
  history_continuations,
  format = "latex",
  booktabs = TRUE,
  longtable = TRUE,
  escape = TRUE,
  caption = "Gate 0 histories, panel D: beliefs and continuation subgames."
) |>
  kableExtra::kable_styling(
    latex_options = "repeat_header",
    font_size = 8
  ) |>
  kableExtra::column_spec(1, width = "0.65in") |>
  kableExtra::column_spec(2, width = "4.1in") |>
  kableExtra::column_spec(3, width = "2.05in") |>
  kableExtra::column_spec(4, width = "1.8in") |>
  kableExtra::landscape()
```

The R2-unanimity row with \(H\) already absent is included to document a
structurally unreachable class: the branch ends when the opt-out occurs because
the original quota \(N\) cannot be met by \(m\) active weak states. It is not a
latent R2 action.

# Information, beliefs, and incentive constraints

## Public histories and beliefs

Let \(h\) denote the public history before a proposal, let \(s_i\) be the
public proposal, and let

\[
I=(h,s_i)
\]

denote a ballot information set before simultaneous votes are cast. After the
ballot, the public history is

\[
h_2=(h,s_i,a_H,v_W,\text{quota outcome},\text{opt-out status}).
\]

Thus continuation strategies may condition on the full proposal and revealed
vote vector, not only on the posterior. Let \(\mu(h)\) be the belief before the
ballot. Whenever the denominator is positive, Bayes' rule requires

\[
\nu(h_2)
=
\frac{
  \mu(h)\Pr(h_2\mid\theta=1,h)
}{
  \mu(h)\Pr(h_2\mid\theta=1,h)
  +\{1-\mu(h)\}\Pr(h_2\mid\theta=0,h)
}.
\tag{G0-Bayes}\label{G0-Bayes}
\]

The posterior \(\nu(h_2)\) and a continuation payoff
\(C_{H,2}^R(\theta,h_2)\) are distinct objects. Histories with identical
posteriors can induce different continuation strategies and payoffs.

The baseline uses the expression **weak-vote-passive assessment**. It is a
maintained assessment of the simultaneous public ballot, not a refinement, D1,
the Intuitive Criterion, a sequential-equilibrium restriction, or a
characterization of all PBEs.

A unilateral vote deviation by a weak state is not treated as direct evidence
about \(\theta\), because weak states do not observe \(\theta\). Following a
weak-voter deviation, the posterior is the posterior implied by the public
proposal and \(H\)'s simultaneous type-contingent strategy and vote, not by the
action of an uninformed deviator. A separating vote by \(H\) can update beliefs
when the continuation keeps \(H\) active and the type remains payoff-relevant.

Every zero-probability continuation in which \(H\) remains active must state
the required off-path posterior and identify which result depends on it.
Off-path posteriors are components of an assessment, not primitives of the
game. They cannot be selected to delete an inconvenient branch. Once \(H\)
opts out, \(\theta\) is payoff-irrelevant for the weak-only continuation and no
posterior is needed to solve it.

## Hegemon's expected ballot IC

Fix an R1 ballot information set \(I=(h,s_i)\) and type \(\theta\). Let
\(v_W\) be the simultaneous vector of non-proposer weak votes, let
\(A_R^Y(v_W)\) indicate that the rule-\(R\) quota is reached when \(H\) votes
yes, and let \(\sigma_W(v_W\mid I)\) be the probability induced by the weak
strategies. The correct R1 payoff from voting yes is

\[
EU_H^R(Y\mid\theta,I)
=
\sum_{v_W}\sigma_W(v_W\mid I)
\left[
A_R^Y(v_W)y
+
\{1-A_R^Y(v_W)\}
\beta C_{H,2}^R\!\left(\theta,h_2(h,s_i,Y,v_W)\right)
\right].
\tag{G0-H-yes-R1}\label{G0-H-yes-R1}
\]

The payoff from voting no is always

\[
EU_H^R(N\mid\theta,I)=o_\theta.
\tag{G0-H-no-R1}\label{G0-H-no-R1}
\]

This equality holds whether a majority weak-only agreement passes immediately
or the weak states continue alone: \(H\) has opted out and receives no further
flow. Sequential rationality requires

\[
EU_H^R(Y\mid\theta,I)\geq o_\theta
\tag{G0-H-IC-R1}\label{G0-H-IC-R1}
\]

for a prescribed yes vote, with the reverse weak inequality for a prescribed
no vote.

At an R2 ballot information set \(I_2=(h_2,s_i)\), a failed ballot with \(H\)
still active pays the terminal outside payoff. Thus, in Round-1 units,

\[
EU_H^R(Y\mid\theta,I_2)
=
\beta\sum_{v_W}\sigma_W(v_W\mid I_2)
\left[
A_R^Y(v_W)y
+
\{1-A_R^Y(v_W)\}o_\theta
\right],
\qquad
EU_H^R(N\mid\theta,I_2)=\beta o_\theta.
\tag{G0-H-IC-R2}\label{G0-H-IC-R2}
\]

At an R1 information set where weak strategies make
\(A_R^Y(v_W)=1\) with probability one, the comparison directly reduces to

\[
y\geq o_\theta.
\tag{G0-direct-cutoff}\label{G0-direct-cutoff}
\]

Therefore the direct threshold \(y_\theta^*=o_\theta\) is derived at a
certain-implementation information set. It is not a global voting rule. If
weak-caused failure has positive probability, equation
\eqref{G0-H-IC-R1} must integrate current implementation and the legitimate
R2 continuation after a yes vote.

R2 has a special terminal algebra that does not change this scope statement.
If

\[
p_R(I_2)=\Pr\!\left(A_R^Y(v_W)=1\mid I_2\right),
\]

then equation \eqref{G0-H-IC-R2} implies

\[
EU_H^R(Y\mid\theta,I_2)-EU_H^R(N\mid\theta,I_2)
=
\beta p_R(I_2)(y-o_\theta).
\tag{G0-H-R2-difference}\label{G0-H-R2-difference}
\]

Consequently, every \(p_R(I_2)>0\) produces the same sign as
\(y-o_\theta\), while \(p_R(I_2)=0\) makes \(H\) indifferent for every \(y\)
and the voting tie rule prescribes yes. This is a consequence of terminal
failure payoffs, not a global primitive cutoff and not a license to replace
the R1 expected IC when failure has positive probability.

## Weak-voter approval IC

Fix a non-proposing weak voter \(j\) at ballot information set \(I=(h,s_i)\).
Its strategy is chosen before the simultaneous ballots are revealed. Let
\(u_j(a_j,v_{-j};I)\) be the payoff
specified by the history table: an implemented \(x_j\) if the current agreement
passes, the appropriate discounted continuation if R1 fails, or zero after
terminal failure. Its approval condition is

\[
\sum_{v_{-j}}\Pr(v_{-j}\mid I)
u_j(Y,v_{-j};I)
\geq
\sum_{v_{-j}}\Pr(v_{-j}\mid I)
u_j(N,v_{-j};I).
\tag{G0-W-IC}\label{G0-W-IC}
\]

The expectation includes \(H\)'s type-contingent simultaneous vote. A voter
cannot condition on that vote before the ballot closes. If voter \(j\) is
pivotal with probability one, equation \eqref{G0-W-IC} reduces to the usual
comparison between \(x_j\) and the continuation induced by a no vote. If its
vote never changes implementation or continuation, it is indifferent and votes
yes. Offers are implemented for named recipients whenever the agreement passes,
even if the recipient voted no.

## Tie-breaking

Every player votes yes when indifferent. Among proposals that maximize the
weak proposer's payoff, the selected proposal minimizes \(H\)'s expected
payoff. No third tie-break is available. If two proposals tie for both the weak
proposer and \(H\) but imply different downstream objects, the result must
remain set-valued or exclude that boundary by a stated strict theorem domain;
the proof may not select one silently.

# Gate 0 audit checklist

The implementer records the following mechanical assessment, which does not
substitute for the independent verdict.

| Requirement | Candidate evidence | Implementer status |
|---|---|---|
| Every required history has a payoff and continuation | Versioned history tables, rows G01--G21 | checked |
| No double payment to H | Contract identities and all H-no rows | checked |
| No reinclusion after opt-out | Rows G03, G04, G07--G09, G12, G13, and G16--G21 | checked |
| No discounted continuation as payoff from H-no in R1 | Equation \eqref{G0-H-no-R1} and table H-no rows | checked |
| Same simultaneous ballot under U and M | Equation \eqref{G0-quotas} | checked |
| Post-opt-out electorate and quotas defined | Common-ballot section and rows G19--G21 | checked |
| H-including and weak-only budgets close | Equations \eqref{G0-H-budget}--\eqref{G0-budget-closure} | checked |
| H IC integrates simultaneous weak votes | Equations \eqref{G0-H-yes-R1}--\eqref{G0-H-IC-R2} | checked |
| Beliefs stated whenever type remains relevant | Beliefs section and G02/G06 | checked |
| No equilibrium result presumed as a primitive | Status, primitives, and proof ledger | checked |

**Independent Gate 0 verdict:** **PASS without substantive reservation** on
commit `fff4a35`. The read-only reviewer found no critical, major, or minor
substantive issue after the repair round. Gate 0 fixes the game; it does not
presume that a PBE exists for every parameter vector. The backward-induction
results below therefore leave Gate 0 closed while reporting endogenous
existence restrictions faithfully.

# Analytical rederivation

## Belief discipline and theorem domains

The distinction between an on-path and a globally off-path proposal is
essential. A proposal made with positive probability by an uninformed weak
proposer preserves the current belief and subsequent positive-probability
histories obey Bayes' rule. After a globally zero-probability proposal, the
posterior is a component of the PBE assessment and must be declared. The
weak-vote-passive assessment prevents an uninformed unilateral vote deviation
from being treated as a direct signal, but it does not silently turn weak PBE
into sequential equilibrium.

In particular, when the proposal itself has probability zero, weak PBE does
not impose consistency between the belief held at its ballot and the posterior
at a later zero-probability continuation inside that same off-path subtree.
The constructions below therefore state both beliefs. Some completions use a
positive probability of the low type at the ballot and a pooling-selecting
posterior after a separating \(H\)-yes history. That pair is admissible under
weak PBE because both information sets are globally off path; it would require
additional justification under sequential equilibrium. No result below calls
this freedom a refinement or claims robustness to a stronger solution concept.

All payoffs from deviations are evaluated under the true prior \(\mu\), even
when voters at an off-path ballot hold a declared posterior \(\nu\). This
separation is what generates the security-value discontinuities below.

The main theorems use the **regular interior domain**

\[
\mu\in(0,1),\qquad
\beta\in(0,1),\qquad
0<o_0<o_1<1.
\tag{regular-domain}\label{regular-domain}
\]

The loci \(o_0=0\), \(o_1=1\), \(\beta=1\), and degenerate priors are stated
separately. They are not imported into the regular theorem by continuity.

## Round 2 under unanimity

In terminal R2 unanimity, every non-proposing weak voter approves. A unilateral
no either leaves failure unchanged or turns failure into an agreement paying
that voter a nonnegative offer; equality invokes the yes-vote tie rule. Hence
\(H\) is pivotal and the R2 IC in \eqref{G0-H-R2-difference} yields the direct
cutoff \(y\geq o_\theta\).

At posterior \(\nu\), define

\[
P=1-o_1,
\qquad
L(\nu)=(1-\nu)(1-o_0),
\qquad
S(\nu)=\max\{P,L(\nu)\}.
\tag{R2-U-value}\label{R2-U-value}
\]

A pooling proposal sets \(y=o_1\) and leaves weak total \(P\). A low-only
proposal sets \(y=o_0\) and leaves expected weak total \(L(\nu)\). At equality,
the proposal tie-break selects low-only because it gives \(H\) the lower
expected payoff. Thus

\[
\nu_2^*=\frac{o_1-o_0}{1-o_0}
\tag{R2-U-cutoff}\label{R2-U-cutoff}
\]

is the terminal belief cutoff, the total weak value is \(S(\nu)\), and each
weak state's ex ante R2 value is \(S(\nu)/m\). This is an all-PBE payoff result
for the terminal subgame; allocation among weak states can still depend on the
recognition draw.

## Round 2 under majority

In terminal R2 majority, weak approval is weakly dominant: failure pays zero,
an implemented named offer is nonnegative, and ties go to yes. All weak states
therefore approve. A weak proposer sets all named weak offers to zero, the
proposal passes even after \(H\)-no, and the proposer retains the unit pie.
The conclusion is the same in the active-\(H\) and post-opt-out weak-only
subgames:

\[
C_{W,2}^M=\frac1m,
\qquad
c:=\beta C_{W,2}^M=\frac\beta m,
\qquad
C_{H,2}^M(\theta)=o_\theta.
\tag{R2-M-values}\label{R2-M-values}
\]

The outside payoff remains external to the institutional pie. This proposition
is selection-free.

# Round 1 under unanimity

## On-path candidates

For the regular domain, write

\[
D_0=1-o_0,
\qquad
L=(1-\mu)D_0,
\qquad
\delta=\frac{\beta(m-1)}m,
\qquad
a=1-\delta.
\tag{U-notation}\label{U-notation}
\]

If pooling is on path, a weak deviation after \(H\)-yes preserves posterior
\(\mu\). Each non-proposer must receive \(\beta S(\mu)/m\), and the maximal
proposer payoff is

\[
\pi_P=P-\delta S(\mu).
\tag{U-onpath-P}\label{U-onpath-P}
\]

If low-only is on path, \(H\)-yes reveals the low type. Each non-proposer must
receive \(\beta D_0/m\), so

\[
\pi_L=aL.
\tag{U-onpath-L}\label{U-onpath-L}
\]

There is no high-branch continuation term in \(\pi_L\): high-type \(H\)-no is
immediate opt-out and ends unanimity. For \(o_0>0\) and \(\beta<1\), an
on-path weak-caused failure cannot be sustained because both types prefer
immediate opt-out once Bayes disciplines the continuation.

## Off-path completion lemma

Assume first that \(\beta o_1\geq o_0\), and define

\[
d=\frac{\beta P}{m},
\qquad
G_L=(1-\mu)(D_0-\delta P),
\qquad
G_P=aP.
\tag{U-security}\label{U-security}
\]

For each globally off-path proposal used in the upper-bound construction,
retain the prior ballot belief \(\rho=1-\mu\in(0,1)\) on the low type, because
the weak proposer is uninformed. After a history with
low-type \(H\)-yes and exactly one weak no, declare
\(\nu=1>\nu_2^*\), so terminal R2 selects pooling. These two beliefs are not
linked by Bayes because the proposal itself has probability zero. The low type
then obtains \(\beta o_1\) after weak-caused failure and is willing to vote
yes; the high type strictly opts out. The following completion holds every
proposal to an explicit upper bound.

1. If \(y<o_0\), prescribe all weak voters yes and both types of \(H\) no. The
   proposer receives zero.
2. If \(o_0\leq y<o_1\) and every \(x_j\geq d\), prescribe all weak voters
   yes, low-type \(H\) yes, and high-type \(H\) no. A pooling posterior after a
   unilateral weak deviation makes \(d\) the exact approval price. The best
   such deviation gives the proposer \(G_L\).
3. If \(y\geq o_1\) and every \(x_j\geq d\), prescribe all weak voters and both
   types of \(H\) yes. The best such deviation gives \(G_P\).
4. If some \(x_j<d\), designate one underpaid weak rejector, prescribe low-type
   \(H\) yes and high-type \(H\) no, and use the pooling continuation. The
   rejector strictly prefers continuation. The proposer's payoff is at most
   \((1-\mu)\beta P/m<G_L\).

Two or more weak no votes cannot occur because each rejector would be
nonpivotal and the tie rule would require yes. A nondegenerate ballot mixture
cannot enlarge the set: mixing requires indifference, where the same rule
assigns yes with probability one.

The reverse, guarantee direction is separate. Consider first

\[
y=o_0,\qquad x_j=d\quad\text{for every }j\ne i.
\tag{U-low-guarantee}\label{U-low-guarantee}
\]

In every sequentially rational completion, all weak voters must vote yes.
Two weak no votes make each rejector nonpivotal and therefore require yes by
the tie rule. With exactly one weak no, that voter can strictly prefer
rejection only if its terminal continuation exceeds \(d\), which requires R2
low-only rather than pooling. But low-only gives low-type \(H\) only
\(\beta o_0<o_0\) after voting yes, so that type instead opts out. The lone
weak rejector then becomes nonpivotal and must vote yes. With all weak voters
yes, low-type \(H\) votes yes at \(y=o_0\), high-type \(H\) votes no, and the
proposer obtains exactly \(G_L\).

The same argument applies to

\[
y=o_1,\qquad x_j=d\quad\text{for every }j\ne i.
\tag{U-pool-guarantee}\label{U-pool-guarantee}
\]

A lone weak rejector again needs more than \(d\), hence a low-only R2
continuation. Following weak-caused failure, low-type \(H\) then prefers
immediate opt-out because \(\beta o_0<o_0\), while high-type \(H\) prefers
immediate opt-out because \(\beta o_1<o_1\). Rejection therefore makes the
weak voter nonpivotal and contradicts the yes tie rule. All weak voters and
both types of \(H\) vote yes, so the proposer obtains exactly \(G_P\).
Thus \(G_L\) and \(G_P\) are genuine guarantees against every sequentially
rational completion, whereas the four-case construction above supplies the
matching upper bound.

If \(\beta o_1<o_0\), step 4 is impossible. The feasible proposal
\(y=o_1,x_j=0\) then has no sequentially rational ballot completion: all weak
yes makes each weak voter prefer the positive R2 continuation, two weak no
votes violate the tie rule, and with exactly one weak no both types of \(H\)
opt out. Thus the full game has no PBE on that open subdomain.

## Unanimity existence theorem

Suppose \eqref{regular-domain} holds. Define

\[
D_U=D_0-\delta P>0,
\qquad
\mu_E=1-\frac{aP}{D_U}.
\tag{U-existence-cutoff}\label{U-existence-cutoff}
\]

Then a global unanimity PBE exists if and only if

\[
\boxed{
\beta o_1\geq o_0
\quad\text{and}\quad
G_P>G_L
}
\tag{U-existence}\label{U-existence}
\]

or, equivalently, \(\beta o_1\geq o_0\) and \(\mu>\mu_E\).

**Proof.** The guarantee part of the lemma shows that the proposer can force
both \(G_L\) and \(G_P\) against every sequentially rational completion; its
declared-belief construction completes every proposal at no more than their
maximum. Yet the on-path low-only price is
\(\beta D_0/m>d\), so

\[
G_L-\pi_L=(1-\mu)\delta(D_0-P)>0.
\tag{U-shaving-gap}\label{U-shaving-gap}
\]

Low-only can therefore never be a regular PBE: the proposer shaves every weak
offer to \(d\) when the deviation is off path, but Bayes would restore the
higher price if that proposal became on path. Pooling can attain its off-path
security payoff only when \(P>L\), in which case
\(\pi_P=aP=G_P\). Direct algebra gives

\[
\mu_E-\nu_2^*
=
\frac{\delta P(D_0-P)}{D_0D_U}>0,
\tag{U-cutoff-order}\label{U-cutoff-order}
\]

so \(G_P>G_L\) already implies \(P>L\). If equality holds, the intermediate
deviation ties the proposer payoff and pays less to \(H\); the proposal
tie-break selects it, but its on-path approval price jumps back to
\(\beta D_0/m\). No fixed point exists at equality. Under the strict
inequality, pooling strictly dominates every completed deviation and the
completion lemma constructs a global PBE. \(\square\)

Every regular unanimity PBE in the existence domain has the same on-path
outcome:

\[
y=o_1,
\qquad
T_W^U=P,
\qquad
V_W^U=\frac{P}{m},
\qquad
E[u_H^U]=o_1.
\tag{U-regular-payoffs}\label{U-regular-payoffs}
\]

This is a pooling result, not a resurrection of the historical P/L/R
characterization and not a claim about parameter vectors outside
\eqref{U-existence}.

## Unanimity boundary results

Throughout this section \(\mu\in(0,1)\). The propositions cover every
single-parameter boundary and their intersections in the primitive domain.
As in the regular theorem, they characterize pure on-path proposal classes;
mixing is possible only over proposal ties that survive the stated tie-break.

### Zero low outside option

Let \(\beta<1\), \(o_0=0\), and \(0<o_1\leq1\). Terminal low-only gives each
weak state \(1/m\), so it can be used after a globally off-path proposal to
make the approval price \(\beta/m\). Define the exact off-path security value

\[
K_0=\max\{aL,P-\delta\}.
\tag{U-zero-low-security}\label{U-zero-low-security}
\]

The first term is the on-path low-only proposer payoff. The second is the
payoff from a pooling proposal that pays \(\beta/m\) to every non-proposer and
therefore forces their approval against every continuation when that proposal
is feasible; if \(P-\delta<0\), the second component is inactive in the
maximum.

**Proposition.** A global PBE always exists. If \(L\geq P\), low-only is the
selected accepted class and gives \(T_W^U=L\). If \(P>L\), canonical pooling
exists and gives \(T_W^U=P\); low-only also exists if and only if

\[
P-\delta\leq aL.
\tag{U-zero-low-multiplicity}\label{U-zero-low-multiplicity}
\]

At equality the proposal tie-break favors low-only. If \(P>L\), an overpaid
pooling class with \(o_1<y\leq\bar y\), \(T=1-y\), exists exactly when

\[
T\in[1-\bar y,P),
\qquad T-\delta P>K_0.
\tag{U-zero-low-overpay}\label{U-zero-low-overpay}
\]

**Proof.** Under low-only, a Bayes-consistent weak deviation leads to terminal
low-only, so each non-proposer receives \(\beta/m\) and the proposer obtains
\(aL\). Any off-path low-inclusion proposal can be held to that value by the
same continuation. A pooling proposal can force approval only by paying
\(\beta/m\) to every non-proposer, which yields \(P-\delta\). These two
constructions give both active sides of \(K_0\); when the pooling proposal is
infeasible, its negative component cannot bind. R1 rejection gives its original
proposer only \((1-\mu)\beta/m<aL\), since
\(a-\beta/m=1-\beta>0\).

If \(L\geq P\), then \(aL\geq P-\delta\), so low-only attains \(K_0\).
If \(P>L\), canonical pooling gives \(aP>K_0\). Low-only remains optimal
under a different off-path completion exactly when
\(aL\geq P-\delta\). At \(L=P\), low-only and pooling tie in proposer payoff,
but low-only pays less to \(H\) and is selected. An overpaid pooling proposal
gives the proposer \(T-\delta P\); it is sustainable precisely when it
strictly exceeds both security components. Equality is selected against it by
a deviation that pays less to \(H\). \(\square\)

Low-only gives weak total \(L\) and \(H\) expected payoff \(\mu o_1\);
canonical pooling gives weak total \(P\) and \(H\) payoff \(o_1\); overpaid
pooling gives weak total \(T\) and \(H\) payoff \(1-T\). Rejection is never
on path on this boundary.

### No terminal weak surplus

If \(\beta<1\), \(o_0>0\), \(o_1=1\), and \(\mu\in(0,1)\), no PBE exists.
Here \(P=0\). The proposal \(y=o_0,x_j=0\) forces all weak voters to approve:
two no votes make each rejector nonpivotal; with exactly one no, a positive
weak continuation requires terminal low-only, but then low-type \(H\) obtains
only \(\beta o_0<o_0\) after yes and instead opts out, making the rejector
nonpivotal; a zero continuation invokes the weak yes tie rule. Low-type \(H\)
therefore approves, high-type \(H\) opts out, and the deviation guarantees the
proposer \(L\). On-path low-only gives only \(aL<L\), pooling leaves zero weak
surplus, and R1 rejection gives the original proposer at most
\(\beta L/m<L\). No on-path class can attain the guarantee, so no fixed point
exists.

### No discounting

Let \(\beta=1\) and \(0\leq o_0<o_1\leq1\). Rejection/continuation is
admissible and gives the R1 proposer \(S(\mu)/m\) by uniform R2 recognition.
Because \(D_0>P\), terminal low-only provides the largest per-capita weak
continuation and can enforce the approval price \(D_0/m\) after any globally
off-path proposal. Define

\[
K_1=\max\left\{\frac Lm,
P-\frac{(m-1)D_0}{m}\right\}.
\tag{U-beta-one-security}\label{U-beta-one-security}
\]

If \(L\geq P\), low-only and rejection survive and pooling is selected
against. If \(P>L\), pooling and rejection survive; low-only additionally
survives if and only if

\[
P-\frac{(m-1)(1-o_0)}m\leq\frac{L}{m}.
\tag{U-beta-one-low}\label{U-beta-one-low}
\]

For \(P>L\), an overpaid pooling class with
\(o_1<y\leq\bar y\), \(T=1-y\), exists
exactly when

\[
T-\frac{m-1}{m}P>K_1.
\tag{U-beta-one-overpay}\label{U-beta-one-overpay}
\]

**Proof.** With no discounting, after weak-caused failure low-type \(H\) is
indifferent between terminal low-only and immediate opt-out, and high-type
\(H\) is indifferent between its terminal opt-out and immediate opt-out; the
yes tie rule sustains the punitive continuation. Low-only therefore gives and
can guarantee \(L/m\), while a pooling proposal that pays \(D_0/m\) to every
non-proposer guarantees the second component of \(K_1\) when feasible; a
negative component is inactive. A deliberately
rejected proposal with one strict weak rejector gives \(S/m\). If \(L\geq P\),
low-only and rejection attain that value; at equality they pay weakly less to
\(H\) than pooling. If \(P>L\), canonical pooling and rejection give \(P/m\),
which strictly exceeds \(K_1\); low-only survives exactly under
\eqref{U-beta-one-low}. The overpaid condition compares its on-path payoff
with both security components; equality is again selected against the larger
payment to \(H\). Mixtures over the payoff- and \(H\)-payoff-equivalent
low/rejection or pooling/rejection classes can survive. \(\square\)

Writing \(\bar o:=(1-\mu)o_0+\mu o_1\), when \(L\geq P\), low-only and
rejection give weak total \(L\) and \(H\) expected payoff \(\bar o\). When
\(P>L\), canonical pooling and rejection
give weak total \(P\) and \(H\) payoff \(o_1\). An overpaid pooling class gives
weak total \(T\) and \(H\) payoff \(1-T\).

No timing uniqueness is claimed. This is the only baseline locus on which the
historical rejection/delay label survives on path. The cases
\(o_0=0\) and \(o_1=1\) at \(\beta=1\) are included in this proposition, so
the boundary intersections are not filled in by continuity.

# Round 1 under majority

## Voting reduction and security value

For the regular domain, define

\[
q=\left\lfloor\frac N2\right\rfloor+1,
\qquad
k=q-1,
\qquad
c=\frac\beta m,
\tag{M-notation}\label{M-notation}
\]

and

\[
E=1-kc,
\qquad
B_M=(1-\mu)(1-o_0)+\mu c,
\qquad
P=1-o_1,
\qquad
F_M=\max\{E,B_M,P\}.
\tag{M-security}\label{M-security}
\]

For any non-proposing weak voter,

\[
EU_j(Y)-EU_j(N)
=(x_j-c)\Pr(j\text{ is pivotal}).
\tag{M-weak-IC}\label{M-weak-IC}
\]

If \(z\geq q-1\), \(H\)-yes implements and type \(\theta\) approves iff
\(y\geq o_\theta\). If \(z\leq q-2\), \(H\)-yes still fails and pays
\(\beta o_\theta<o_\theta\); both types opt out. Outside all-weak-yes, the only
regular pure ballot size is therefore \(z=q-1\).

The security value \(F_M\) is exact. Let
\(X=\sum_{j\ne i}x_j\), and let \(r\) count non-proposers offered at least
\(c\).

1. If \(y<o_0\) and \(r\leq k-1\), a \(z=q-1\) rejection completion pays the
   proposer \(c\); if \(r\geq k\), weak-only passage gives at most
   \(1-X\leq E\).
2. If \(o_0\leq y<o_1\), declare the off-path belief \(\nu=1\). With
   \(r\leq k-1\), choose \(z=q-1\); the true low state passes, the true high
   state continues, and the proposer receives at most \(B_M\). With
   \(r\geq k\), weak-only passage is forced and the proposer receives less
   than \(E\).
3. If \(y\geq o_1\), all weak voters are nonpivotal when all vote yes, and the
   proposer receives at most \(P\).

Conversely, paying \(c\) to \(k\) partners guarantees \(E\); the proposal
\(y=o_1,x=0\) guarantees \(P\); and \(y=o_0,x=0\) guarantees at least
\(B_M\). Hence no proposal system can give the proposer less than \(F_M\),
and off-path completions can hold every deviation to at most \(F_M\).

## Majority characterization by group size

For \(N=3,4\), the statements below classify every pure proposal in the
support. Mixtures over payoff-tied classes are admitted only when the proposal
tie-break does not select one support element; their expected payoffs remain
in the convex hull of the listed class payoffs.

### Three states

For \(N=3\), the exact proposer payoff is

\[
V_{\mathrm{prop}}^M=F_M.
\tag{M-N3-value}\label{M-N3-value}
\]

Exclusion is selected when \(E\) maximizes; low-only is selected when \(B_M\)
maximizes; pooling requires \(P>\max\{E,B_M\}\). A tie \(E=B_M\) remains
set-valued because both classes give \(H\) exactly its outside payoff. Their
weak totals are

\[
T_W^E=1,
\qquad
T_W^B=(1-\mu)(1-o_0)+\mu\beta,
\qquad
T_W^P=P.
\tag{M-N3-totals}\label{M-N3-totals}
\]

### Four states

For \(N=4\), the deviation \(y=o_0,x=0\) guarantees \(B_M\) when off path
but has no Bayes-consistent pure ballot on path. The only on-path classes that
can attain the security value are exclusion and pooling. Therefore

\[
\boxed{
\text{a majority PBE exists iff }
E\geq B_M\ \text{ or }\ P>B_M
}
\tag{M-N4-existence}\label{M-N4-existence}
\]

or, equivalently, it fails to exist iff \(B_M>E\) and \(B_M\geq P\).
Exclusion is selected iff \(E\geq\max\{B_M,P\}\), and pooling iff
\(P>\max\{B_M,E\}\). The boundary \(B_M=P>E\) also has no PBE: the
unattainable deviation ties pooling in proposer payoff and the tie-break favors
its lower payoff to \(H\).

### Five or more states

For \(N\geq5\), all-weak-yes is oversized. A majority PBE always exists, and
the exact set of proposer payoffs is

\[
V_{\mathrm{prop}}^M\in[F_M,1].
\tag{M-large-value}\label{M-large-value}
\]

Every regular pure on-path proposal belongs to one of three classes. If an
equilibrium mixes over proposal ties, every pure proposal in its support is in
the listed classes, and convex combinations preserve the payoff bounds below;
the classification does not exclude such mixing.

1. **Exclusion:** \(y<o_0\), with weak total \(T_W^M=1\) and \(H\) receiving
   \(o_\theta\). Redundant weak transfers support proposer payoffs throughout
   \([F_M,1]\).
2. **Separating current passage:** \(o_0\leq y<o_1\), low-type \(H\) votes
   yes and high-type \(H\) votes no, but the oversized weak coalition passes
   the agreement in both states. If the total named offers are \(X\),

   \[
   v=1-X-(1-\mu)y\geq F_M,
   \qquad
   T_W^M=1-(1-\mu)y,
   \qquad
   E[u_H^M]=(1-\mu)y+\mu o_1.
   \tag{M-large-separating}\label{M-large-separating}
   \]

   The canonical \(y=o_0,X=0\) outcome exists iff
   \(1-(1-\mu)o_0\geq E\). At a proposer-payoff tie, a positive rent
   \(y-o_0\) survives only when it also satisfies the stated proposal
   tie-break.
3. **Pooling:** \(y=o_1,X=0\), which exists iff
   \(P>\max\{E,B_M\}\). It gives \(T_W^M=P\) and \(E[u_H^M]=o_1\).

The interval in \eqref{M-large-value} has an explicit sufficiency
construction. For any \(v\in[F_M,1]\), select on path an exclusion proposal
with \(y=0\) and total named weak offers \(X=1-v\). Since \(N\geq5\) implies
\(q\leq m-1\), every non-proposer is nonpivotal at the all-weak-yes profile
and votes yes by the tie rule; the proposal passes weak-only and gives the
proposer \(v\). Complete every deviation with the security construction at no
more than \(F_M\). At \(v=F_M\), exclusion gives \(H\) its outside payoff,
the lowest payoff consistent with its opt-out, so the proposal tie-break does
not select a payoff-tied deviation against it. This proves sufficiency at the
floor as well as in the strict interior.

Low-only with delay is not a regular large-\(N\) PBE because its maximal
proposer payoff is strictly below \(B_M\). Across all three classes,

\[
F_M\leq T_W^M\leq1,
\qquad
\frac{F_M}{m}\leq V_W^M\leq\frac1m.
\tag{M-weak-bounds}\label{M-weak-bounds}
\]

The same weak-total bounds hold for \(N=3\) and for every existing \(N=4\)
PBE, using the class-specific totals above.

## No-Cheap-H changes

The historical condition \(o_0\geq c\) only compares exclusion with a
minimal-coalition low-only proposal. Under weak PBE, the binding off-path
comparison is \(B_M\) versus \(E\):

\[
B_M-E
=kc-o_0+\mu(o_0+c-1).
\tag{M-no-cheap-gap}\label{M-no-cheap-gap}
\]

Because the endpoint gaps are \(kc-o_0\) and \(qc-1\leq0\), the uniform
condition becomes

\[
o_0\geq kc=(q-1)\frac\beta m.
\tag{M-no-cheap-strong}\label{M-no-cheap-strong}
\]

It selects exclusion for \(N=3\), and guarantees existence with exclusion for
\(N=4\). For \(N\geq5\), even this stronger condition does not remove
separating current passage for sufficiently high priors. Majority
no-screening is therefore not an all-PBE result.

## Majority boundary warning

At \(\beta=1\), or at \(o_0=0\) with \(\beta<1\), \(H\)-yes can be optimal
even when \(z=q-2\) and the current ballot still fails. Rejection classes
reappear, and the regular security proof must be replaced. For \(N\geq4\),
the corresponding algebraic floors are, respectively,

\[
F_{M,1}=\max\left\{
1-kc,
(1-\mu)[1-o_0-(k-1)c]+\mu c,
1-o_1-(k-1)c
\right\},
\tag{M-beta-one-floor}\label{M-beta-one-floor}
\]

and

\[
F_{M,0}=\max\left\{
1-kc,
(1-\mu)[1-(k-1)c]+\mu c,
1-o_1-(k-1)c
\right\}.
\tag{M-zero-low-floor}\label{M-zero-low-floor}
\]

These formulas are boundary bounds, not substitutes for the regular theorem.
The common institutional comparison below is intentionally restricted to
\eqref{regular-domain}.

# Degenerate priors as one-sided limits

The primitive domain permits \(\mu=0,1\), but a literal degenerate game
requires an additional convention about strategies and beliefs for a
zero-probability type. The Goal permits endpoint cases or limits. We choose
the latter and do not silently add that convention.

Fix all other parameters in the regular domain. Let \(\mathcal E_R(\mu)\) be
the correspondence of regular on-path PBE outcome vectors under rule \(R\),
where

\[
z=(v_{\mathrm{prop}},T_W,V_W,E[u_H]),
\qquad V_W=T_W/m.
\]

For \(e\in\{0,1\}\), define the one-sided cluster correspondence

\[
\mathcal L_e^R=
\left\{
\lim_{n\to\infty}z_n:
\mu_n\in(0,1),\ \mu_n\to e,\
z_n\in\mathcal E_R(\mu_n)
\right\}.
\tag{endpoint-limit}\label{endpoint-limit}
\]

This object records limits of equilibrium outcomes and payoffs, not complete
assessments. An empty correspondence means that no sequence of regular PBEs
approaches the endpoint; it is not a nonexistence claim about a separately
specified degenerate game.

## Unanimity limits

Since

\[
\mu_E=\frac{D_0-P}{D_U}\in(0,1),
\]

the regular existence theorem gives

\[
\mathcal L_0^U=\varnothing,
\qquad
\mathcal L_1^U=
\begin{cases}
\{(aP,P,P/m,o_1)\},&\beta o_1\geq o_0,\\
\varnothing,&\beta o_1<o_0.
\end{cases}
\tag{U-endpoint-limits}\label{U-endpoint-limits}
\]

The weak inequality includes \(\beta o_1=o_0\) because the low type votes yes
when indifferent. The cutoff itself remains a nonexistence point even though
the pooling vector is the limit from \(\mu>\mu_E\). In terminal R2,
\(S(\mu)\to D_0\) as \(\mu\to0^+\) and \(S(\mu)\to P\) as
\(\mu\to1^-\).

## Majority limits

Write

\[
B_0^+=D_0,\qquad B_1^-=c,
\qquad F_0^+=\max\{E,D_0\},
\qquad F_1^-=\max\{E,P\}.
\tag{M-endpoint-floors}\label{M-endpoint-floors}
\]

Here \(E>c\) because \(E-c=1-q\beta/m>0\). The exact small-group
correspondences are:

| Group | \(\mu\to0^+\) | \(\mu\to1^-\) |
|---|---|---|
| \(N=3\) | low-only \((D_0,D_0,D_0/m,o_0)\) if \(D_0>E\); otherwise exclusion \((E,1,1/m,o_0)\) | pooling \((P,P,P/m,o_1)\) if \(P>E\); otherwise exclusion \((E,1,1/m,o_1)\) |
| \(N=4\) | empty if \(D_0>E\); otherwise exclusion \((E,1,1/m,o_0)\) | pooling \((P,P,P/m,o_1)\) if \(P>E\); otherwise exclusion \((E,1,1/m,o_1)\) |

At \(D_0=E\), \(B_M(\mu)-E=\mu(c-E)<0\), so the lateral tie belongs to
exclusion. At \(P=E\), pooling's strict condition fails, so that tie also
belongs to exclusion.

For \(N\geq5\), the lower-endpoint correspondence is

\[
\mathcal L_0^M=
\begin{cases}
\{(v,1,1/m,o_0):v\in[E,1]\},&D_0<E,\\[3pt]
\{(v,1,1/m,o_0):v\in[D_0,1]\}
\mathbin{\cup}
\{(D_0,t,t/m,o_0):t\in[D_0,1]\},&D_0\geq E.
\end{cases}
\tag{M-lower-limit}\label{M-lower-limit}
\]

The second set records mixtures over tied canonical separating and exclusion
proposals. Separating rents vanish in the limit because any surviving
separating proposal must have \(y\to o_0\). At the upper endpoint,

\[
\mathcal L_1^M=
\{(v,1,1/m,o_1):v\in[F_1^-,1]\}
\mathbin{\cup}
\begin{cases}
\{(P,P,P/m,o_1)\},&P>E,\\
\varnothing,&P\leq E.
\end{cases}
\tag{M-upper-limit}\label{M-upper-limit}
\]

**Proof.** The affine guarantee satisfies
\(B_M(\mu)\to D_0\) at \(0^+\) and \(B_M(\mu)\to c\) at \(1^-\).
Substitution in the exact \(N=3\) and \(N=4\) characterizations gives the two
small-group rows and their strict pooling conditions. For \(N\geq5\), the
exclusion construction yields the stated proposer intervals. Near zero, a
separating class has weak total
\(1-(1-\mu)y\); attaining the floor forces \(y\to o_0\) when
\(D_0\geq E\), while \(D_0<E\) excludes separation in a sufficiently small
right neighborhood. Mixtures at the tie fill the second segment in
\eqref{M-lower-limit}. Near one, every exclusion or separating class converges
to weak total one and \(H\) payoff \(o_1\); canonical pooling adds its distinct
limit exactly when \(P>E\). These are all the regular classes, so no other
cluster point exists. \(\square\)

Thus there is no common lower-endpoint comparison because
\(\mathcal L_0^U\) is empty. At the upper endpoint a common limit exists iff
\(\beta o_1\geq o_0\), and every majority weak-payoff limit is at least
\(P/m\). The entry nesting and three cost regions below therefore carry to
each common upper-endpoint selection. Every corresponding \(H\)-payoff limit
equals \(o_1\), so the interior weak advantage of unanimity for \(H\)
converges to equality.

# Conditional institutional comparison

## Common PBE domain

Let the regular unanimity conditions in \eqref{U-existence} hold. For
\(N=3\) or \(N\geq5\), this is already a common existence domain. For
\(N=4\), additionally impose \eqref{M-N4-existence}. These are derived
existence conditions, not ad hoc primitive restrictions.

On this common domain unanimity has the single on-path payoff in
\eqref{U-regular-payoffs}. Majority may be set-valued, but its security value
satisfies

\[
F_M=\max\{E,B_M,P\}\geq P.
\tag{comparison-floor}\label{comparison-floor}
\]

Combining this inequality with \eqref{M-weak-bounds} yields, in every majority
PBE,

\[
V_W^M\geq\frac{F_M}{m}
\geq\frac{P}{m}
=V_W^U.
\tag{entry-nesting}\label{entry-nesting}
\]

Thus collective formation sets satisfy

\[
\mathcal F_U\subseteq\mathcal F_M
\tag{formation-nesting}\label{formation-nesting}
\]

selection-free on the common domain. Given a selected majority payoff
\(v_M=V_W^M\):

1. if \(\chi\leq P/m\), both institutions form;
2. if \(P/m<\chi\leq v_M\), only majority forms;
3. if \(\chi>v_M\), neither forms.

There is no region in which only unanimity forms.

For \(H\), define its expected outside payoff

\[
\bar o=(1-\mu)o_0+\mu o_1.
\]

Every regular majority class satisfies

\[
\bar o\leq E[u_H^M]\leq o_1=E[u_H^U].
\tag{H-payoff-bounds}\label{H-payoff-bounds}
\]

Consequently, when both institutions form, unanimity weakly favors \(H\), with
equality only under majority pooling. When only majority forms, majority
weakly favors \(H\) relative to nonformation under unanimity. When neither
forms, \(H\) receives \(\bar o\) under both rules. This is a conditional
institutional classification over the common PBE domain, not global dominance
over the primitive parameter space.

# Reproducible verification

The analytical proofs are accompanied by separate R scripts. They check
identities, cutoff order, strict and weak inequalities, named counterexamples,
quota geometry, boundary formulas, and formation margins; they do not claim to
prove PBE by numerical enumeration.

- `verify_clean_optout_protocol_piH0.R`: 36/36 PASS.
- `verify_clean_optout_R2_piH0.R`: 8/8 PASS.
- `verify_clean_optout_R1_piH0.R`: 11/11 PASS.
- `verify_clean_optout_majority_piH0.R`: 18/18 PASS.
- `verify_clean_optout_entry_classification_piH0.R`: 10/10 PASS.
- `verify_clean_optout_boundaries_piH0.R`: 13/13 PASS.

Machine-readable outputs are stored in `tables/`; timestamped execution logs
are stored in `quality_reports/logs/`.

# Survival matrix

| Historical result | Clean immediate-opt-out status | Evidence |
|---|---|---|
| \(\pi_H=0\), \(b_\theta=0\), immediate \(o_\theta\) | survives as baseline primitives | Gate 0 contract and history table |
| Direct threshold \(y_\theta^*=o_\theta\) | survives locally | derived only under certain implementation and in terminal R2 |
| Terminal unanimity P/L choice | survives | \(S(\nu)=\max\{P,L(\nu)\}\) |
| Regular R1 pooling | survives conditionally | exists exactly under \eqref{U-existence} |
| Regular R1 low-only | disappears | payment-shaving gap \eqref{U-shaving-gap} |
| R1 rejection/delay | becomes a boundary result | survives at \(\beta=1\), not in the regular domain |
| Exact reduction to historical P/L/R | rejected | boundary overpay pooling and weak-PBE multiplicity add classes |
| Rejected-history reduction lemma | rejected for the clean game | its hypotheses do not survive the new opt-out timing and belief discontinuity |
| Global PBE existence over the primitive domain | rejected | unanimity and \(N=4\) majority have derived nonexistence regions |
| Majority no-screening | rejected as all-PBE | large groups admit separating current passage and rents |
| Historical No-Cheap-H \(o_0\geq\beta/m\) | changes | uniform majority condition is \eqref{M-no-cheap-strong} |
| Entry nesting | survives conditionally and selection-free | equation \eqref{formation-nesting} on the common PBE domain |
| H institutional ranking | survives conditionally | equation \eqref{H-payoff-bounds} plus formation status |
| Delayed continuation, hybrid exit, and \(t_\theta=d_\theta-b_\theta\) | extension/history only | excluded from every baseline payoff and theorem above |

# Proof ledger

| Object | Status | Notes |
|---|---|---|
| Gate -1 provenance | proved | Clean baseline commit, user-approved annotated tag, and initial v6 hash recorded. |
| Gate 0 extensive form | independently verified | PASS on candidate commit `fff4a35`; 36/36 protocol checks pass. |
| R2 unanimity | proved | Terminal low-only/pooling value and cutoff derived. |
| R2 majority | proved | Active and weak-only continuation values derived selection-free. |
| R1 unanimity, regular domain | proved; independently verified | Exact existence condition, nonexistence regions, and pooling payoff derived. |
| R1 unanimity, boundaries | proved or bounded as stated; independently verified | Zero outside option, no discounting, and zero terminal surplus are separated. |
| R1 majority, regular domain | proved; independently verified | Exact security value and \(N=3,4,\geq5\) characterization derived. |
| Majority boundaries | bounded; no comparison promoted | Separate floors recorded; regular theorem not extrapolated. |
| Degenerate priors | one-sided limits proved; independently verified | Cluster correspondences are derived without asserting PBEs of an underspecified literal endpoint game. |
| Entry, nesting, and classification | proved conditionally; independently verified | Selection-free on the explicitly derived common PBE domain. |
| Computational verification | passed; independent R PASS | Protocol 36/36, R2 8/8, R1 unanimity 11/11, majority 18/18, entry/classification 10/10, and boundary 13/13 checks pass (96/96 total). |
| Final formal and adversarial reviews | PASS without substantive reservation | Both read-only rereviews examined commit `db52b20`; no critical, major, or minor substantive finding remained. |
| Delayed continuation and other architectures | extension/history only | Not part of this Goal's clean baseline. |
